---
title: "801 Download Labware sample data"
author: "DHJ"
date: "5 5 2020"
output: html_document
---

This script reads sample data (including info on pooling of samples etc.) from Nivabasen     
  
NOTE:   
* A sample is in principle defined by AQUAMONITOR_CODE, SAMPLED_DATE, SPECIES, TISSUE, BIOTA_SAMPLENO   
    - However, BIOTA_SAMPLENO may be set to 0 for cod bile samples (for PAH matabolites- see 4b)  
* X_BULK_BIO (machine readable) and DESCRIPTION (human-readable) both describes pooling or not  
* SAMPLED_DATE is sometimes empty. Therefore, it is safe to collect data on the basis of PROSJEKT or ACCOUNT   
* Data may be from to types of accounts: Milkys and 'lange tidsserier'. The same sample may be registered in one or both (see 4c)      
    - However, X_BULK_BIO should be the same for both (see 4b-4c 4b below)    



Produces the R data file **Labware_samples_2019_[date].rds**which is used by scripts  
- 109_Adjust_for_fish_length.Rmd  
- 201_Make_big_excel_file.Rmd                 

- __NOTE: This script must be used on your personal PC, it doesn't work n Jupyterhub!__   
Prerequsites for this script to work on your PC:  
- Install 32-bit Oracle client    
- Get password to Nivabasen from JVE (see the readme file for niRvana)   
- Install package niRvana (https://github.com/NIVANorge/niRvana)  
  
  

## 1. Libraries  
```{r, message=FALSE}

library(dplyr)
library(niRvana)
library(lubridate)
library(knitr)       # kable()

```

```{r}

oracle_fix_norw_letters <- function(string){
  result = gsub("<f8>", "ø", string, fixed = TRUE)
  result = gsub("<c5>", "Å", result, fixed = TRUE)
  result = gsub("<e5>", "å", result, fixed = TRUE)
  result = gsub("<c6>", "Æ", result, fixed = TRUE)
  result = gsub("<e6>", "æ", result, fixed = TRUE)
  result
}

# Test
if (FALSE){
  oracle_fix_norw_letters("O 15330 ANA14 MILKYS - Milj<f8>overv<e5>king i kystomr<e5>dene")
  oracle_fix_norw_letters("23B B<f8>mlo north - Torsk lever 18 og 19")
}

```


## 2. Set username and pasword for Nivabasen  
```{r}

set_credentials()

```

## 3. Check samples from all years  
```{r}

# Get projects, but doesn't really make sense since the projects names fiffer from those in PROJECT in table LABWARE_CHECK_SAMPLE
if (FALSE){
  df_proj <- get_projects()
  grep("CEMP", df_proj$PROJECT_NAME, value = TRUE)
}

dat <- get_nivabase_data("select * from NIVADATABASE.LABWARE_CHECK_SAMPLE where upper(PROSJEKT) like '%MILKYS%'") %>%
  mutate(PROSJEKT = oracle_fix_norw_letters(PROSJEKT),
         DESCRIPTION = oracle_fix_norw_letters(DESCRIPTION)
  )
nrow(dat) # 3568

dat %>%
  count(ACCOUNT_NUMBER, PROSJEKT)

tail(dat$SAMPLED_DATE, 200)
sel <- grepl("2019", dat$TEXT_ID)
sum(sel)   # 656
mean(sel)  # 0.18

```

## 4. 2018 data  

### a. Get data  
```{r}

labware_2018_1 <- get_nivabase_data("select * from NIVADATABASE.LABWARE_CHECK_SAMPLE where ACCOUNT_NUMBER in (823,824)") %>%
  mutate(PROSJEKT = oracle_fix_norw_letters(PROSJEKT),
         DESCRIPTION = oracle_fix_norw_letters(DESCRIPTION)
         )

labware_2018_1 %>%
  count(ACCOUNT_NUMBER, PROSJEKT)


```

### b. Check X_BULK_BIO and DESCRIPTION   
Check whether data whether data given oth the same BIOTA_SAMPLENO    
* Small spelling difference in DESCRIPTION between the two ACCOUNTs for a blue mussel station  
    - So they are actually the same sample  
* Bile (NO: galle) samples are given with BIOTA_SAMPLENO = 0 (both are from MILKYS project)   
    - So they are actually a bunch of different samples  
```{r}

check1 <- labware_2018_1 %>%
  group_by(AQUAMONITOR_CODE, SAMPLED_DATE, SPECIES, TISSUE, BIOTA_SAMPLENO) %>%
  summarise(
    ACCOUNT_NUMBER_nunique = length(unique(ACCOUNT_NUMBER)),
    ACCOUNT_NUMBER_unique = paste(unique(ACCOUNT_NUMBER), collapse = " | "),
    X_BULK_BIO_nunique = length(unique(X_BULK_BIO)),
    DESCRIPTION_nunique = length(unique(DESCRIPTION)),
    X_BULK_BIO_unique = paste(unique(X_BULK_BIO), collapse = " | "),
    DESCRIPTION_unique = paste(unique(DESCRIPTION), collapse = " | "),
    .groups = "drop"
    )

cat("Number of unique X_BULK_BIO and ACCOUNT_NUMBER: \n")
xtabs(~X_BULK_BIO_nunique + ACCOUNT_NUMBER_nunique, check1)

cat("\n")
cat("Number of unique X_BULK_BIO and DESCRIPTION: \n")
xtabs(~X_BULK_BIO_nunique + DESCRIPTION_nunique, check1)

# Check the ones with >1 DESCRIPTION for each 'apparent' sample
check1 %>%
  filter(DESCRIPTION_nunique > 1)
  
# Check the ones with BIOTA_SAMPLENO == 0
# They are all bile samples  
check1 %>%
  filter(BIOTA_SAMPLENO == 0)

```
### c. Check 'lange tidsserier' account (824)     
33F and 52A are only on the 'lange tidsserier' account   
```{r}

xtabs(~ACCOUNT_NUMBER_unique, check1)

check1 %>%
  filter(ACCOUNT_NUMBER_unique == 824)
  
```
### d. Summarize data  
So we get only one line per 
We divide this in two parts:  
1. The mussel samples with only small spelling differences in DESCRIPTION  
2. The bile samples where BIOTA_SAMPLENO = 0, so actually only DESCRIPTION is telling us the sample number  
```{r}

# 1. The mussel samples with only small spelling differences in DESCRIPTION but same X_BULK_BIO  
# - we just pick the first DESCRIPTION and X_BULK_BIO  
labware_2018_1_part1 <- labware_2018_1 %>%
  filter(BIOTA_SAMPLENO != 0) %>%
  group_by(AQUAMONITOR_CODE, SAMPLED_DATE, SPECIES, TISSUE, BIOTA_SAMPLENO) %>%
  summarise(DESCRIPTION = first(DESCRIPTION),
            X_BULK_BIO= first(X_BULK_BIO),
            .groups = "drop")


# 2. The bile samples where BIOTA_SAMPLENO = 0, and therefore the samples can only be separated by
#   different DESCRIPTION    
labware_2018_1_part2 <- labware_2018_1 %>%
  filter(BIOTA_SAMPLENO %in% 0) %>%
  group_by(AQUAMONITOR_CODE, SAMPLED_DATE, SPECIES, TISSUE, BIOTA_SAMPLENO, DESCRIPTION) %>%
  summarise(X_BULK_BIO= first(X_BULK_BIO),
            .groups = "drop")

labware_2018_2 <- bind_rows(labware_2018_1_part1, labware_2018_1_part2)
  

```


## 5. 2019 data  

### a. Checking values of PROSJEKT  
* PROSJEKT including 'MILKYS': one found, 'MILKYS2019'   
* PROSJEKT including 'tidsserier': none found     
```{r}

check1 <- get_nivabase_data(
  "select * from NIVADATABASE.LABWARE_CHECK_SAMPLE where upper(PROSJEKT) like '%MILKYS%'")
check1 %>%
  count(ACCOUNT_NUMBER, PROSJEKT)
  

check2 <- get_nivabase_data(
  "select * from NIVADATABASE.LABWARE_CHECK_SAMPLE where lower(PROSJEKT) like '%tidsserier%'")
table(check2$PROSJEKT)
check2 %>%
  count(ACCOUNT_NUMBER, PROSJEKT)

# None fitting (only "O 190142; BL_ID - Lange tidsserier i kystvann; Bl<f8>tbunnsidentifisering")
```


### b. Get data    
```{r}

labware_2019 <- get_nivabase_data(
  "select * from NIVADATABASE.LABWARE_CHECK_SAMPLE where PROSJEKT like '%MILKYS2019%'") %>%
  mutate(PROSJEKT = oracle_fix_norw_letters(PROSJEKT),
         DESCRIPTION = oracle_fix_norw_letters(DESCRIPTION)
         )

labware_2019 %>%
  count(ACCOUNT_NUMBER, PROSJEKT)
#     ACCOUNT_NUMBER PROSJEKT                       n
# <dbl> <chr>                      <int>
#   1            925 O 19330;ANA19 - MILKYS2019   595

# head(df1, 2)


```



## 6. Save  
Data should then be uploaded to Jupyterhub Milkys2, folder 'Input_data', for use there       
```{r}

# Needs only be save if different from previously downloaded data

file_date <- Sys.time() %>% substr(1, 10)
# file_date

filename <- paste0("Labware_samples_2018_", file_date, ".rds")
saveRDS(labware_2018_2, paste0("Data_to_upload_to_Jupyterhub/", filename))
cat("File", sQuote(filename), "saved for 2018 \n\n")

filename <- paste0("Labware_samples_2019_", file_date, ".rds")
saveRDS(labware_2019, paste0("Data_to_upload_to_Jupyterhub/", filename))
cat("File", sQuote(filename), "saved for 2019 \n")

```


