---
title: "802 Get chemical data from NIVAbasen"
output: html_document
---

## Libraries
```{r setup, include=FALSE}
# install.packages("niRvana")
library(dplyr)
library(ggplot2)
library(lubridate)

# And then the niRvana package. If you need to install it, then run thee next line also:
# devtools::install_github("NIVANorge/niRvana")
library(niRvana)

source("802_Get_NIVAbasen_data_functions.R")

```


## Get new(er) data

### Give your username and password to R (only once per R session):
```{r}
set_credentials()
```
### Add encoding to 'get_nivabase_data' function  - until we have fisxed in in niRvana package - see issue there (https://github.com/NIVANorge/niRvana/issues/9)  
```{r}

get_nivabase_data <- function (sql = NULL, source = "NIVABASE") {
  library(svSocket)
  sock_port <- 8642L
  sock_con <- "sv_con"
  dataframe_temporary <<- data.frame()
  table_out <- "dataframe_temporary"
  pwd <- unserialize(sodium::simple_decrypt(db_pwd, db_privkey))
  startSocketServer(port = sock_port, server.name = "oracle_query_32", 
                    local = TRUE)
  expr <- "library(svSocket)"
  expr <- c(expr, sprintf("con <- DBI::dbConnect(odbc::odbc(), '%3$s', UID='%1$s', PWD='%2$s', encoding = 'ISO-8859-1')", 
                          db_username, pwd, source))
  expr <- c(expr, sprintf("%1$s <- DBI::dbGetQuery(con, \"%2$s\")", 
                          table_out, sql))
  expr <- c(expr, sprintf("%s <- socketConnection(port=%i)", 
                          sock_con, sock_port))
  expr <- c(expr, sprintf("evalServer(%s, %s, %s)", sock_con, 
                          table_out, table_out))
  expr <- c(expr, sprintf("close(%s)", sock_con))
  expr <- paste(expr, collapse = ";")
  prog <- file.path(R.home(), "bin", "i386", "Rscript.exe")
  system2(prog, args = c("-e", shQuote(expr)), stdout = NULL, 
          wait = TRUE, invisible = TRUE)
  stopSocketServer(port = sock_port)
  fact2char_df(dataframe_temporary)
}

```

### Get the list of projects
```{r}
df_projects <- get_projects()   # we call it 'df_projects' (the default name used by 'get_stations_from_project')
```

### Get a list of the stations in the CEMP_Biota project
```{r}
df_stations <- get_stations_from_project("CEMP_Biota", ignore.case = FALSE)
```

### Get all specimens collected at these stations (20 seconds or so)
```{r}
df_specimens <- get_specimens_from_stationdata(df_stations)
```

### Get species  
```{r}

taxoncode_id <- unique(df_specimens$TAXONOMY_CODE_ID)

df1 <- get_nivabase_selection(
  "TAXONOMY_CODE_ID, CODE, NIVA_TAXON_ID",
  "TAXONOMY_CODES",
  "TAXONOMY_CODE_ID",
  taxoncode_id
)

df2 <- get_nivabase_selection(
  "NIVA_TAXON_ID, LATIN_NAME",
  "TAXONOMY",
  "NIVA_TAXON_ID",
  df1$NIVA_TAXON_ID
)

df_species <- left_join(df1, df2, by = "NIVA_TAXON_ID") %>% 
  select(TAXONOMY_CODE_ID, LATIN_NAME)

```

## Get the data itself  
Get data frame of chemical data for all samples from the measurement year 2016 (30 seconds or so)
```{r, results = FALSE}

df_2020_temp <- get_biota_chemistry(
  years = 2020:2021,              # in case some specimens wrongly  
                                  #    have date_caught = 2020
  specimendata = df_specimens, 
  stationdata = df_stations,
  report_samples = TRUE) %>%
  select(-LATIN_NAME)             # we remove the 'LATIN_NAME', which comes from the BIOTA_SAMPES table
                                  #   and is empty

df_2020 <- df_2020_temp %>%
  left_join(df_species, by = "TAXONOMY_CODE_ID")

```
### Check species  
```{r}

table(addNA(df_2020$LATIN_NAME))

```


### Save 
```{r}
#
# 2020 data
#
# dir.create("Files_to_Jupyterhub_2020")

# Make timestamp (date stamp actually)
t <- Sys.time()
timestamp <- substr(t, 1, 10)
filename <- paste0("Files_to_Jupyterhub_2020/01_df_2020_notstandard_", timestamp, ".rds")

# Set to TRUE/FALSE depending on whether file should be saved or not  
if (FALSE){

  saveRDS(df_2020, file = filename)
  cat("2020 data saved as rds,", sQuote(filename), "\n")
  filename_txt <- sub("rds", "csv", filename)
  write.csv(df_2020, file = filename_txt, fileEncoding = "UTF-8")
  cat("2020 data saved as csv,", sQuote(filename_txt), "\n")

}

# Set to TRUE/FALSE depending on whether file should be read or not  
if (FALSE){
  # df_2020 <- readRDS(filename)
  df_2020 <- readRDS("Files_to_Jupyterhub_2020/01_df_2020_notstandard_2021-09-30.rds")
}

```


