---
title: "802 Get chemical data from NIVAbasen"
output: html_document
---

## Libraries
```{r setup, include=FALSE}
# install.packages("niRvana")

library(dplyr)
library(ggplot2)
library(lubridate)

# And then the niRvana package. If you need to install it, then run thee next line also:
# devtools::install_github("NIVANorge/niRvana")
library(niRvana)

source("802_Get_NIVAbasen_data_functions.R")

```


## Get new(er) data

### Give your username and password to R (only once per R session):
```{r}
set_credentials()
```

### Get the list of projects

```{r}

df_projects <- get_projects()   # we call it 'df_projects' (the default name used by 'get_stations_from_project')

```

### Get a list of the stations in the CEMP_Biota project

```{r}

df_stations <- get_stations_from_project("CEMP_Biota", ignore.case = FALSE)

```

### Get all specimens collected at these stations (20 seconds or so)
- Includes LATIN_NAME  
```{r}

df_specimens <- get_specimens_from_stationdata(df_stations, years = 2020:2021)

```

### Plot sample dates in 2020 (or later)    
```{r, fig.height=6, fig.width=7}

df_specimens %>%
  filter(year(DATE_CAUGHT) >= 2020) %>%
  ggplot(aes(DATE_CAUGHT)) +
  geom_histogram(binwidth = 3600*24*10) +   # 10 days
  scale_x_datetime(date_breaks = "1 month", date_labels = "%m %Y") +
  facet_wrap(vars(LATIN_NAME), ncol = 1, scales = "free_y")

```



## Get the data itself  
Get data frame of chemical data for all samples from the measurement year 2016 (30 seconds or so)  
- LATIN_NAME is included  
```{r, results = FALSE}

# debugonce(get_biota_chemistry)
# debugonce(get_samples_from_sampleid)

df_2020 <- get_biota_chemistry(
  years = 2020:2021,              # in case some specimens wrongly  
                                  #    have date_caught = 2020
  specimendata = df_specimens, 
  stationdata = df_stations,
  report_samples = TRUE)

```





### Save 
```{r}
#
# 2020 data
#
# dir.create("Files_to_Jupyterhub_2020")

# Make timestamp (date stamp actually)
t <- Sys.time()
timestamp <- substr(t, 1, 10)
filename <- paste0("Files_to_Jupyterhub_2020/01_df_2020_notstandard_", timestamp, ".rds")

# Set to TRUE/FALSE depending on whether file should be saved or not  
if (FALSE){

  saveRDS(df_2020, file = filename)
  cat("2020 data saved as rds,", sQuote(filename), "\n")
  filename_txt <- sub("rds", "csv", filename)
  write.csv(df_2020, file = filename_txt, fileEncoding = "UTF-8")
  cat("2020 data saved as csv,", sQuote(filename_txt), "\n")

}

# Set to TRUE/FALSE depending on whether file should be read or not  
if (FALSE){
  # df_2020 <- readRDS(filename)
  df_2020 <- readRDS("Files_to_Jupyterhub_2020/01_df_2020_notstandard_2021-09-30.rds")
}

```


