---
title: "71_Import_til_NIVAbasen_NILU_eider_2019data"
author: "DHJ"
date: "2 9 2019"
output: 
  html_document:
    keep_md: true
---

Make tables for inserting into NIVAbase   
  
For eider duck data from NILU (script 31)  


## 0. Libraries and functions
```{r}
library(dplyr)
library(purrr)
library(lubridate)
library(stringr)
library(tidyr)
# library(ggplot2)
library(niRvana)
library(safejoin)   # https://github.com/moodymudskipper/safejoin

source("71_Import_til_NIVAbasen_functions.R")

# "Restart" for lazy people that don't want to write the 
#    database password again (deletes all objects except database password/username)
if (FALSE){
  obj <- ls()
  obj <- obj[!obj %in% c("db_privkey", "db_pubkey", "db_pwd", "db_username")]
  rm(list = obj)
  rm(obj)
}


```

### For connectng to Nivabasen    
Store your username and password to R (only once per R session)  
```{r}

set_credentials()

# Check these codes (press F2) for reminder about what comes from which tables:
# niRvana::get_biota_chemistry
# niRvana::get_samples_from_sampleid

```

## 1. Read data


### a. Year
```{r}

selected_year <- 2019

```


### b. Get most recent data produced in Jupyterhub  
We only keep the data for the curret year
```{r}
# dir("Data")
# Data at sample level (chemical data, biological efect parameters and VDSI)
data_all <- readRDS(file = "Input_data_2019/101_data_updated_2020-08-05.rds")

dat_eider <- data_all %>%
  filter(MYEAR %in% selected_year & LATIN_NAME %in% "Somateria mollissima")

```


### c. Get samples in Labware file     
```{r}

# March-December data
sql <- paste(
  "select * from NIVADATABASE.LABWARE_CHECK_SAMPLE", 
  "where extract(YEAR from SAMPLED_DATE) =", selected_year, 
  "and extract(MONTH from SAMPLED_DATE) >= 3", 
  "and UPPER(PROSJEKT) like '%MILKYS%';")
sql
df_labware_01 <- get_nivabase_data(sql)

# January-February next year data
sql <- paste(
  "select * from NIVADATABASE.LABWARE_CHECK_SAMPLE", 
  "where extract(YEAR from SAMPLED_DATE) =", selected_year + 1, 
  "and extract(MONTH from SAMPLED_DATE) <= 2", 
  "and UPPER(PROSJEKT) like '%MILKYS%';")
df_labware_02 <- get_nivabase_data(sql)

df_samples <- bind_rows(df_labware_01, df_labware_02) %>%
  filter(AQUAMONITOR_CODE %in% dat$STATION_CODE) %>%
  # Add 'Specimen_no'
  mutate(DESCRIPTION = gsub("<f8>", "ø", DESCRIPTION, fixed = TRUE),
         DESCRIPTION = gsub("<c5>", "Å", DESCRIPTION, fixed = TRUE),
         DESCRIPTION = gsub("<e5>", "å", DESCRIPTION, fixed = TRUE),
         DESCRIPTION = gsub("<c6>", "Æ", DESCRIPTION, fixed = TRUE),
         DESCRIPTION = gsub("<e6>", "æ", DESCRIPTION, fixed = TRUE),
         DESCRIPTION2 = DESCRIPTION %>% stringr::str_sub(start = nchar(AQUAMONITOR_CODE) + 1),
         Specimen_no = stringr::str_extract(DESCRIPTION2, "[0-9]+") %>% as.numeric())


if (FALSE){
  df_samples %>%
    select(TEXT_ID, DESCRIPTION, SPECIES, TISSUE, Specimen_no, BIOTA_SAMPLENO, X_BULK_BIO) %>%
    arrange(TISSUE, BIOTA_SAMPLENO) %>%
    mutate()
}


```

### d. Plot samples  
```{r, fig.width=7, fig.height=3}

ggplot(df_samples, aes(BIOTA_SAMPLENO, Specimen_no, color = is.na(X_BULK_BIO))) +
  geom_point() +
  facet_grid(vars(AQUAMONITOR_CODE), vars(TISSUE))

```

### e. Get some generic lookup tables  
```{r}

df_tissue <- get_nivabase_data(
  "select TISSUE_ID,TISSUE_NAME from NIVADATABASE.BIOTA_TISSUE_TYPES")

df_taxon <- get_nivabase_selection(
  "NIVA_TAXON_ID, LATIN_NAME", 
  "TAXONOMY", 
  "LATIN_NAME",
  unique(dat_eider$LATIN_NAME), values_are_text = TRUE
  )

# Get a list of projects
df_projects <- get_projects()   # we call it 'df_projects' (the default name)

# Get a list of stations
df_stations <- get_stations_from_project("CEMP_Biota", ignore.case = FALSE)

# Get df_methods (NILU methods only)
df_methods <- get_nivabase_selection(
  "*", 
  "METHOD_DEFINITIONS", 
  "LABORATORY", 
  "NILU", values_are_text = TRUE)

```

### f. Read existing Nivabase data  
- % C, % N    
- Delta 13C, Delta 15N  
- PFAS  
```{r}

df2_specimens_allyrs <- get_specimens_from_stationdata(
  df_stations %>% filter(STATION_CODE %in% dat_eider$STATION_CODE))

# Get data frame of chemical results (30 seconds or so)
df_chem <- get_biota_chemistry(
  years = selected_year, 
  specimendata = df2_specimens_allyrs, 
  stationdata = df_stations,
  report_samples = TRUE)

cat("\nDownloaded", nrow(df_chem), "records\n")

xtabs(~NAME + STATION_CODE, df_chem)

```

### g. Define parameters that we can drop   
Baswd on table above  
```{r}

params_already_in_database <- c(
  "Delta13C", "Delta15N", "% C", "% N", "C/N", "PFAS", "PFBS", 
  "PFDcA", "PFHpA", "PFHxA", "PFHxS", "PFNA", "PFOA", "PFOS", "PFOSA", 
  "PFUdA")

# Sum parameters etc. that we fon't need to enter
params_to_drop <- c("BDE6S", "BDESS", "HBCDD", "CB_S7", "Sample")

```



## 2. Data to insert in NIVAbase
1. Eider duck data are from specimens/samples that ARE in the NIVAbase (for 2018 data, they were NOT)  
    - New records in BIOTA_SAMPLES only (SAMPLE_NO taken from data, get SAMPLE_ID from database)
    - Also note that although pairs of eggs and blood were unrelated (I think!),
      the same SPECIMEN_ID was giben for pairs of blood/egg samples  

2. NILU cod data (siloxans; D4,D5 and D6) are from specimens that ARE in the NIVAbase, 
but new samples (liver) are numbered corresponding to the muscle samples
   - Use existing record in BIOTA_SINGLE_SPECIMENS (existing SPECIMEN_ID)
   - New records in BIOTA_SAMPLES_SPECIMENS - relationship between SAMPLE_NO and SPECIMEN_ID
   should be the same as for muscle
   - New records in BIOTA_SAMPLES (new SAMPLE_ID, SAMPLE_NO taken from data)

3. Biological effects in cod - from specimens that ARE in the NIVAbase, with the 
exception of a single cod individual (Fish_no = 15, station = 53B). New samples (bile, blood,
liver fraction). Samples are numbered corresponding to the muscle samples (as NILU cod)
    - Use existing records in BIOTA_SINGLE_SPECIMENS (existing SPECIMEN_ID)  
    - New records in BIOTA_SAMPLES_SPECIMENS (new SAMPLE_ID, but existing SPECIMEN_ID)  
    - New records in BIOTA_SAMPLES (new SAMPLE_ID, SAMPLE_NO taken from data)  


### a. Check SAMPLE_NO2  
- Each tissue should have SAMPLE_NO2 = 1-15 (corresponds to individuals)   
- For 2017, this was a bit of a mess, with some SAMPLE_NO2 > 1000  
    - See code for 2017 if that is the case  
```{r}

# SAMPLE_NO2 are only 1-15 and 9064-9093
xtabs(~SAMPLE_NO2 + TISSUE_NAME, dat_eider)

```


### b. Make `df_lookup_sample_id`   
`SAMPLE_ID` already exists for some of the data (as they are already in the database)
- we use these existing records to find the correct `SAMPLE_ID`  
- which is used when we insert records in `BIOTA_CHEMISTY_VALUES`  
```{r}

df_lookup_sample_id <- df_chem %>%
  distinct(SAMPLE_NO, TISSUE_NAME, SAMPLE_ID) %>%
  arrange(TISSUE_NAME, SAMPLE_NO)

# Check 30 B (cod station) just to be sure this is the right thing  
if (FALSE){

    station_id_example <- subset(df_stations, STATION_CODE == "30B")$STATION_ID
  
  # Get 2017 data for check
  dat_example <- get_biota_chemistry(
    years = 2017, 
    specimendata = df2_specimens_allyrs %>% filter(STATION_ID == station_id_example),  # 30B
    stationdata = df_stations,
    report_samples = TRUE)
  
  
  xtabs(~SAMPLE_NO + TISSUE_NAME, dat_example)  
  xtabs(~SAMPLE_ID + TISSUE_NAME, dat_example)  
  
}

```


### c. Make lookups for the other ID variabless   
`BIOTA_SAMPLES_SPECIMENS_ID` and `SPECIMEN_ID` in table `BIOTA_SAMPLES_SPECIMENS`  
`SPECIMEN_NO` in table `BIOTA_SINGLE_SPECIMENS`  
NOT NEEDED! These tables don't have to be added to the database    
```{r}


if (FALSE){
  
  df_lookup_samples_specimens_id <- niRvana::get_nivabase_selection(
    columns = "SAMPLE_ID, SPECIMEN_ID, BIOTA_SAMPLES_SPECIMENS_ID", 
    table = "BIOTA_SAMPLES_SPECIMENS", 
    selection_column = "SAMPLE_ID", 
    selection_values = df_lookup_sample_id$SAMPLE_ID
  )
  
  df_lookup_specimen_no <- niRvana::get_nivabase_selection(
    columns = "STATION_ID, SPECIMEN_ID, SPECIMEN_NO, DATE_CAUGHT, TAXONOMY_CODE_ID", 
    table = "BIOTA_SAMPLES_SPECIMENS", 
    selection_column = "SAMPLE_ID", 
    selection_values = df_lookup_sample_id$SAMPLE_ID
  )
  
}


```




## 3. Eider duck data

### a. Set STATION_ID
```{r}

data_for_input <- dat_eider %>%
  safe_left_join(
    df_stations %>% select(STATION_CODE, STATION_ID),
    by = "STATION_CODE", na_matches = "never",
    check = "BMV")

```


### Tissue, check  
If 'Tissues are correct' we can just keep this column
```{r}

df_tissue <- get_nivabase_data("select TISSUE_ID,TISSUE_NAME from NIVADATABASE.BIOTA_TISSUE_TYPES")

cat("Tissues:\n")
unique(data_for_input$TISSUE_NAME)

check <- unique(data_for_input$TISSUE_NAME) %in% df_tissue$TISSUE_NAME
if (sum(!check) == 0){
  cat("Tissue names exist in NIVAbasen\n")
}

```


### Species, check  
If 'Species are correct' we can just keep this column
```{r}

cat("Species:\n")
unique(data_for_input$LATIN_NAME)

df_species <- get_nivabase_selection("SPECIES_ID,LATIN_NAME", "SPECIES", 
                                     "LATIN_NAME", unique(data_for_input$LATIN_NAME), values_are_text = TRUE)

check <- unique(data_for_input$LATIN_NAME) %in% df_species$LATIN_NAME
if (sum(!check) == 0){
  cat("Species names exist in NIVAbasen\n")
}

```



## 4. Eider data to insert, start  
Date is from   
`K:\Prosjekter\Sjøvann\JAMP\2017\opparbeiding biota\sjøfugl\Felttabell egg 2017 Ærfugl Kongsfjord MILKYS.xlsx`
```{r}


data_insert_1 <- data_for_input %>%
  filter(!PARAM %in% c(params_already_in_database, params_to_drop)) %>% 
  filter(!PARAM %in% params_to_drop) %>% 
  select(STATION_CODE, STATION_ID, LATIN_NAME, TISSUE_NAME, SAMPLE_NO2,
         PARAM, UNIT,
         VALUE_WW, FLAG1) %>%
  rename(SAMPLE_NO = SAMPLE_NO2,
         UNIT_orig = UNIT) %>%
  safe_left_join(
    df_chem %>% distinct(STATION_CODE, TISSUE_NAME, SAMPLE_DATE),
    by = c("STATION_CODE", "TISSUE_NAME"), na_matches = "never",
    check = "BMV"
  )

```


## 5. Make lookup table for parameter/unit  
To add correct parameter (NAME and UNIT) to the data  
  
### Parameters (NAME) used in 2018 data      
```{r}

if (FALSE)

  dat_eider_2018 %>%
    pull(NAME) %>% unique()


```

### Start lookup table  
Starting with NAME  
```{r}

df_paramnames1 <- data_insert_1 %>%
  distinct(PARAM, UNIT_orig) %>% #  View()       # get one record per PARAM, UNIT combination
  mutate(
    NAME = case_when(
      PARAM == "HCB"              ~ "Hexachlorobenzene (HCB)",
      PARAM == "2,2-Bisf.-F"      ~ "2,2-bisphenol F",
      PARAM == "4,4-Bisf.-F"      ~ "4,4-bisphenol F",
      PARAM == "4-T-NP"           ~ "4-tert-Nonylphenol",
      PARAM == "4-T-OP"           ~ "4-tert-Octylphenol",
      PARAM == "4,2-Bisf.-F"      ~ "4,2-bisphenol F",
      PARAM == "Dodekylphenol"    ~ "Dodecylphenol",
      PARAM == ""      ~ "",
      PARAM == ""      ~ "",
      substr(PARAM,1,3) == "BDE"  ~ sub("BDE", "BDE-", PARAM),
      substr(PARAM,1,2) == "CB"   ~ sub("CB", "PCB-", PARAM),
      substr(PARAM,1,4) == "HBCD" ~ paste0(tolower(substr(PARAM,5,5)), "-HBCD"),
      nchar(PARAM) == 2           ~ str_to_sentence(PARAM),
      TRUE ~ PARAM
    )
    )

# df_paramnames

test <- df_paramnames1 %>%
  left_join(df_methods, by = "NAME") %>%
  group_by(NAME) %>%
  summarise(n = sum(!is.na(UNIT)), .groups = "drop")
# test 

cat("No. of parameters not matching df_methods:", sum(test$n == 0), "\n")

# test %>% filter(n == 0)



```


### Lookup table, add UNIT 
```{r}

df_paramnames2 <- df_paramnames1 %>%
  mutate(
    UNIT = case_when(
      substr(PARAM,1,4) == "HBCD"               ~ "µg/kg",     # was "ng/g (w.w.)"
      PARAM == "HG"                             ~ "ng/g w.w.",
      PARAM %in% c("D4","D5","D6")              ~ "ng/g w.w.",
      PARAM %in% c("BPA","TBBPA","TBA", "PeCB") ~ "µg/kg",
      grepl("Sum.+CB$", PARAM)                  ~ "µg/kg",
      nchar(PARAM) == 2                         ~ "µg/g w.w.",  # All metals
      # PARAM %in% c("CO","SN")                   ~ "µg/g w.w.",  # Exceptions to the metals
      PARAM == "Fett"                           ~ "%",
      TRUE                                      ~ "ng/g w.w."
      
    )
    )

test <- df_paramnames2 %>%
  left_join(df_methods, by = c("NAME", "UNIT")) %>%
  group_by(NAME) %>%
  summarise(n = sum(!is.na(LABORATORY)), .groups = "drop")

cat("No. of parameters/units not matching df_methods:", sum(test$n == 0), "\n")

# test %>% filter(n == 0)

if (sum(test$n == 0) > 1){
  
  df_methods %>%
    filter(NAME %in% subset(test, n == 0)$NAME & toupper(MATRIX) == "BIOTA")

}


```

### For finding correct NAME and UNIT - if needed   
Modify and run the lines you need  
- Not needed in 2019, all 2017 parameters were found  
```{r}


if (FALSE) {
  
  get_nivabase_data("select * from NIVADATABASE.METHOD_DEFINITIONS where rownum < 4")
  # get_nivabase_data("select * from NIVADATABASE.METHOD_DEFINITIONS where rownum < 4") %>% colnames()
  
  df_methods %>%
    filter(grepl("bis", NAME, ignore.case = TRUE) & toupper(MATRIX) == "BIOTA")
  
}


```



### Lookup table, add multiplier  
For those cases (only 1) where original and new unit are not equivalent  
```{r}

if (FALSE){
  
  # Check combinations of UNIT_orig, UNIT manually
  df_paramnames2 %>%
    count(UNIT_orig, UNIT)
  
}

# Add multiplier
df_paramnames2 <- df_paramnames2 %>%
  mutate(Multiplier = 
           case_when(
             UNIT_orig == "MG_P_KG" & UNIT == "ng/g w.w." ~ 1000,
             TRUE                                         ~ 1
             )
         )

df_paramnames2 %>%
  count(UNIT_orig, UNIT, Multiplier)

```

## 6. Finish eider data to insert  
Check number of missing values  
```{r}

data_insert_2 <- data_insert_1 %>%
  safe_left_join(df_paramnames2, 
                 by = c("PARAM", "UNIT_orig"), na_matches = "never",
                 check = "BCV"
                 ) %>%
  mutate(VALUE = VALUE_WW*Multiplier)

# Check number of missing values
apply(is.na(data_insert_2), 2, sum)

# Check visually:
# data_insert_2
```

## 7. METHOD_DEFINITIONS  

### For reference - this year's Nivabasen data   
For instance, shows that SPECIES_ID is not used but TAXONOMY_CODE_ID is  
```{r}

if (FALSE)
  apply(!is.na(df_chem), 2, mean) %>% round(2)

```



### METHOD_DEFINITIONS   
Not needed in 2019  
```{r}

if (FALSE){
  n <- 2  
  new_methods <- tibble(
    NAME = c("Co", "Sn"),
    UNIT = rep("µg/g w.w.", n),
    LABORATORY = rep("NILU", n),
    METHOD_REF = rep("ICP-MS", n),
    MATRIX = rep("BIOTA", n),
    CAS = c(NA, NA),
    MATRIX_ID = rep(1, n),
  )
  
  new_methods
  
  
  ### METHOD_DEFINITIONS II: Make SQL for use in SQL developer  
  # Copy-paste into SQL developer
  # * Remember "commit;"!
  
  # make_sql_chemistry_values(1, data = dat_summ)
  
  sql_list <- 1:nrow(new_methods) %>% 
    map_chr(make_sql_methods, data = new_methods)
  cat("Number of records: \n")
  length(sql_list)  
  
  i <- 1:length(sql_list)
  sql <- paste(sql_list[i], collapse = ";\n")
  sql <- paste0(sql, ";\n")
  writeLines(sql, "clipboard-1024")   # copies SQLs to clipboard - go to SQL Developer and paste
  # "clipboard-1024" instead of "clipboard": increases avaliable
  #    for the clipboard
  
  cat("\nSQLs: \n")
  sql_list[1:2]
  
  
  
  ### Retrieve METHOD_ID by re-reading data    
  
  # Get df_methods (NILU methods only)
  df_methods <- get_nivabase_selection(
    "*", 
    "METHOD_DEFINITIONS", 
    "LABORATORY", 
    "NILU", values_are_text = TRUE)
  
  
  
}


```


## 8. BIOTA_CHEMISTRY_VALUES

 
### a. Make lookup table for METHOD_ID   
df_lookup_methodid   

```{r}

#
# The following will be used - seems to be the common ones used in 2016 data:  
#
# Co, Sn                  mg/kg         -     	      BIOTA
# Hg                      ng/g w.w.     ICP-MS 	      BIOTA
# Cu + the other metals   Âµg/g w.w.	    ICP-MS	      BIOTA
# PCB-118	                ng/g w.w.	    GC-MS         BIOTA
# BDE-47	                ng/g w.w.	    GC-MS	        BIOTA
# D4	                    ng/g w.w.	    CSR-LVI-GCMS	BIOTA
# a-HBCD	                ng/g (w.w.)   -             -
# Hexachlorobenzene (HCB)	ng/g w.w.	    GC-MS         -
# SCCP	                  ng/g w.w.	    GC-MS	        BIOTA
# Fett                    %             -             Biota 


# Try simple left join of the data
check1 <- data_insert_2 %>%
  count(NAME, UNIT) %>%
  left_join(df_methods %>%
              filter(toupper(MATRIX) %in% "BIOTA") %>%
              select(NAME, UNIT, METHOD_ID, METHOD_REF, MATRIX), 
            by = c("NAME", "UNIT")) %>%
  group_by(NAME, UNIT) %>%
  mutate(n = n())

# We may get a lot of duplicates
cat("Number of duplicates in df_methods, first check: \n")
sum(check1$n > 1)    # 130

# Visual check:
# check1

# Pick one method per NAME,UNIT 
df_lookup_methodid <- check1 %>%
  filter(n == 1 | !is.na(MATRIX))  # We pick those methods which either have no alternatives,
                                   #   or the ones with Biota/BIOTA in MATRIX
# Visual check:
# df_lookup_methodid

# Check for duplicates
check2 <- df_lookup_methodid %>%
  count(NAME, UNIT, wt = 1)
# We get no duplicates
cat("Number of duplicates in df_methods, second check: \n")
sum(check2$n > 1)      # should be zero

# Visual check:
# check1

# Check the othRr way (that all NAME/UNIT are repesented)
check3 <- data_insert_2 %>%
  left_join(df_lookup_methodid %>% select(NAME, UNIT, METHOD_ID, METHOD_REF, MATRIX),
            by = c("NAME", "UNIT"))

cat("Number of parameter/unit combinations not found: \n")
sum(is.na(check3$METHOD_ID))  # should be zero

if (sum(is.na(check3$METHOD_ID)) > 0)  # should be zero
  check3 %>% filter(is.na(check3$METHOD_ID)) %>% count(PARAM, NAME)

```

### b. Start `biota_chemistry_values_eider`   
By selecting columns and adding `METHOD_ID`   
```{r}
# df <- get_nivabase_data("select * from NIVADATABASE.BIOTA_CHEMISTRY_VALUES where rownum < 4")
# df
# df %>% colnames() %>% dput()

# "VALUE_ID"              - Let the database decide
# "SAMPLE_ID"             - From the database, after BIOTA_SAMPLES have been inserted
# "METHOD_ID"             - Lookup based on NAME and UNIT
# "VALUE"                 - From data
# "FLAG1"                 - From data
# "FLAG2"                 - NA
# "ENTERED_BY"            - DHJ
# "ENTERED_DATE"          - date, see above
# "REMARK"                - NA
# "DETECTION_LIMIT"       - NA
# "UNCERTAINTY"           - NA
# "QUANTIFICATION_LIMIT"  - NA
# "APPROVED"              - NA?

biota_chemistry_values1 <- data_insert_2 %>%
  safe_left_join(df_lookup_methodid, by = c("NAME", "UNIT"), 
                 check = "VM", na_matches = "never")

biota_chemistry_values1 %>%
  select(STATION_CODE, TISSUE_NAME, SAMPLE_NO, METHOD_ID, VALUE, FLAG1)

```

### c. Add SAMPLE_ID  
```{r}

biota_chemistry_values2 <- biota_chemistry_values1 %>%
  safe_left_join(df_lookup_sample_id, by = c("TISSUE_NAME", "SAMPLE_NO"), 
                 check = "BCVMn", na_matches = "never")

biota_chemistry_values2$SAMPLE_ID %>% unique() %>% sort()

cat("\n")
cat("Number of SAMPLE_ID lacking (must be zero): \n")
sum(is.na(biota_chemistry_values2$SAMPLE_ID))

```


### d. Make SQLs  
And copies them to clipboard  
PASTE INTO SQL DEVELOPER TO ADD THE RECORDS  
```{r}

sql_list <- 1:nrow(biota_chemistry_values2) %>% 
  map_chr(make_sql_chemistry_values, data = biota_chemistry_values2)
cat("Number of sql sentences: \n") # 2340
length(sql_list) # 2340

i <- 1:length(sql_list)
sql <- paste(sql_list[i], collapse = ";\n") # join sqls together, with "new line" character between
sql <- paste0(sql, ";\n")                   # add final "new line" character at the end
writeLines(sql, "clipboard-1024")   # copies SQLs to clipboard - go to SQL Developer and paste
                                    # "clipboard-1024" instead of "clipboard": increases avaliable
                                    #    for the clipboard

cat("\n") # 2340
cat("First sql sentences: \n") # 2340
head(sql_list, 3)

cat("\n") # 2340
cat("Last sql sentence: \n") # 2340
tail(sql_list, 1)


```



## 9. Save all
```{r}
if (FALSE){
  saveRDS(data_insert_2, 
          paste0("Data/71_data_insert_", selected_year, ".rds"))
  saveRDS(biota_chemistry_values2, 
          paste0("Data/71_biota_chemistry_values_eider_", selected_year, ".rds"))
}

# For reading from disk:
if (FALSE){
  data_insert_2 <- readRDS("Data/71_data_insert_2019.rds")
  biota_chemistry_values2 <- readRDS("Data/71_biota_chemistry_values_eider_2019.rds")
}


```

## 10. Check  

### Reread data 
```{r}

# Get data frame of chemical results (30 seconds or so)
df_chem <- get_biota_chemistry(
  years = selected_year, 
  specimendata = df2_specimens_allyrs, 
  stationdata = df_stations,
  report_samples = TRUE)

cat("\nDownloaded", nrow(df_chem), "records\n")

xtabs(~NAME + STATION_CODE, df_chem)

```

### Test plot  
```{r}

df_chem %>%
  filter(grepl("BDE", NAME)) %>%
  ggplot(aes(NAME, VALUE, color = TISSUE_NAME)) + 
  geom_jitter(width = 0.1) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))

df_chem %>%
  filter(grepl("PCB", NAME)) %>%
  ggplot(aes(NAME, VALUE, color = TISSUE_NAME)) + 
  geom_jitter(width = 0.1) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))

# Metals + dechlorans  
df_chem %>%
  filter(nchar(NAME) == 2) %>%
  ggplot(aes(NAME, VALUE, color = TISSUE_NAME)) + 
  geom_jitter(width = 0.1) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))

```

