---
title: "812 - Import eider duck data from NILU to Nivabasen, 2020"
author: "DHJ"
date: "2 9 2019"
output: 
  html_document:
    keep_md: true
---

Make tables for inserting into NIVAbase   
  
For eider duck data from NILU  
- Note: source of files the same as for cod siloxans (script 210)  

NOTE: samples mosty exist in Nivabasen, but two samples MUST BE ADDED   


## 0. Libraries and functions
```{r}

library(dplyr)
library(purrr)
library(lubridate)
library(stringr)
library(tidyr)
library(ggplot2)
library(niRvana)
library(safejoin)   # https://github.com/moodymudskipper/safejoin

source("812_Import_til_NIVAbasen_NILU_functions.R")

# "Restart" for lazy people that don't want to write the 
#    database password again (deletes all objects except database password/username)
if (FALSE){
  obj <- ls()
  obj <- obj[!obj %in% c("db_privkey", "db_pubkey", "db_pwd", "db_username")]
  rm(list = obj)
  rm(obj)
}


```

### For connectng to Nivabasen    
Store your username and password to R (only once per R session)  
```{r}

set_credentials()

# Check these codes (press F2) for reminder about what comes from which tables:
# niRvana::get_biota_chemistry
# niRvana::get_samples_from_sampleid

```

## 1. Read data


### a. Year
```{r}

selected_year <- 2020

```


### b. Get most recent data produced in Jupyterhub  
We only keep the data for the curret year
```{r}
# dir("Data")
# Data at sample level (chemical data, biological efect parameters and VDSI)
data_all <- readRDS(file = "Files_to_Jupyterhub_2020/01_df_2020_notstandard_2021-06-15.rds")

dat_eider_nivabase <- data_all %>%
  filter(LATIN_NAME %in% "Somateria mollissima")

```


### c. Get samples in Labware file     
```{r}

# March-December data
sql <- paste(
  "select * from NIVADATABASE.LABWARE_CHECK_SAMPLE", 
  "where extract(YEAR from SAMPLED_DATE) =", selected_year, 
  "and extract(MONTH from SAMPLED_DATE) >= 3", 
  "and UPPER(PROSJEKT) like '%MILKYS%';")
sql
df_labware_01 <- get_nivabase_data(sql)

# January-February next year data (if any)
sql <- paste(
  "select * from NIVADATABASE.LABWARE_CHECK_SAMPLE", 
  "where extract(YEAR from SAMPLED_DATE) =", selected_year + 1, 
  "and extract(MONTH from SAMPLED_DATE) <= 2", 
  "and UPPER(PROSJEKT) like '%MILKYS%';")
df_labware_02 <- get_nivabase_data(sql)

# Remove everything except Eider duck stations
# Extract 'Specimen_no' and 'TISSUE_NAME'
df_samples <- bind_rows(df_labware_01, df_labware_02) %>%
  # Add 'Specimen_no'
  mutate(DESCRIPTION = gsub("<f8>", "ø", DESCRIPTION, fixed = TRUE),
         DESCRIPTION = gsub("<c5>", "Å", DESCRIPTION, fixed = TRUE),
         DESCRIPTION = gsub("<e5>", "å", DESCRIPTION, fixed = TRUE),
         DESCRIPTION = gsub("<c6>", "Æ", DESCRIPTION, fixed = TRUE),
         DESCRIPTION = gsub("<e6>", "æ", DESCRIPTION, fixed = TRUE),
         DESCRIPTION2 = DESCRIPTION %>% stringr::str_sub(start = nchar(AQUAMONITOR_CODE) + 1),
         Specimen_no = stringr::str_extract(DESCRIPTION2, "[0-9]+") %>% as.numeric(),
         TISSUE_NAME = stringr::str_sub(TISSUE, start = 4))

# Remove everything except Eider duck stations
df_samples_eider <- df_samples %>%
  filter(AQUAMONITOR_CODE %in% dat_eider_nivabase$STATION_CODE)

if (FALSE){
  df_samples_eider %>%
    select(TEXT_ID, DESCRIPTION, SPECIES, TISSUE, Specimen_no, BIOTA_SAMPLENO, X_BULK_BIO) %>%
    arrange(TISSUE, BIOTA_SAMPLENO)
}


```

### d. Plot samples    
* Note blood samples numbered 90-95 (probably means that they have no corresponding egg data)   
```{r, fig.width=7, fig.height=3}

ggplot(df_samples_eider, aes(BIOTA_SAMPLENO, Specimen_no, color = is.na(X_BULK_BIO))) +
  geom_point() +
  facet_grid(vars(AQUAMONITOR_CODE), vars(TISSUE))

```

### e. Get some generic lookup tables  
* Tissues  
* Species names  
* Projects
* Methods  
```{r}

df_tissue <- get_nivabase_data(
  "select TISSUE_ID,TISSUE_NAME from NIVADATABASE.BIOTA_TISSUE_TYPES")

df_taxon <- get_nivabase_selection(
  "NIVA_TAXON_ID, LATIN_NAME", 
  "TAXONOMY", 
  "LATIN_NAME",
  unique(dat_eider_nivabase$LATIN_NAME), values_are_text = TRUE
  )

# Get a list of projects
df_projects <- get_projects()   # we call it 'df_projects' (the default name)

# Get a list of stations
df_stations <- get_stations_from_project("CEMP_Biota", ignore.case = FALSE)

# Get df_methods (NILU methods only)
df_methods <- get_nivabase_selection(
  "*", 
  "METHOD_DEFINITIONS", 
  "LABORATORY", 
  "NILU", values_are_text = TRUE)

```

### f1. Read existing Nivabase chemical data  
- % C, % N    
- Delta 13C, Delta 15N  
- PFAS  
```{r}

df_chem_specimens <- get_specimens_from_stationdata(
  df_stations %>% filter(STATION_CODE %in% dat_eider_nivabase$STATION_CODE))

# Add species 
df_chem_species <- get_nivabase_data(paste(
  "select a.TAXONOMY_CODE_ID, b.LATIN_NAME",
  "from NIVADATABASE.TAXONOMY_CODES a left join NIVADATABASE.TAXONOMY b",
  "on a.NIVA_TAXON_ID = b.NIVA_TAXON_ID",
  "where a.TAXONOMY_CODE_ID =",
  unique(df_chem_specimens$TAXONOMY_CODE_ID)
  ))
df_chem_specimens <- left_join(df_chem_specimens, df_chem_species, by = "TAXONOMY_CODE_ID")

# Get data frame of chemical results (30 seconds or so)
df_chem <- get_biota_chemistry(
  years = selected_year, 
  specimendata = df_chem_specimens, 
  stationdata = df_stations,
  report_samples = TRUE)

cat("\nDownloaded", nrow(df_chem), "records\n")

```

### f2 Table of SPECIMEN_NO (shows number of specimens)
```{r}

df_chem_specimens %>%
  filter(year(DATE_CAUGHT) == selected_year) %>%
  pull(SPECIMEN_NO) %>%
  table()

```

### f3. Table of SAMPLE_NO (shows number of measurements) 
```{r}

xtabs(~SAMPLE_NO + TISSUE_NAME, dat_eider_nivabase)

```


### f4. Table of parameters   
```{r}

tab <- xtabs(~NAME + STATION_CODE, df_chem)
tab

```



### g. DROP?? Define parameters that we can drop  
* Based on table above (but note that names are not identical)    
```{r}

params_already_in_database <- c(
  "Delta13C", "Delta15N", "% C", "% N", "C/N", "PFAS", "PFBS", 
  "PFDcA", "PFHpA", "PFHxA", "PFHxS", "PFNA", "PFOA", "PFOS", "PFOSA", 
  "PFUdA")

# Sum parameters etc. that we don't need to enter
params_to_drop <- c("BDE6S", "BDESS", "HBCDD", "CB_S7", "Sample")

```



## 2a. Data to insert in NIVAbase
1. Eider duck data are from specimens/samples that MOSTLY ARE in the NIVAbase (for 2018 data, they were NOT) 
    - Sample ID given as Labware ID, e.g. 'NR-2020-09188' 
    - New records in BIOTA_SAMPLES only (SAMPLE_NO taken from data, get SAMPLE_ID from database)
    - Also note that although pairs of eggs and blood were unrelated (I think!),
      the same SPECIMEN_ID was giben for pairs of blood/egg samples     

2. NILU cod data (siloxans; D4,D5 and D6) are from specimens that MOSTLY ARE in the NIVAbase, 
but new samples (liver) are numbered corresponding to the muscle samples
   - Sample ID should normally be given as station number + sample number (1-100)
   - This was added manually in the excel file (Samperapport_edited.xlsx)
     based on the 'fake Labware ID' (e.g. 'NR-2020-09188') added (which doesn't correspond to
     the actual Labware IDs)
   - Use existing record in BIOTA_SINGLE_SPECIMENS (existing SPECIMEN_ID)
   - New records in BIOTA_SAMPLES_SPECIMENS - relationship between SAMPLE_NO and SPECIMEN_ID
   should be the same as for muscle
   - New records in BIOTA_SAMPLES (new SAMPLE_ID, SAMPLE_NO taken from data)

3. Biological effects in cod - from specimens that ARE in the NIVAbase, with the 
exception of a single cod individual (Fish_no = 15, station = 53B). New samples (bile, blood,
liver fraction). Samples are numbered corresponding to the muscle samples (as NILU cod)
    - Use existing records in BIOTA_SINGLE_SPECIMENS (existing SPECIMEN_ID)  
    - New records in BIOTA_SAMPLES_SPECIMENS (new SAMPLE_ID, but existing SPECIMEN_ID)  
    - New records in BIOTA_SAMPLES (new SAMPLE_ID, SAMPLE_NO taken from data)  

## 2b. Eider duck

### a. Read NILU data   
From script 210 (2020 version)  
```{r}

dat_eider <- readRDS("Data/210_dat_eiderduck_2020.rds") %>%
  mutate(TEXT_ID = sub("Nr. ", "NR-", Sample_no, fixed = TRUE))

# Show all
# xtabs(~TEXT_ID + Group, dat_eider)

```

### b. Eider duck: Check TEXT_ID    
- Except siloxan, all data have sample ID (as can be seen in the excel file)   
```{r}

# Get unique IDs
id_eider <- data.frame(TEXT_ID = names(table(dat_eider$TEXT_ID))) %>%
  left_join(df_samples_eider %>% select(TEXT_ID, SPECIES, AQUAMONITOR_ID, AQUAMONITOR_CODE,
                                        BIOTA_SAMPLENO, X_BULK_BIO, TISSUE, DESCRIPTION2, Specimen_no),
            by = "TEXT_ID") %>%
  mutate(TISSUE_NAME = stringr::str_sub(TISSUE, start = 4))
id_cod <- data.frame(TEXT_ID = names(table(dat_cod$TEXT_ID))) %>%
  left_join(df_samples %>% select(TEXT_ID, SPECIES, AQUAMONITOR_ID, AQUAMONITOR_CODE,
                                  BIOTA_SAMPLENO, X_BULK_BIO, TISSUE, DESCRIPTION2, Specimen_no),
            by = "TEXT_ID") %>%
  mutate(TISSUE_NAME = stringr::str_sub(TISSUE, start = 4))

# Check TEXT_ID is found in Labware samples 

sel <- !is.na(id_eider$AQUAMONITOR_ID)

if (mean(sel) < 1){
  warning("Eider duck: NOT all TEXT_ID is found in Labware samples \n")
  cat("Lacking: \n")
  id_eider$TEXT_ID[!sel]
} else {
  message("Eider duck: All TEXT_ID found in Labware samples \n")
}

dat_eider_lacking_id <- dat_eider %>%
  filter(is.na(TEXT_ID))

if (nrow(check) > 0){
  warning("Eider duck: NOT all samples have TEXT_ID \n")
  cat("Lacking: \n")
  table(dat_eider_lacking_id$Group)
} else {
  message("Eider duck: All samples have TEXT_ID \n")
}

#
# Code for cod, NOT USED as we use SPECIMEN_NO instead
#
#
# sel <- !is.na(id_cod$AQUAMONITOR_ID)
# 
# if (mean(sel) < 1){
#   warning("Cod: NOT all TEXT_ID is found in Labware samples \n")
#   id_cod$TEXT_ID[!sel]
# } else {
#   message("Cod: All TEXT_ID found in Labware samples \n")
# }
# 
# # I don't understand why not all TEXT_IDs are found for cod,
# #   or why they have TEXT_ID at all. Last year they had just station + no 1-15,
# #   which fits with the "opparbeiding" excel files from Bjørnar
# write.csv2(id_cod, "Data/210_Cod_siloxan_IDs_raw.csv")


```

### c. Eider duck: Make lookup table for samples, including SAMPLE_ID    
- Lacking SAMPLE_ID for 10 blood samples  
- These data have not entered NIVAbase  
- The reason is probably that X_BULK_BIO contains a letter instead of just the number  
```{r}

# Get the link (tissue, SAMPLE_NO) -> SAMPLE_ID by "extracting" from Nivabasen
df_lookup_sample_id_1 <- dat_eider_nivabase %>%
  # Get the sample IDs by "extracting" from Nivabasen
  distinct(SAMPLE_NO, TISSUE_NAME, SAMPLE_ID) %>%
  arrange(TISSUE_NAME, SAMPLE_NO)

# Start with labware samples 
df_lookup_sample_id <- df_samples_eider %>%
  select(AQUAMONITOR_CODE, TISSUE_NAME, BIOTA_SAMPLENO, X_BULK_BIO, TEXT_ID, Specimen_no) %>%
  # ...and add SAMPLE_ID
  left_join(df_lookup_sample_id_1, 
            by = c("TISSUE_NAME", "BIOTA_SAMPLENO" = "SAMPLE_NO"))

# Lacking SAMPLE_ID for 10 samples  
table(is.na(df_lookup_sample_id$SAMPLE_ID))

#
# For sending to Espen, for fixing the error
#
if (FALSE){
  
  df_samples_eider %>% 
    select(TISSUE_NAME, BIOTA_SAMPLENO, TEXT_ID, X_BULK_BIO, DESCRIPTION) %>%
    writexl::write_xlsx("Data/Prøver_ærfugl_2020_raw.xlsx")
  
  df_samples_eider %>% 
    writexl::write_xlsx("Data/Prøver_ærfugl_2020_allekolonner.xlsx")
}

```

### d. Eider duck: Add SAMPLE_ID     
- Using TEXT_ID for all except siloxans (part a)  
- Using TISSUE_NAME + SPECIMEN_NO for siloxans (part b)  
- In addition to lacking : Two samples (blod 10 + 11) have siloxan data but lack in Labware    
   - These two samples MUST BE ADDED    
```{r}

# Join this table 
dat_eider2_a <- dat_eider %>%
  filter(!is.na(TEXT_ID)) %>%
  left_join(df_lookup_sample_id %>% select(TEXT_ID, SAMPLE_ID), 
            by = "TEXT_ID")

dat_eider2_b <- dat_eider %>%
  filter(is.na(TEXT_ID)) %>%
  mutate(SPECIMEN_NO = as.numeric(SPECIMEN_NO)) %>%
  left_join(df_lookup_sample_id %>% select(TISSUE_NAME, Specimen_no, SAMPLE_ID), 
            by = c("TISSUE_NAME", "SPECIMEN_NO" = "Specimen_no"))

# A lot of NAs in both cases, see previous chunk   
xtabs(~addNA(SAMPLE_ID), dat_eider2_a)
xtabs(~addNA(SAMPLE_ID), dat_eider2_b)


```

### x. NOTE: ALL METHOD_ID were set in script 210  
 
```{r}

table(addNA(dat_eider$METHOD_ID))

```


## 4. Eider data to insert, start  
Date is from   
`K:\Prosjekter\Sjøvann\JAMP\2017\opparbeiding biota\sjøfugl\Felttabell egg 2017 Ærfugl Kongsfjord MILKYS.xlsx`
```{r}


data_insert_1 <- data_for_input %>%
  filter(!PARAM %in% c(params_already_in_database, params_to_drop)) %>% 
  filter(!PARAM %in% params_to_drop) %>% 
  select(STATION_CODE, STATION_ID, LATIN_NAME, TISSUE_NAME, SAMPLE_NO2,
         PARAM, UNIT,
         VALUE_WW, FLAG1) %>%
  rename(SAMPLE_NO = SAMPLE_NO2,
         UNIT_orig = UNIT) %>%
  safe_left_join(
    df_chem %>% distinct(STATION_CODE, TISSUE_NAME, SAMPLE_DATE),
    by = c("STATION_CODE", "TISSUE_NAME"), na_matches = "never",
    check = "BMV"
  )

```




## 8. BIOTA_CHEMISTRY_VALUES

 
### a. Make lookup table for METHOD_ID   
df_lookup_methodid   

```{r}

#
# The following will be used - seems to be the common ones used in 2016 data:  
#
# Co, Sn                  mg/kg         -     	      BIOTA
# Hg                      ng/g w.w.     ICP-MS 	      BIOTA
# Cu + the other metals   Âµg/g w.w.	    ICP-MS	      BIOTA
# PCB-118	                ng/g w.w.	    GC-MS         BIOTA
# BDE-47	                ng/g w.w.	    GC-MS	        BIOTA
# D4	                    ng/g w.w.	    CSR-LVI-GCMS	BIOTA
# a-HBCD	                ng/g (w.w.)   -             -
# Hexachlorobenzene (HCB)	ng/g w.w.	    GC-MS         -
# SCCP	                  ng/g w.w.	    GC-MS	        BIOTA
# Fett                    %             -             Biota 


# Try simple left join of the data
check1 <- data_insert_2 %>%
  count(NAME, UNIT) %>%
  left_join(df_methods %>%
              filter(toupper(MATRIX) %in% "BIOTA") %>%
              select(NAME, UNIT, METHOD_ID, METHOD_REF, MATRIX), 
            by = c("NAME", "UNIT")) %>%
  group_by(NAME, UNIT) %>%
  mutate(n = n())

# We may get a lot of duplicates
cat("Number of duplicates in df_methods, first check: \n")
sum(check1$n > 1)    # 130

# Visual check:
# check1

# Pick one method per NAME,UNIT 
df_lookup_methodid <- check1 %>%
  filter(n == 1 | !is.na(MATRIX))  # We pick those methods which either have no alternatives,
                                   #   or the ones with Biota/BIOTA in MATRIX
# Visual check:
# df_lookup_methodid

# Check for duplicates
check2 <- df_lookup_methodid %>%
  count(NAME, UNIT, wt = 1)
# We get no duplicates
cat("Number of duplicates in df_methods, second check: \n")
sum(check2$n > 1)      # should be zero

# Visual check:
# check1

# Check the othRr way (that all NAME/UNIT are repesented)
check3 <- data_insert_2 %>%
  left_join(df_lookup_methodid %>% select(NAME, UNIT, METHOD_ID, METHOD_REF, MATRIX),
            by = c("NAME", "UNIT"))

cat("Number of parameter/unit combinations not found: \n")
sum(is.na(check3$METHOD_ID))  # should be zero

if (sum(is.na(check3$METHOD_ID)) > 0)  # should be zero
  check3 %>% filter(is.na(check3$METHOD_ID)) %>% count(PARAM, NAME)

```

### b. Start `biota_chemistry_values_eider`   
By selecting columns and adding `METHOD_ID`   
```{r}
# df <- get_nivabase_data("select * from NIVADATABASE.BIOTA_CHEMISTRY_VALUES where rownum < 4")
# df
# df %>% colnames() %>% dput()

# "VALUE_ID"              - Let the database decide
# "SAMPLE_ID"             - From the database, after BIOTA_SAMPLES have been inserted
# "METHOD_ID"             - Lookup based on NAME and UNIT
# "VALUE"                 - From data
# "FLAG1"                 - From data
# "FLAG2"                 - NA
# "ENTERED_BY"            - DHJ
# "ENTERED_DATE"          - date, see above
# "REMARK"                - NA
# "DETECTION_LIMIT"       - NA
# "UNCERTAINTY"           - NA
# "QUANTIFICATION_LIMIT"  - NA
# "APPROVED"              - NA?

biota_chemistry_values1 <- data_insert_2 %>%
  safe_left_join(df_lookup_methodid, by = c("NAME", "UNIT"), 
                 check = "VM", na_matches = "never")

biota_chemistry_values1 %>%
  select(STATION_CODE, TISSUE_NAME, SAMPLE_NO, METHOD_ID, VALUE, FLAG1)

```

### c. Add SAMPLE_ID  
```{r}

biota_chemistry_values2 <- biota_chemistry_values1 %>%
  safe_left_join(df_lookup_sample_id, by = c("TISSUE_NAME", "SAMPLE_NO"), 
                 check = "BCVMn", na_matches = "never")

biota_chemistry_values2$SAMPLE_ID %>% unique() %>% sort()

cat("\n")
cat("Number of SAMPLE_ID lacking (must be zero): \n")
sum(is.na(biota_chemistry_values2$SAMPLE_ID))

```


### d. Make SQLs  
And copies them to clipboard  
PASTE INTO SQL DEVELOPER TO ADD THE RECORDS  
```{r}

sql_list <- 1:nrow(biota_chemistry_values2) %>% 
  map_chr(make_sql_chemistry_values, data = biota_chemistry_values2)
cat("Number of sql sentences: \n") # 2340
length(sql_list) # 2340

i <- 1:length(sql_list)
sql <- paste(sql_list[i], collapse = ";\n") # join sqls together, with "new line" character between
sql <- paste0(sql, ";\n")                   # add final "new line" character at the end
writeLines(sql, "clipboard-1024")   # copies SQLs to clipboard - go to SQL Developer and paste
                                    # "clipboard-1024" instead of "clipboard": increases avaliable
                                    #    for the clipboard

cat("\n") # 2340
cat("First sql sentences: \n") # 2340
head(sql_list, 3)

cat("\n") # 2340
cat("Last sql sentence: \n") # 2340
tail(sql_list, 1)


```



## 9. Save all
```{r}
if (FALSE){
  saveRDS(data_insert_2, 
          paste0("Data/71_data_insert_", selected_year, ".rds"))
  saveRDS(biota_chemistry_values2, 
          paste0("Data/71_biota_chemistry_values_eider_", selected_year, ".rds"))
}

# For reading from disk:
if (FALSE){
  data_insert_2 <- readRDS("Data/71_data_insert_2019.rds")
  biota_chemistry_values2 <- readRDS("Data/71_biota_chemistry_values_eider_2019.rds")
}


```

## 10. Check  

### Reread data 
```{r}

# Get data frame of chemical results (30 seconds or so)
df_chem <- get_biota_chemistry(
  years = selected_year, 
  specimendata = df_chem_specimens, 
  stationdata = df_stations,
  report_samples = TRUE)

cat("\nDownloaded", nrow(df_chem), "records\n")

xtabs(~NAME + STATION_CODE, df_chem)

```

### Test plot  
```{r}

df_chem %>%
  filter(grepl("BDE", NAME)) %>%
  ggplot(aes(NAME, VALUE, color = TISSUE_NAME)) + 
  geom_jitter(width = 0.1) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))

df_chem %>%
  filter(grepl("PCB", NAME)) %>%
  ggplot(aes(NAME, VALUE, color = TISSUE_NAME)) + 
  geom_jitter(width = 0.1) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))

# Metals + dechlorans  
df_chem %>%
  filter(nchar(NAME) == 2) %>%
  ggplot(aes(NAME, VALUE, color = TISSUE_NAME)) + 
  geom_jitter(width = 0.1) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))

```

## APPENDIX  
```{r}

```{r}

df1 <- get_nivabase_selection(
  "*",
  "BIOTA_SINGLE_SPECIMENS",
  "STATION_ID",
  69750,
  extra_where = " and extract(YEAR from DATE_CAUGHT) = 2020")

df2 <- get_nivabase_selection(
  "*",
  "BIOTA_SAMPLES_SPECIMENS",
  "SPECIMEN_ID",
  df1$SPECIMEN_ID)

df3 <- get_nivabase_selection(
  "*",
  "BIOTA_SAMPLES",
  "SAMPLE_ID",
  df2$SAMPLE_ID)

df3 <- get_nivabase_selection(
  "*",
  "BIOTA_SAMPLES",
  "SAMPLE_ID",
  df2$SAMPLE_ID)

df3b <- get_nivabase_selection(
  "*",
  "BIOTA_SAMPLES",
  "SAMPLE_ID",
  245960)

df3b <- get_nivabase_selection(
  "*",
  "BIOTA_SAMPLES",
  "STATION_ID",
  69750,
  extra_where = " and extract(YEAR from SAMPLE_DATE) = 2020")

```
```

