---
title: "74_Import_til_NIVAbasen_imposex_2019data"
author: "DHJ"
date: "2 9 2019"
output: 
  html_document:
    keep_md: true
---

Make tables for inserting into NIVAbase   
  
For imposex (and intersex) in snails (Nucella + Littorina)  

SKIPPING INTERSEX (71G) until later (cannot find code in METHODS)


## 0. Libraries and functions
```{r}

library(dplyr)
library(purrr)
library(lubridate)
library(stringr)
# library(ggplot2)
library(niRvana)

library(safejoin) # https://github.com/moodymudskipper/safejoin

source("71_Import_til_NIVAbasen_functions.R")

# "Restart" for lazy people that don't want to write the 
#    database password again (deletes all objects except database password/username)
if (FALSE){
  obj <- ls()
  obj <- obj[!obj %in% c("db_privkey", "db_pubkey", "db_pwd", "db_username")]
  rm(list = obj)
  rm(obj)
}

```


## 1. Read data

### a. Year
```{r}

selected_year <- 2019

```


### b. Data to insert in NIVAbase
1. Eider duck data are from specimens/samples NOT in the NIVAbase  
2. NILU cod data are from specimens that ARE in the NIVAbase, 
but new samples (liver) are numbered corresponding to the muscle samples
3. Biological effects in cod - existing specimens, new samples
4. VDSI (imposex) in snails - existing 'individuals' (pooled samples) and samples

```{r}

# Create medians from raw data  
dat <- readRDS(file = "Input_data_2019/101_data_updated_2020-08-05.rds") %>%
  filter(MYEAR == selected_year & PARAM %in% "VDSI" & !is.na(VALUE_WW)) %>%
  group_by(MYEAR, STATION_CODE, LATIN_NAME, TISSUE_NAME, PARAM) %>%
  summarise(VALUE = median(VALUE_WW)) %>%
  mutate(STATION_CODE = case_when(
    STATION_CODE == "227G" ~ "227G2",
    TRUE ~ STATION_CODE)
    )

dat %>%
  select(MYEAR, STATION_CODE, LATIN_NAME, PARAM, VALUE)
      
```


### c. Get samples in Labware file     
```{r}

#
# 2019 data  
#

# March-December data
sql <- paste(
  "select * from NIVADATABASE.LABWARE_CHECK_SAMPLE", 
  "where extract(YEAR from SAMPLED_DATE) =", selected_year, 
  "and extract(MONTH from SAMPLED_DATE) >= 3", 
  "and UPPER(PROSJEKT) like '%MILKYS%';")
sql
df_labware_01 <- get_nivabase_data(sql)

# January-February next year data
sql <- paste(
  "select * from NIVADATABASE.LABWARE_CHECK_SAMPLE", 
  "where extract(YEAR from SAMPLED_DATE) =", selected_year + 1, 
  "and extract(MONTH from SAMPLED_DATE) <= 2", 
  "and UPPER(PROSJEKT) like '%MILKYS%';")
df_labware_02 <- get_nivabase_data(sql)

df_samples <- bind_rows(df_labware_01, df_labware_02) %>%
  filter(AQUAMONITOR_CODE %in% dat$STATION_CODE)

```

### d. Add dates to data    
```{r}

dat <- dat %>%
  safe_left_join(df_samples, 
                 na_matches = "never",
                 by = c("STATION_CODE" = "AQUAMONITOR_CODE"),
                 check = "BCMV") %>%
  rename(SAMPLE_DATE = SAMPLED_DATE)

```


## 2.Connection to NIVAbasen  

### Store your username and password to R (only once per R session):
```{r}

set_credentials()

# Check these codes (press F2) for reminder about what comes from which tables:
# niRvana::get_biota_chemistry
# niRvana::get_samples_from_sampleid

```
### Get some generic lookup tables  
```{r}

df_tissue <- get_nivabase_data(
  "select TISSUE_ID,TISSUE_NAME from NIVADATABASE.BIOTA_TISSUE_TYPES")

df_species <- get_nivabase_data(
  "select * from NIVADATABASE.SPECIES")

df_taxoncodes <- 
  get_nivabase_data(
    "select NIVA_TAXON_ID, TAXONOMY_CODE_ID, CODE from NIVADATABASE.TAXONOMY_CODES")

df_taxon <- get_nivabase_selection(
  "NIVA_TAXON_ID, LATIN_NAME", 
  "TAXONOMY", 
  "LATIN_NAME",
  unique(dat$LATIN_NAME), values_are_text = TRUE
  )

# Get a list of projects
df_projects <- get_projects()   # we call it 'df_projects' (the default name)

# Get a list of stations
df_stations <- get_stations_from_project("CEMP_Biota", ignore.case = FALSE)

```


## 3. Check parameter   
Mostly used: VDSI (METHOD_ID = 15631)  
- exception: 2015 used VDSI with METHOD_ID = 28589  
- has been submitted on an individual basis up to 2009, then as average values in 2015 and 2018    
```{r}

pars <- c("Imposex", "VDSI", "Intersex")

# Check
df <- get_nivabase_selection("*", 
                       "METHOD_DEFINITIONS",
                       "NAME", 
                       pars, 
                       values_are_text = TRUE)
df

# Tabulate species, years and methods
par <- "VDSI"
df_param <- df[grepl(par, df$NAME),]
df1 <- get_nivabase_selection("*", "BIOTA_CHEMISTRY_VALUES", "METHOD_ID", df_param$METHOD_ID)
nrow(df1)
df2 <- get_nivabase_selection("*", "BIOTA_SAMPLES", "SAMPLE_ID", df1$SAMPLE_ID) %>%
  left_join(df1 %>% select(SAMPLE_ID, METHOD_ID))  %>%
  left_join(df %>% select(METHOD_ID, NAME, UNIT, METHOD_REF, MATRIX))  %>%
  left_join(df_species) %>%
  left_join(df_tissue) %>%
  mutate(Year = year(SAMPLE_DATE))

xtabs(~TISSUE_NAME, df2)   # "Hel organisme"
# xtabs(~LATIN_NAME, df2)
df2 %>% 
  count(LATIN_NAME, Year, NAME, UNIT, METHOD_ID, MATRIX)  # 15361
  
# Pick methods
df_methods <- df %>%
  filter(METHOD_ID %in% c(15631))

df_methods

# Pick parameters for later
pars <- df_methods$NAME

```

## 4. Check existing data, former years  
THIS WHOLE THING CAN BE SKIPPED (I think)
Starting from VDSI, via samples  
```{r}
# df2 <- get_nivabase_selection("*", "BIOTA_SAMPLES", "SAMPLE_ID", df1$SAMPLE_ID)

df3 <- get_nivabase_selection("SAMPLE_ID, SPECIMEN_ID",
                              "BIOTA_SAMPLES_SPECIMENS",
                              "SAMPLE_ID",
                              unique(df2$SAMPLE_ID))

df4 <- get_nivabase_selection("STATION_ID, SPECIMEN_ID, SPECIMEN_NO, DATE_CAUGHT, TAXONOMY_CODE_ID",
                              "BIOTA_SINGLE_SPECIMENS",
                              "SPECIMEN_ID",
                              unique(df3$SPECIMEN_ID))

nrow(df1)
nrow(df2)
nrow(df3)
nrow(df4)

# Checl 11G 
df4 %>%
  filter(STATION_ID == 47219) %>%  # 47219 = 11G
  arrange(SPECIMEN_NO)
# 2010  - a bunch
# 2015 - only 1
```


### Joining tables, starting from sample  
```{r}

df5 <- df1 %>%
  select(SAMPLE_ID, METHOD_ID, VALUE, FLAG1, DETECTION_LIMIT, QUANTIFICATION_LIMIT) %>%
  safe_left_join(df2, na_matches = "never", check = "v")

df3 %>%
  count(SAMPLE_ID, SPECIMEN_ID) %>%
  filter(n > 1)  # zero

# Most (491) samples have one specimen, but 8 samples have 20-30 specimens
df3 %>%
  count(SAMPLE_ID) %>%
  count(n)

# Pick one of those 8 samples
spec_check <- df3 %>%
  filter(SAMPLE_ID == 57374) %>% # View("df3")
  pull(SPECIMEN_ID)

# Checking specimens + stations - no station codes found
df4 %>%
  filter(SPECIMEN_ID %in% spec_check) %>%
  safe_left_join(
    get_nivabase_selection(
      "STATION_ID, STATION_CODE, STATION_NAME", "PROJECTS_STATIONS",
      "STATION_ID", .$SPECIMEN_ID), 
    na_matches = "never"
    ) # %>%  View("df4")


```





### Stations found  
```{r}
tab <- table(df4$STATION_ID)
length(tab)
tab[tab > 1]

# QUery PROJECTS_STATIONS for stations
vdsi_stations_all <- get_nivabase_selection(
  "STATION_ID, STATION_CODE, STATION_NAME", "PROJECTS_STATIONS",
  "STATION_ID", names(tab))
# Includes a bunch of Scottish stations

```


### Add year, stations + species    
```{r}
vdsi_stations <- vdsi_stations_all %>%
  filter(STATION_CODE %in% unique(dat$STATION_CODE)) %>%
  group_by(STATION_ID, STATION_CODE) %>%
  summarise(STATION_NAME = first(STATION_NAME))   # several 

df4 <- df4 %>%
  mutate(Year = year(DATE_CAUGHT)) %>%
  safe_left_join(vdsi_stations , check = "V")

# Add species
df_species_tax <- df4$TAXONOMY_CODE_ID %>%
  unique() %>%
  map_df(taxid_to_latin)

df4 <- df4 %>%
  safe_left_join(df_species_tax, check = "V")

# Check 11G 
df4 %>%
  filter(STATION_ID == 47219) %>%  # 47219 = 11G
  arrange(SPECIMEN_NO)

```

### Present MILKYS stations 
```{r}

vdsi_stations %>%
  arrange(STATION_ID)

```

### Tabulate year x stations
```{r}
df4 %>%
  filter(STATION_CODE %in% unique(dat$STATION_CODE)) %>%  # only MILKYS
  xtabs(~Year + STATION_CODE, .)

```

### Tabulate species x stations
```{r}
df4 %>%
  filter(STATION_CODE %in% unique(dat$STATION_CODE)) %>%
  xtabs(~LATIN_NAME + STATION_CODE, .)

```

### Example 2010      
47219 = 11G in 2010  
Many specimens per sample
```{r}
df4_sel <- df4 %>%
  filter(Year %in% 2010 & STATION_CODE %in% "11G") %>% 
  arrange(SPECIMEN_NO)
df4_sel

df3_sel <- df3 %>%
  filter(SPECIMEN_ID %in% df4_sel$SPECIMEN_ID)
df3_sel %>%
  xtabs(~SAMPLE_ID, .)

df2_sel <- df2 %>%
  filter(SAMPLE_ID %in% unique(df3_sel$SAMPLE_ID))
df2_sel  

df1_sel <- get_nivabase_selection("SAMPLE_ID, METHOD_ID, VALUE, FLAG1",
                                  "BIOTA_CHEMISTRY_VALUES",
                                  "SAMPLE_ID", df2_sel$SAMPLE_ID)
df1_sel

df0_sel <- get_nivabase_selection("METHOD_ID, NAME, UNIT",
                                  "METHOD_DEFINITIONS",
                                  "METHOD_ID", df1_sel$METHOD_ID)
df0_sel
```


## 6. Check TBT data for the selected year    

### Get chemical data  
```{r}

df2_specimens_allyrs <- get_specimens_from_stationdata(
  df_stations %>% filter(STATION_CODE %in% dat$STATION_CODE))

# Get data frame of chemical results (30 seconds or so)
df_snailchem <- get_biota_chemistry(
  years = selected_year, 
  specimendata = df2_specimens_allyrs, 
  stationdata = df_stations,
  report_samples = TRUE)

cat("\nDownloaded", nrow(df_snailchem), "records\n")

xtabs(~NAME + STATION_CODE, df_snailchem)

```

### Get SAMPLE_ID used  
```{r}

station_sampleid <- df_snailchem %>%
  count(SAMPLE_ID, STATION_CODE)
station_sampleid

```


## 14. BIOTA_CHEMISTRY_VALUES

### a. Read data, if necessary  
Add METHOD_ID and FLAG1  
```{r}

xtabs(~PARAM, dat)

dat_summ <- dat %>%
  filter(PARAM %in% c("VDSI", "Intersex")) %>%
  mutate(
    METHOD_ID = df_methods$METHOD_ID[1],
    FLAG1 = as.character(NA)
  )

```

### b. Add SAMPLE_ID  
Which is all we need (see code for 'make_sql_chemistry_values' in '71...functions')
```{r}

cn1 <- dat_summ %>% colnames()

# Add SAMPLE_ID to chemistry_values data
if (!("SAMPLE_ID" %in% cn1)){    # checks that SAMPLE_ID has not already been added
  dat_summ <- dat_summ %>%
    safe_left_join(station_sampleid,
                   by = "STATION_CODE",
                   na_matches = "never",
                   check = "MV")
}


```


### c. Make SQLs  
PASTE INTO SQL DEVELOPER TO ADD THE RECORDS   
Note to DHJ: use "SQL developer (latest version)" from desktop. Don't use the start menu.  
- remember `commit;` after running the insert sentences  
```{r}
# make_sql_chemistry_values(1, data = dat_summ)

sql_list <- 1:nrow(dat_summ) %>% 
  map_chr(make_sql_chemistry_values, data = dat_summ)

i <- 1:length(sql_list)
sql <- paste(sql_list[i], collapse = ";\n")
sql <- paste0(sql, ";\n")
writeLines(sql, "clipboard-1024")   # copies SQLs to clipboard - go to SQL Developer and paste
                                    # "clipboard-1024" instead of "clipboard": increases available
                                    #    for the clipboard

cat("Number of SQLs: \n")
length(sql_list)  # 144

cat("\nSample of SQLs: \n")
sql_list[1:3]


```

## 15. Reread data  
```{r}

# Get data frame of chemical results (30 seconds or so)
df_snailchem <- get_biota_chemistry(
  years = selected_year, 
  specimendata = df2_specimens_allyrs, 
  stationdata = df_stations,
  report_samples = TRUE)

cat("\nDownloaded", nrow(df_snailchem), "records\n")

xtabs(~NAME + STATION_CODE, df_snailchem)

```
### Plot data   
Very boring plot
```{r, fig.width=5, fig.height=2}

df_snailchem %>%
  filter(NAME == "VDSI") %>%
  ggplot(aes(STATION_CODE, VALUE)) + 
  geom_point()

```




