---
title: "74_Import_til_NIVAbasen_imposex_2020data"
author: "DHJ"
date: "2 9 2019"
output: 
  html_document:
    keep_md: true
---

Make tables for inserting into NIVAbase   
  
For imposex (and intersex) in snails (Nucella + Littorina)  

For the first time, Intersex was added (including adding it in METHODS)   


## 0. Setup

### a. Libraries and functions
```{r}

library(dplyr)
library(purrr)
library(lubridate)
library(stringr)
# library(ggplot2)
library(niRvana)

# If needed
# remotes::install_github("https://github.com/moodymudskipper/safejoin")
library(safejoin) # https://github.com/moodymudskipper/safejoin

source("810_Import_til_NIVAbasen_functions.R")

# "Restart" for lazy people that don't want to write the 
#    database password again (deletes all objects except database password/username)
if (FALSE){
  obj <- ls()
  obj <- obj[!obj %in% c("db_privkey", "db_pubkey", "db_pwd", "db_username")]
  rm(list = obj)
  rm(obj)
}

```

### b. Creds for connectng to Nivabasen    
Store your username and password to R (only once per R session)  
```{r}

set_credentials()

# Check these codes (press F2) for reminder about what comes from which tables:
# niRvana::get_biota_chemistry
# niRvana::get_samples_from_sampleid

```



## 1. Read data

### a. Year
```{r}

selected_year <- 2020

```


### b. Data to insert in NIVAbase   
1. Eider duck data are from specimens/samples NOT in the NIVAbase  
2. NILU cod data are from specimens that ARE in the NIVAbase, 
but new samples (liver) are numbered corresponding to the muscle samples
3. Biological effects in cod - existing specimens, new samples
4. VDSI (imposex) in snails - existing 'individuals' (pooled samples) and samples
  
```{r, message=FALSE}

folder_vdsi <- "K:/Prosjekter/Sjøvann/JAMP/2020/opparbeiding/Snegl"

#
# 1. Imposex in Littorina
#

fn <- dir(folder_vdsi, pattern = ".xlsx")
fn <- fn[!grepl("AAMAL", fn)]
fn <- fn[!grepl("LITTORINA", fn)]
fn <- fn[!grepl("~", fn, fixed = TRUE)]

# Names = Extract all characters to the left of the first underscore  
names(fn) <- stringr::str_extract(fn, "[^_]+")
names(fn)[names(fn) == "131"] <- "131G"
# fn


read_excel_snail <- function(fn){
  df <- readxl::read_excel(paste0(folder_vdsi, "/", fn), skip = 6)
  df <- df[1:6]
  df <- df[!is.na(df$`Analysert av`),]
  df
  }

dat_vdsi_list <- fn %>% map(read_excel_snail)

# 15G has 51 rows:
# dat_vdsi_list %>% map_int(nrow)

# Check 15G
# dat_vdsi_list[["15G"]] %>% View()
# dat_vdsi_list[["36G"]] %>% View()
dat_vdsi_list[["15G"]] <- dat_vdsi_list[["15G"]][1:50,]  # delete last line

# Pick only number, sex and Imposex 
# ...we will deal with length and penis length later  
dat_vdsi <- dat_vdsi_list %>% 
  map(~.[c(1,3,5)]) %>%
  bind_rows(.id = "STATION_CODE") %>%
  rename(VALUE_WW = VDSI, 
         Sex = Kjønn) %>%
  mutate(PARAM = "VDSI",
         UNIT = "idx",
         LATIN_NAME = "Nucella lapillus")

# dat_vdsi

cat("\n------------------------------------------\n")
cat("Imposex data: \n------------------------------------------\n")
xtabs(~is.na(VALUE_WW) + addNA(Sex), dat_vdsi)


#
# 2. Intersex
#
fn <- "71G_fugløy_intersex_LITTORINA_2020.xlsx"
df <- readxl::read_excel(paste0(folder_vdsi, "/", fn), range = "A3:G53")

dat_intersex <- df %>%
  filter(F == 1) %>%
  mutate(STATION_CODE = "71G", 
         Sex = "f",
         PARAM = "Intersex",
         UNIT = "PERCENT",
         LATIN_NAME = "Littorina littorea") %>%
  mutate(VALUE_WW = `ISI   (intersex-stadie)`*100) %>%
  select(STATION_CODE, Sex, VALUE_WW, PARAM, LATIN_NAME)

#
# 3. Combine and summarize   
#

dat <- bind_rows(dat_vdsi, dat_intersex) %>%
  filter(Sex == "f") %>%
  group_by(STATION_CODE, PARAM, LATIN_NAME) %>%
  summarise(VALUE = mean(VALUE_WW), .groups = "drop")

dat$MYEAR <- selected_year
dat$TISSUE_NAME <- "Whole soft body"

cat("\n------------------------------------------\n")
cat("Combined data: \n------------------------------------------\n")
xtabs(~PARAM, dat )

# For 2019 data:
# dat_2019 <- readRDS(file = "../Milkys/Input_data_2019/101_data_updated_2020-08-05.rds") %>%
#   filter(MYEAR == selected_year & PARAM %in% "VDSI" & !is.na(VALUE_WW)) %>%
#   group_by(MYEAR, STATION_CODE, LATIN_NAME, TISSUE_NAME, PARAM) %>%
#   summarise(VALUE = median(VALUE_WW)) %>%
#   mutate(STATION_CODE = case_when(
#     STATION_CODE == "227G" ~ "227G2",
#     TRUE ~ STATION_CODE)
#     )


```

### c1. Get samples in Labware file     
```{r}

#
# For data the same year    
#

# March-December data
sql <- paste(
  "select * from NIVADATABASE.LABWARE_CHECK_SAMPLE", 
  "where extract(YEAR from SAMPLED_DATE) =", selected_year, 
  "and extract(MONTH from SAMPLED_DATE) >= 3", 
  "and UPPER(PROSJEKT) like '%MILKYS%';")
# sql
df_labware_01 <- get_nivabase_data(sql)

# January-February next year data
sql <- paste(
  "select * from NIVADATABASE.LABWARE_CHECK_SAMPLE", 
  "where extract(YEAR from SAMPLED_DATE) =", selected_year + 1, 
  "and extract(MONTH from SAMPLED_DATE) <= 2", 
  "and UPPER(PROSJEKT) like '%MILKYS%';")
df_labware_02 <- get_nivabase_data(sql)

df_samples <- bind_rows(df_labware_01, df_labware_02) %>%
  filter(AQUAMONITOR_CODE %in% dat$STATION_CODE)

```

### c2. Check that we have all stations  

```{r}

a <- dat$STATION_CODE
b <- df_samples$AQUAMONITOR_CODE
check <- a[!a %in% b]

if (length(check) == 0){
  message("'df_samples' contains all stations we need")
} else {
  stop("'df_samples' does NOT contain all stations we need!")
}

```


### d. Add dates to data    
```{r}

check <- df_samples %>%
  add_count(AQUAMONITOR_CODE) %>%
  filter(n > 1)

if (nrow(check) > 0){
  stop("'df_samples' have > 1 lines per station. Check 'check'")
}

# In this case, two records of 71G, both with same date
# Delete the "extra" one
nrow(df_samples)
df_samples <- df_samples %>%
  filter(!grepl("Snegl - Ekstra", DESCRIPTION))
nrow(df_samples)

# Merge
dat <- dat %>%
  safe_left_join(df_samples, 
                 na_matches = "never",
                 by = c("STATION_CODE" = "AQUAMONITOR_CODE"),
                 check = "BCMV") %>%
  rename(SAMPLE_DATE = SAMPLED_DATE)

```


## 2.Connection to NIVAbasen  

### Get some generic lookup tables  
```{r}

df_tissue <- get_nivabase_data(
  "select TISSUE_ID,TISSUE_NAME from NIVADATABASE.BIOTA_TISSUE_TYPES")

df_species <- get_nivabase_data(
  "select * from NIVADATABASE.SPECIES")

df_taxoncodes <- 
  get_nivabase_data(
    "select NIVA_TAXON_ID, TAXONOMY_CODE_ID, CODE from NIVADATABASE.TAXONOMY_CODES")

df_taxon <- get_nivabase_selection(
  "NIVA_TAXON_ID, LATIN_NAME", 
  "TAXONOMY", 
  "LATIN_NAME",
  unique(dat$LATIN_NAME), values_are_text = TRUE
  )

# Get a list of projects
df_projects <- get_projects()   # we call it 'df_projects' (the default name)

# Get a list of stations
df_stations <- get_stations_from_project("CEMP_Biota", ignore.case = FALSE)

```


## 3. Check parameter   

### Check in METHOD_DEFINITIONS  
Mostly used: VDSI (METHOD_ID = 15631)  
- exception: 2015 used VDSI with METHOD_ID = 28589  
- has been submitted on an individual basis up to 2009, then as average values in 2015 and 2018    
- Intersex does not exist
```{r}

pars <- c("Imposex", "VDSI", "Intersex")

# Check
df <- get_nivabase_selection("*", 
                       "METHOD_DEFINITIONS",
                       "NAME", 
                       pars, 
                       values_are_text = TRUE)
df

# Tabulate species, years and methods
par <- "VDSI"
df_param <- df[grepl(par, df$NAME),]
df1 <- get_nivabase_selection("*", "BIOTA_CHEMISTRY_VALUES", "METHOD_ID", df_param$METHOD_ID)
nrow(df1)
df2 <- get_nivabase_selection("*", "BIOTA_SAMPLES", "SAMPLE_ID", df1$SAMPLE_ID) %>%
  left_join(df1 %>% select(SAMPLE_ID, METHOD_ID))  %>%
  left_join(df %>% select(METHOD_ID, NAME, UNIT, METHOD_REF, MATRIX))  %>%
  left_join(df_species) %>%
  left_join(df_tissue) %>%
  mutate(Year = year(SAMPLE_DATE))

xtabs(~TISSUE_NAME, df2)   # "Hel organisme"
# xtabs(~LATIN_NAME, df2)
df2 %>% 
  count(LATIN_NAME, Year, NAME, UNIT, METHOD_ID, MATRIX)  # 15361
  
# Pick methods
df_methods <- df %>%
  filter(METHOD_ID %in% c(15631))

df_methods

# Pick parameters for later
pars <- df_methods$NAME

```
### b. Add Intersex method  
```{r}

# check1 <- get_nivabase_selection("NAME, LABORATORY, DESCR", "METHOD_DEFINITIONS", "UNIT", '%', values_are_text = TRUE)
# # 167
# check2 <- get_nivabase_selection("NAME, LABORATORY, DESCR", "METHOD_DEFINITIONS", "UNIT", 'PERCENT', values_are_text = TRUE)
# 32, Nivabasen chemistry  

  # name <- df[i, 'NAME']
  # unit <- df[i, 'UNIT']
  # lab <- df[i, 'LABORATORY']
  # method_ref <- df[i, 'METHOD_REF']
  # matrix <- df[i, 'MATRIX']
  # cas <- df[i, 'CAS']

df_intersex <- data.frame(
  NAME = "Intersex",
  UNIT = "PERCENT",
  LABORATORY = "NIVA",
  METHOD_REF = NA,
  MATRIX = "Biota",
  CAS = NA,
  MATRIX_ID = 1
)

make_sql_methods(1, df_intersex) %>% writeLines("clipboard")

# see Milkys_Methods_add_intersex.sql  

```

### Add Intersex to df_methods   
```{r}

# Get the added Intersex method  
df_methods_to_add <- get_nivabase_data(
  "select * from NIVADATABASE.METHOD_DEFINITIONS where TRUNC(ENTERED_DATE) = TO_DATE('2021-09-10', 'yyyy-mm-dd') and ENTERED_BY = 'DHJ';")



if (!"Intersex" %in% df_methods$NAME){
  # Pick methods
  df_methods <- bind_rows(df_methods, df_methods_to_add)
}



```


## 4. Check existing data, former years  
THIS WHOLE THING CAN BE SKIPPED (I think)
Starting from VDSI, via samples  

## 6. Check TBT data for the selected year    

### Get chemical data  
```{r}

df2_specimens_allyrs <- get_specimens_from_stationdata(
  df_stations %>% filter(STATION_CODE %in% dat$STATION_CODE))

# Get data frame of chemical results (30 seconds or so)
df_snailchem <- get_biota_chemistry(
  years = selected_year, 
  specimendata = df2_specimens_allyrs, 
  stationdata = df_stations,
  report_samples = TRUE)

cat("\nDownloaded", nrow(df_snailchem), "records\n")

# Show top of table  
xtabs(~NAME + STATION_CODE, df_snailchem)[1:8,]

```

### Get SAMPLE_ID used  
```{r}

station_sampleid <- df_snailchem %>%
  count(SAMPLE_ID, STATION_CODE)
station_sampleid

```


## 14. BIOTA_CHEMISTRY_VALUES

### a. Read data, if necessary  
Add METHOD_ID and FLAG1  
```{r}

xtabs(~PARAM, dat)

dat_summ <- dat %>%
  filter(PARAM %in% c("VDSI", "Intersex")) %>%
  mutate(
    FLAG1 = as.character(NA)
  ) %>%
  left_join(
    df_methods %>% select(NAME, METHOD_ID),
    by = c("PARAM" = "NAME")
  )

xtabs(~addNA(METHOD_ID), dat_summ)

```

### b. Add SAMPLE_ID  
Which is all we need (see code for 'make_sql_chemistry_values' in '71...functions')
```{r}

cn1 <- dat_summ %>% colnames()

n1 <- nrow(dat_summ)

# Add SAMPLE_ID to chemistry_values data
if (!("SAMPLE_ID" %in% cn1)){    # checks that SAMPLE_ID has not already been added
  dat_summ <- dat_summ %>%
    left_join(station_sampleid,
              by = "STATION_CODE",
              na_matches = "never")
}

n2 <- nrow(dat_summ)

if (n2 > n1){
  stop("Right-hand table not unique!")
} else {
  message("Left join ok, using unique values")
}

if (sum(is.na(dat_summ$SAMPLE_ID)) > 0){
  stop("Some SAMPLE_ID lacking!")
} else {
  message("No SAMPLE_ID lacking")
}


  # dat_summ <- dat_summ %>%
  #   safe_left_join(station_sampleid,
  #                  by = "STATION_CODE",
  #                  na_matches = "never",
  #                  check = "MV")
  # 
```

### c. Make SQLs  
PASTE INTO SQL DEVELOPER TO ADD THE RECORDS   
Note to DHJ: use "SQL developer (latest version)" from desktop. Don't use the start menu.  
- remember `commit;` after running the insert sentences   
- See 'Milkys_2020_Intersex_Biota_chemistry_values.sql' in folder 'SQL_scripts'
```{r}
# make_sql_chemistry_values(1, data = dat_summ)

sql_list <- 1:nrow(dat_summ) %>% 
  map_chr(make_sql_chemistry_values, data = dat_summ)

i <- 1:length(sql_list)
sql <- paste(sql_list[i], collapse = ";\n")
sql <- paste0(sql, ";\n")
writeLines(sql, "clipboard-1024")   # copies SQLs to clipboard - go to SQL Developer and paste
                                    # "clipboard-1024" instead of "clipboard": increases available
                                    #    for the clipboard

cat("Number of SQLs: \n")
length(sql_list)  # 144

cat("\nSample of SQLs: \n")
sql_list[1:3]


```

## 15. Reread data  
```{r}

# Get data frame of chemical results (30 seconds or so)
df_snailchem <- get_biota_chemistry(
  years = selected_year, 
  specimendata = df2_specimens_allyrs, 
  stationdata = df_stations,
  report_samples = TRUE)

cat("\nDownloaded", nrow(df_snailchem), "records\n")

xtabs(~NAME + STATION_CODE, df_snailchem)

```
### Plot data   
Very boring plot
```{r, fig.width=5, fig.height=2}

library(ggplot2)

df_snailchem %>%
  filter(NAME %in% c("VDSI", "Intersex")) %>%
  ggplot(aes(STATION_CODE, VALUE, color = NAME)) + 
  geom_point()

```




