---
title: "814 - Import cod siloxan data from NILU to Nivabasen, 2020"
author: "DHJ"
date: "2 9 2019"
output: 
  html_document:
    keep_md: true
---

Make tables for inserting into NIVAbase   
  
For eider duck and cod (siloxans only) data from NILU  


## 0. Libraries and functions
```{r}

library(dplyr)
library(purrr)
library(lubridate)
library(stringr)
library(tidyr)
library(ggplot2)
library(niRvana)
library(safejoin)   # https://github.com/moodymudskipper/safejoin

source("812_Import_til_NIVAbasen_NILU_functions.R")

# "Restart" for lazy people that don't want to write the 
#    database password again (deletes all objects except database password/username)
if (FALSE){
  obj <- ls()
  obj <- obj[!obj %in% c("db_privkey", "db_pubkey", "db_pwd", "db_username")]
  rm(list = obj)
  rm(obj)
}


```

### For connectng to Nivabasen    
Store your username and password to R (only once per R session)  
```{r}

set_credentials()

# Check these codes (press F2) for reminder about what comes from which tables:
# niRvana::get_biota_chemistry
# niRvana::get_samples_from_sampleid

```

## 1. Read data


### a. Year
```{r}

selected_year <- 2020

```


### b. Get most recent data produced in Jupyterhub  
We only keep the data for the curret year
```{r}
# dir("Data")
# Data at sample level (chemical data, biological efect parameters and VDSI)
data_all <- readRDS(file = "Files_to_Jupyterhub_2020/01_df_2020_notstandard_2021-06-15.rds")

```


### c. Get samples in Labware file     
```{r}

# March-December data
sql <- paste(
  "select * from NIVADATABASE.LABWARE_CHECK_SAMPLE", 
  "where extract(YEAR from SAMPLED_DATE) =", selected_year, 
  "and extract(MONTH from SAMPLED_DATE) >= 3", 
  "and UPPER(PROSJEKT) like '%MILKYS%';")
sql
df_labware_01 <- get_nivabase_data(sql)

# January-February next year data (if any)
sql <- paste(
  "select * from NIVADATABASE.LABWARE_CHECK_SAMPLE", 
  "where extract(YEAR from SAMPLED_DATE) =", selected_year + 1, 
  "and extract(MONTH from SAMPLED_DATE) <= 2", 
  "and UPPER(PROSJEKT) like '%MILKYS%';")
df_labware_02 <- get_nivabase_data(sql)

# Remove everything except Eider duck stations
# Extract 'Specimen_no' and 'TISSUE_NAME'
df_samples <- bind_rows(df_labware_01, df_labware_02) %>%
  # Add 'Specimen_no'
  mutate(DESCRIPTION = gsub("<f8>", "ø", DESCRIPTION, fixed = TRUE),
         DESCRIPTION = gsub("<c5>", "Å", DESCRIPTION, fixed = TRUE),
         DESCRIPTION = gsub("<e5>", "å", DESCRIPTION, fixed = TRUE),
         DESCRIPTION = gsub("<c6>", "Æ", DESCRIPTION, fixed = TRUE),
         DESCRIPTION = gsub("<e6>", "æ", DESCRIPTION, fixed = TRUE),
         DESCRIPTION2 = DESCRIPTION %>% stringr::str_sub(start = nchar(AQUAMONITOR_CODE) + 1),
         Specimen_no = stringr::str_extract(DESCRIPTION2, "[0-9]+") %>% as.numeric(),
         TISSUE_NAME = stringr::str_sub(TISSUE, start = 4))

```


### d. Get some generic lookup tables  
* Tissues  
* Species names  
* Projects
* Methods  
```{r}

df_tissue <- get_nivabase_data(
  "select TISSUE_ID,TISSUE_NAME from NIVADATABASE.BIOTA_TISSUE_TYPES")

df_taxon <- get_nivabase_selection(
  "NIVA_TAXON_ID, LATIN_NAME", 
  "TAXONOMY", 
  "LATIN_NAME",
  unique(dat_eider_nivabase$LATIN_NAME), values_are_text = TRUE
  )

# Get a list of projects
df_projects <- get_projects()   # we call it 'df_projects' (the default name)

# Get a list of stations
df_stations <- get_stations_from_project("CEMP_Biota", ignore.case = FALSE)

# Get df_methods (NILU methods only)
df_methods <- get_nivabase_selection(
  "*", 
  "METHOD_DEFINITIONS", 
  "LABORATORY", 
  "NILU", values_are_text = TRUE)

```




## 2. Data to insert in NIVAbase
1. Eider duck data are from specimens/samples that MOSTLY ARE in the NIVAbase (for 2018 data, they were NOT) 
    - Sample ID given as Labware ID, e.g. 'NR-2020-09188' 
    - New records in BIOTA_SAMPLES only (SAMPLE_NO taken from data, get SAMPLE_ID from database)
    - Also note that although pairs of eggs and blood were unrelated (I think!),
      the same SPECIMEN_ID was giben for pairs of blood/egg samples     

2. NILU cod data (siloxans; D4,D5 and D6) are from specimens that MOSTLY ARE in the NIVAbase, 
but new samples (liver) are numbered corresponding to the muscle samples
   - Sample ID should normally be given as station number + sample number (1-100)
   - This was added manually in the excel file (Samperapport_edited.xlsx)
     based on the 'fake Labware ID' (e.g. 'NR-2020-09188') added (which doesn't correspond to
     the actual Labware IDs)
   - Use existing record in BIOTA_SINGLE_SPECIMENS (existing SPECIMEN_ID)
   - New records in BIOTA_SAMPLES_SPECIMENS - relationship between SAMPLE_NO and SPECIMEN_ID
   should be the same as for muscle
   - New records in BIOTA_SAMPLES (new SAMPLE_ID, SAMPLE_NO taken from data)

3. Biological effects in cod - from specimens that ARE in the NIVAbase, with the 
exception of a single cod individual (Fish_no = 15, station = 53B). New samples (bile, blood,
liver fraction). Samples are numbered corresponding to the muscle samples (as NILU cod)
    - Use existing records in BIOTA_SINGLE_SPECIMENS (existing SPECIMEN_ID)  
    - New records in BIOTA_SAMPLES_SPECIMENS (new SAMPLE_ID, but existing SPECIMEN_ID)  
    - New records in BIOTA_SAMPLES (new SAMPLE_ID, SAMPLE_NO taken from data)  


### a. NILU data  
```{r}

dat_cod <- readRDS("Data/210_dat_cod-siloxans_2020.rds") %>%
  mutate(TEXT_ID = sub("Nr. ", "NR-", Sample_no, fixed = TRUE))

# xtabs(~TEXT_ID + Group, dat_cod)

```


### b. Add station code    
- Starting 'data_for_input'  
```{r}

# Make lookup table for finding station code + ID
df_lookup_stations <- data_all %>%
  distinct(STATION_CODE, STATION_NAME, STATION_ID)

# Check contents of 'Sample' field
table(substr(dat_cod$Sample, 13, 20))

# Start 'data_for_input'  
data_for_input_1 <- dat_cod %>%
  mutate(
    STATION_CODE = case_when(
      grepl("oslo", Sample) ~ "30B",
      grepl("bergen", Sample) ~ "24B",
      grepl("svalbard", Sample) ~ "19B",
      grepl("tromsø", Sample) ~ "43B2",
      grepl("varanger", Sample) ~ "10B",
      TRUE ~ "",
  )) %>%
  left_join(df_lookup_stations, by = "STATION_CODE") %>%
  mutate(
    LATIN_NAME = "Gadus morhua"
  ) %>%
  rename(
    TISSUE_NAME = Tissue,
    UNIT = Unit
  )

table(addNA(data_for_input_1$STATION_ID))

```

### c. Tissue, check  
If 'Tissues are correct' we can just keep this column
```{r}

df_tissue <- get_nivabase_data("select TISSUE_ID,TISSUE_NAME from NIVADATABASE.BIOTA_TISSUE_TYPES")

cat("Tissues:\n")
unique(data_for_input_1$TISSUE_NAME)

check <- unique(data_for_input_1$TISSUE_NAME) %in% df_tissue$TISSUE_NAME
if (sum(!check) == 0){
  cat("Tissue names exist in NIVAbasen\n")
} else {
  warning("NOT ALL tissue names exist in NIVAbasen!")
}
```


### d. Species, check  
If 'Species are correct' we can just keep this column
```{r}
cat("Species:\n")
unique(data_for_input_1$LATIN_NAME)

df_species <- get_nivabase_selection("SPECIES_ID,LATIN_NAME", "SPECIES", 
                                     "LATIN_NAME", unique(data_for_input_1$LATIN_NAME), values_are_text = TRUE)

check <- unique(data_for_input_1$LATIN_NAME) %in% df_species$LATIN_NAME
if (sum(!check) == 0){
  cat("Species names exist in NIVAbasen\n")
} else {
  warning("NOT ALL species names exist in NIVAbasen!")
}

```

### e. Date - not needed       
```{r}
# from 
# data_for_input_1 %>% 
#   count(STATION_CODE, SAMPLE_DATE)
```

### f. Parameters  
- We just check what we used last year  
```{r}

df_lookup_methods_1 <- get_nivabase_selection(
  "*", 
  "METHOD_DEFINITIONS", 
  "NAME", 
  c("D4","D5","D6"), values_are_text = TRUE)

df_lookup_methods_2 <- df_lookup_methods_1 %>%
  filter(LABORATORY == "NILU" & grepl("w.w.", UNIT))

df2 <- get_nivabase_selection("*", "BIOTA_CHEMISTRY_VALUES", "METHOD_ID", df_lookup_methods$METHOD_ID)
df3 <- get_nivabase_selection("*", "BIOTA_SAMPLES", "SAMPLE_ID", df2$SAMPLE_ID)
df_lookup_methods <- df_lookup_methods_2 %>% 
  select(METHOD_ID, NAME, UNIT) %>%
  right_join(df2, by = "METHOD_ID") %>%
  left_join(df3, by = "SAMPLE_ID") %>%
  filter(year(SAMPLE_DATE) == 2019) %>%
  distinct(NAME, UNIT, METHOD_ID)

if(sum(table(df_lookup_methods$NAME) > 1)){
  stop("Some values of NAME has more than one METHOD_ID!")
} else {
  message("All NAME values are unique in 'df_lookup_methods'")
}

# Add METHOD_ID 
data_for_input_2 <- data_for_input_1 %>%
  left_join(
    df_lookup_methods %>% rename(UNIT_Method = UNIT),
    by = c("Parameter" = "NAME")
  )

if (sum(is.na(data_for_input$METHOD_ID)) > 0){
stop("Some values of Parameter not found in 'df_lookup_methods'!")
} else {
  message("All METHOD_ID found")
}

# df4 <- get_nivabase_selection("*", "BIOTA_SAMPLES_SPECIMENS", "SPECIMEN_ID", df3$SAMPLE_ID)


```
### g. Units, check manually    
- Must be correspondence between UNIT and UNIT_Method!  
```{r}

data_for_input %>%
  distinct(Parameter, UNIT, UNIT_Method)

```





## 3. Plot connections  
```{r, fig.width = 8, fig.height=4}

stations <- dat_cod$ STATION_CODE %>% unique()

for (station in stations){
  gg <- allexisting_samples %>%
    left_join(allexisting_sampspec, by = "SAMPLE_ID") %>%
    filter(STATION_CODE %in% station) %>%
    mutate(SAMPLE_NO = factor(SAMPLE_NO), SPECIMEN_NO = factor(SPECIMEN_NO)) %>%
    ggplot(aes(SAMPLE_NO, SPECIMEN_NO)) +
    geom_point() +
    facet_grid(cols = vars(TISSUE_NAME)) +
    theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
    labs(title = station)
  print(gg)
}

```




## APPENDIX  
```{r}

```{r}

df1 <- get_nivabase_selection(
  "*",
  "BIOTA_SINGLE_SPECIMENS",
  "STATION_ID",
  69750,
  extra_where = " and extract(YEAR from DATE_CAUGHT) = 2020")

df2 <- get_nivabase_selection(
  "*",
  "BIOTA_SAMPLES_SPECIMENS",
  "SPECIMEN_ID",
  df1$SPECIMEN_ID)

df3 <- get_nivabase_selection(
  "*",
  "BIOTA_SAMPLES",
  "SAMPLE_ID",
  df2$SAMPLE_ID)

df3 <- get_nivabase_selection(
  "*",
  "BIOTA_SAMPLES",
  "SAMPLE_ID",
  df2$SAMPLE_ID)

df3b <- get_nivabase_selection(
  "*",
  "BIOTA_SAMPLES",
  "SAMPLE_ID",
  245960)

df3b <- get_nivabase_selection(
  "*",
  "BIOTA_SAMPLES",
  "STATION_ID",
  69750,
  extra_where = " and extract(YEAR from SAMPLE_DATE) = 2020")

```
```

