---
title: "816_CCP_chlorinated_paraffins"
author: "DHJ"
date: "2022-12-10"
output: html_document
---


## 1. Libraries
```{r setup, include=FALSE}
# install.packages("niRvana")

library(dplyr)
library(ggplot2)
library(lubridate)
library(stringr)

# And then the niRvana package. If you need to install it, then run thee next line also:
# devtools::install_github("NIVANorge/niRvana")
library(niRvana)

source("802_Get_NIVAbasen_data_functions.R")

```

### Settings  
```{r}

sampling_year <- 2021

```


## 2. Check data in LIMS

- Project 'O 210200.ANAIN' is hard-coded  
- NOTE: no selection for year (but see 3)

```{r}

df_labware_meta <- get_nivabase_data("select * from NIVADATABASE.LABWARE_CHECK_SAMPLE where PROSJEKT = 'O 210200.ANAIN'")
nrow(df_labware_meta) # 982

# get_nivabase_data("select * from NIVADATABASE.LABWARE_IMPORT where rownum < 4")

# One station  
test <- get_nivabase_data(
  "select * from NIVADATABASE.LABWARE_IMPORT where PROSJEKT = 'O 210200.ANAIN' and AQUAMONITOR_CODE = '30B'")

# xtabs(~addNA(PARAM), test)  
# - MCCP and SCCP - not giving 
# xtabs(~addNA(PARAM_I_AQUAMONITOR), test)  # nothing  


```

### Normal indication of <LOQ data  

- ENTRY_QUALIFIER is NA or "<"  
- Example: BDEs  

```{r}

test %>%
  filter(REPORTED_NAME %in% c("BDE99", "BDE100", "BDE209")) %>%
  xtabs(~REPORTED_NAME + addNA(ENTRY_QUALIFIER), .)

```



#### Indication of <LOQ data for CCPs    

- FORMATTED_ENTRY is a number or "ND"  

```{r}

test %>%
  filter(REPORTED_NAME %in% c("SCCP eksl. LOQ", "SCCP inkl. LOQ", "MCCP eksl. LOQ", "MCCP inkl. LOQ")) %>%
  xtabs(~REPORTED_NAME + (FORMATTED_ENTRY == "ND"), .)

```


#### SAMPLE_NUMBER    
```{r}

test %>%
  count(SAMPLE_NUMBER) %>%
  arrange(SAMPLE_NUMBER) %>%
  group_by(n) %>%
  summarize(
    n = n(),
    min = min(SAMPLE_NUMBER),
    max = max(SAMPLE_NUMBER),
    diff = max-min)

```


#### TEXT_ID    
```{r}

test %>%
  count(TEXT_ID) %>%
  arrange(TEXT_ID) %>%
  mutate(TEXT_ID_no = as.numeric(substr(TEXT_ID, 9, 13))) %>%
  group_by(n) %>%
  summarize(
    n = n(),
    min = min(TEXT_ID_no),
    max = max(TEXT_ID_no),
    diff = max-min)

```


## 3. Get data from LIMS  

### Select by project and parameters   

```{r}

# paste(names(test), collapse = ", ")

vars <- "SAMPLE_TYPE, SAMPLED_DATE, AQUAMONITOR_ID, AQUAMONITOR_CODE, REPORTED_NAME, ENTRY_QUALIFIER, FORMATTED_ENTRY, NUMERIC_ENTRY, UNITS, DESCRIPTION, STATUS, STATUSX, TEXT_ID, PROSJEKT, ANALYSEOPPDRAG, SAMPLE_NUMBER"

df_ccp_all <- get_nivabase_selection(
  vars,
  "LABWARE_IMPORT",
  "REPORTED_NAME",
  c("SCCP eksl. LOQ", "SCCP inkl. LOQ", "MCCP eksl. LOQ", "MCCP inkl. LOQ"), values_are_text = TRUE, 
  extra_sql = "and PROSJEKT = 'O 210200.ANAIN'") %>%
  mutate(
    Month = month(SAMPLED_DATE),
    Year = year(SAMPLED_DATE),
    MYEAR = case_when(
      Month <= 2 ~ Year-1,
      Month >= 3 ~ Year))

xtabs(~MYEAR, df_ccp_all)
xtabs(~STATUS + STATUSX + MYEAR, df_ccp_all)

# View(df_ccp_all %>% filter(MYEAR == 2022))

```

### Select by status (A = accepted) and year  

```{r}

df_ccp <- df_ccp_all %>%
  filter(
    STATUS %in% "A",
    MYEAR  %in% sampling_year)

```


#### Tabulate data  

- Eksl. LOQ may be numeric or 'ND'  
- Inkl. LOQ is numeric  
- For one cod station, many samples have 'T' for both   

```{r}

check1 <- df_ccp %>%
  filter(REPORTED_NAME %in% c("SCCP eksl. LOQ", "SCCP inkl. LOQ")) %>%
  mutate(Type = case_when(
    FORMATTED_ENTRY == "ND" ~ "ND",
    FORMATTED_ENTRY == "T" ~ "T",
    !is.na(as.numeric(sub(",", ".", FORMATTED_ENTRY))) ~ "Numeric",
    TRUE ~ "??"))

# xtabs(~AQUAMONITOR_CODE, check1)

cat("\n\n\n")
cat("Eksl. LOQ \n")
xtabs(~AQUAMONITOR_CODE + Type, check1 %>% filter(REPORTED_NAME %in% "SCCP eksl. LOQ")) %>% head(10)

cat("==================================\n")
cat("Inkl. LOQ \n")
xtabs(~AQUAMONITOR_CODE + Type, check1 %>% filter(REPORTED_NAME %in% "SCCP inkl. LOQ")) %>% head(10)

```

## 4. Get adapted SCCP data  

#### Viewing (for understanding)        
```{r}

check <- df_ccp %>%
  filter(REPORTED_NAME %in% c("SCCP eksl. LOQ", "SCCP inkl. LOQ")) %>%
  select(AQUAMONITOR_ID, AQUAMONITOR_CODE, TEXT_ID, REPORTED_NAME, FORMATTED_ENTRY) %>%
  tidyr::pivot_wider(names_from = REPORTED_NAME, values_from = FORMATTED_ENTRY) %>%
  arrange(AQUAMONITOR_ID, AQUAMONITOR_CODE, TEXT_ID)


## View data  

# View(check)

```


#### Make data for export      

- Basis is "eksl." data  
- When "eksl." data is numeric, use this (over LOQ) value   
- When "eksl." = "ND", set FLAG1 = < and use the "inkl. LOQ" value as the LOQ value     

```{r}

params <- c("SCCP eksl. LOQ", "SCCP inkl. LOQ")

df_selected <- df_ccp %>%
  filter(REPORTED_NAME %in% params & !FORMATTED_ENTRY %in% "T") %>%
  select(AQUAMONITOR_ID, AQUAMONITOR_CODE, TEXT_ID, REPORTED_NAME, FORMATTED_ENTRY) %>%
  mutate(Value_temp = as.numeric(sub(',', '.', FORMATTED_ENTRY)))

check <- xtabs(~FORMATTED_ENTRY, df_selected %>% filter(is.na(Value_temp)))
if (identical(names(check), "ND")){
  message("All FORMATTED_ENTRY are either numeric or 'ND', as expected")
} else {
  stop("Some not-numeric data are something else than 'ND'! Check table called 'check'.")
}

df_SCCP <- df_selected %>%
  select(AQUAMONITOR_ID, AQUAMONITOR_CODE, TEXT_ID, REPORTED_NAME, Value_temp) %>%
  tidyr::pivot_wider(names_from = REPORTED_NAME, values_from = Value_temp) %>%
  arrange(AQUAMONITOR_ID, AQUAMONITOR_CODE, TEXT_ID) %>%
  mutate(
    VALUE = case_when(
      !is.na(`SCCP eksl. LOQ`) ~ `SCCP eksl. LOQ`,
      is.na(`SCCP eksl. LOQ`) ~ `SCCP inkl. LOQ`,
    ),
    FLAG1 = case_when(
      !is.na(`SCCP eksl. LOQ`) ~ as.character(NA),
      is.na(`SCCP eksl. LOQ`) ~ "<",
    )
  )


```


## 5, Get adapted MCCP data  

#### Viewing (for understanding)        
```{r}

check <- df_ccp %>%
  filter(REPORTED_NAME %in% c("MCCP eksl. LOQ", "MCCP inkl. LOQ")) %>%
  select(AQUAMONITOR_ID, AQUAMONITOR_CODE, TEXT_ID, REPORTED_NAME, FORMATTED_ENTRY) %>%
  tidyr::pivot_wider(names_from = REPORTED_NAME, values_from = FORMATTED_ENTRY) %>%
  arrange(AQUAMONITOR_ID, AQUAMONITOR_CODE, TEXT_ID)


## View data  

# View(check)

```


#### Make data for export  

- Basis is "eksl." data  
- When "eksl." data is numeric, use this (over LOQ) value   
- When "eksl." = "ND", set FLAG1 = < and use the "inkl. LOQ" value as the LOQ value     

```{r}

params <- c("MCCP eksl. LOQ", "MCCP inkl. LOQ")

df_selected <- df_ccp %>%
  filter(REPORTED_NAME %in% params & !FORMATTED_ENTRY %in% "T") %>%
  select(AQUAMONITOR_ID, AQUAMONITOR_CODE, TEXT_ID, REPORTED_NAME, FORMATTED_ENTRY) %>%
  mutate(Value_temp = as.numeric(sub(',', '.', FORMATTED_ENTRY)))

check <- xtabs(~FORMATTED_ENTRY, df_selected %>% filter(is.na(Value_temp)))
if (identical(names(check), "ND")){
  message("All FORMATTED_ENTRY are either numeric or 'ND', as expected")
} else {
  stop("Some not-numeric data are something else than 'ND'! Check table called 'check'.")
}

df_MCCP <- df_selected %>%
  select(AQUAMONITOR_ID, AQUAMONITOR_CODE, TEXT_ID, REPORTED_NAME, Value_temp) %>%
  tidyr::pivot_wider(names_from = REPORTED_NAME, values_from = Value_temp) %>%
  arrange(AQUAMONITOR_ID, AQUAMONITOR_CODE, TEXT_ID) %>%
  mutate(
    VALUE = case_when(
      !is.na(`MCCP eksl. LOQ`) ~ `MCCP eksl. LOQ`,
      is.na(`MCCP eksl. LOQ`) ~ `MCCP inkl. LOQ`,
    ),
    FLAG1 = case_when(
      !is.na(`MCCP eksl. LOQ`) ~ as.character(NA),
      is.na(`MCCP eksl. LOQ`) ~ "<",
    )
  )


```


## 6. Combine  

```{r}

df_ccp_ready1 <- bind_rows(
  df_SCCP %>% select(AQUAMONITOR_ID, AQUAMONITOR_CODE, TEXT_ID, VALUE, FLAG1) %>% mutate(NAME = "SCCP"),
  df_MCCP %>% select(AQUAMONITOR_ID, AQUAMONITOR_CODE, TEXT_ID, VALUE, FLAG1) %>% mutate(NAME = "MCCP")
)

```


#### Plot  

```{r, fig.width=9, fig.height=7}

df_ccp_ready1 %>% 
  mutate(LOQ = case_when(
    is.na(FLAG1) ~ "Over LOQ",
    !is.na(FLAG1) ~ "Under LOQ")
    ) %>% 
  ggplot(aes(AQUAMONITOR_CODE, VALUE, color = LOQ)) +
  geom_jitter(width = 0.15) +
  facet_grid(vars(NAME))

```




### Add Vannmiljø stations   

- Checking that Kilde_id = AQUAMONITOR_ID with 'NIVA@' in front of it  
- As expected, 20B is not found  
- But can easily be created by adding 'NIVA@' in front of AQUAMONITOR_ID  

```{r}

df_vm <- readxl::read_excel("Files_to_Vannmiljo/2020/ooa63pi kvalsjekk 18.01.2022.xlsm", sheet = "LOK")

df_vm_stations <- df_vm %>%
  select(Kilde_id, Navn) %>%
  mutate(AQUAMONITOR_ID = sub('NIVA@', '', Kilde_id, fixed = TRUE) %>% as.numeric())

check1 <- df_ccp_ready1 %>%
  left_join(df_vm_stations, by = "AQUAMONITOR_ID")

check2 <- check1 %>%
  filter(is.na(Kilde_id)) %>%
  count(AQUAMONITOR_ID, AQUAMONITOR_CODE)

check2

```

## 7. Final data  

```{r}

df_meta <- df_ccp %>%
  distinct(TEXT_ID, SAMPLED_DATE)

df_ccp_ready2 <- df_ccp_ready1 %>%
  mutate(
    Kilde_id = paste0('NIVA@', AQUAMONITOR_ID)
  ) %>%
  left_join(
    df_meta, by = "TEXT_ID"
  ) 


```

## 8. Save

```{r}

saveRDS(df_ccp_ready2, "Files_to_Jupyterhub_2021/816_CCP_2021.rda")

writexl::write_xlsx(df_ccp_ready2, "Files_to_Jupyterhub_2021/816_CCP_2021.xlsx")

```




