---
title: "Report for parameter group"
author: "DHJ"
date: "22 6 2022"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
---

```{r, echo = FALSE}

knitr::opts_chunk$set(echo = FALSE, results = 'hold')
# knitr::opts_chunk(results = 'hold')

```


```{r}

current_year <- 2021

```


```{r packages, results='hide', message=FALSE, warning=FALSE}

library(dplyr)
# library(tidyr)
library(ggplot2)
library(lubridate)
library(flextable)
# library(glue)
library(readxl)
library(purrr)
library(leftcensored)   # DHJ's package (https://github.com/DagHjermann/leftcensored)
library(scico)          # colour palettes incl. "Vik" (https://github.com/thomasp85/scico)

library(ggiraph)
library(cowplot)

# library(safejoin) # https://github.com/moodymudskipper/safejoin

# source("01_Get_chemical_data_NIVAbasen_functions.R")  # for get_standard_parametername()
source("824_Report_for_scientists_functions.R")
source("825_Report_paramgroup_functions.R", encoding = "UTF-8")

```


```{r get_data, warning=FALSE}

params <- get_parametervalues("metals")

# debugonce(get_data_tables)
dat_sample_fish <- get_data("metals", "fish")
dat_sample_muss <- get_data("metals", "mussel")

dat_median <- get_medians(dat_sample_fish, dat_sample_muss)

dat_median_fish <- dat_median %>%
  filter(LATIN_NAME %in% c("Gadus morhua", "Platichthys flesus"))
dat_median_mussel <- dat_median %>%
  filter(LATIN_NAME %in% c("Mytilus edulis"))


```



## Ratio to proref in last year       


### Fish  
* Note: 
    - Region for colour  
    - Concentrations given in tooltip (should do something for the < values for minimum values)   
```{r, fig.height=6, fig.width=11, warning=FALSE}

# **Bruk farger for geografi (kalde farger for nord, varme farger for sør)**  

dat_prorefplot2 <- dat_median_fish %>% 
  filter(MYEAR == current_year) %>%
  mutate(
    Prorefratio = VALUE_WW_med/Proref,
    Unit = gsub("_P_", "/", UNIT, fixed = TRUE) %>% tolower(),
    Tooltip = paste0(Station, "<br>Conc.: (min-median-max): ", VALUE_WW_min, "-", VALUE_WW_med, "-", VALUE_WW_max, " ", Unit))

# str(dat_prorefplot2)

gg <- ggplot(dat_prorefplot2, aes(PARAM, y = Prorefratio)) +
  geom_hline(yintercept = 1) +
  geom_boxplot() +
  geom_jitter_interactive(aes(fill = Region, tooltip = Tooltip, data_id = STATION_CODE), pch = 21, size = 2, width = 0.1) +
  # scale_fill_distiller("Along coast\n(far N/E = blue)", palette = "RdBu", direction = 1) +  # Geogr_position
  scale_fill_brewer("Along coast\n(far N/E = blue)", palette = "RdBu", direction = 1) +
  theme_bw() +
  ggeasy::easy_rotate_x_labels(angle = -45) +
  labs(title = "Ratio concentration/proref, metals in fish 2021")
# gg

# gg <- gg + coord_flip()
# gg

ggr <- girafe(ggobj = plot_grid(gg + guides(fill = "none") + labs(subtitle = "Medians, ordinary scale"),
                         gg + scale_y_log10() + labs(subtitle = "Medians, log scale"), 
                         rel_widths = c(1,1.35)), 
       width_svg = 10, height_svg = 4)

ggr <- girafe_options(ggr, opts_hover(css = "fill:wheat;stroke:orange;r:5pt;") )

ggr
                       
```


```{r test_dotplot}

if (FALSE){
  
  # Looks nice but has no colours for geogrphy
  ggplot(dat_prorefplot2, aes(x = PARAM, y = Prorefratio)) +
    geom_dotplot(binaxis = "y", stackdir = "center", position = "dodge", binwidth = 0.07)
  
  # Looks nice but has no colours for geogrphy
  ggplot(dat_prorefplot2, aes(x = PARAM, fill = Area, y = Prorefratio)) +
    geom_dotplot(binaxis = "y", stackdir = "center", position = "dodge", binwidth = 0.07, )
  
}



```


### Mussel    
*Regions to be added*  
```{r, fig.height=7, fig.width=9, warning=FALSE}

# **Bruk farger for geografi (kalde farger for nord, varme farger for sør)**  

dat_prorefplot2 <- dat_median_mussel %>% 
  filter(MYEAR == current_year) %>%
  mutate(Geogr_position = as.numeric(Station))

gg <- ggplot(dat_prorefplot2, aes(PARAM, VALUE_WW_med/Proref)) +
  geom_hline(yintercept = 1) +
  geom_boxplot() +
  geom_jitter_interactive(aes(fill = Geogr_position, tooltip = Station), pch = 21, size = 2, width = 0.1) +
  scale_fill_distiller("Along coast\n(far N/E = blue)", palette = "RdBu", direction = 1) +
  theme_bw() +
  ggeasy::easy_rotate_x_labels(angle = -45) +
  labs(title = "Ratio concentration/proref, metals in mussel 2021")

girafe(ggobj = plot_grid(gg + guides(fill = "none") + labs(subtitle = "Medians, ordinary scale"),
                         gg + scale_y_log10() + labs(subtitle = "Medians, log scale"), 
                         rel_widths = c(1,1.35)), 
       width_svg = 10, height_svg = 4)



```




```{r}

```





