---
title: "79_Check_Vannmiljo_importfile_2015data."
author: "DHJ"
date: "2 9 2019"
output: 
  html_document:
    keep_md: true
---


Check and fix 2019 data before importing data to Vannmiljø.

- Deal with fixes done in Milkys2 script 101 (in Jupyterhub) **after** getting Nivabasen data:
    1. Station 26A2 with year 2020 instead of 2019   
    2. Records with PARAM = CA should be changed to PARAM = CD  
    3. st. 80B: One AG lacking (SAMPLE_NO2 = 3), values shoud be <0.05 MG_P_KG    
    4. st. 98B1: Zn in liver is 1000x to high for 2 samples (12000 and 11000 intead of 12 and 11 MG_P_KG)   


## 0a. Libraries and functions
```{r}
library(dplyr)
library(purrr)
library(lubridate)
library(stringr)
# library(ggplot2)
library(niRvana)
library(ggplot2)
library(readxl)

library(safejoin) # https://github.com/moodymudskipper/safejoin

source("71_Import_til_NIVAbasen_functions.R")

# "Restart" for lazy people that don't want to write the 
#  password again (deletes all objects except database password/username)
if (FALSE){
  obj <- ls()
  obj <- obj[!obj %in% c("db_privkey", "db_pubkey", "db_pwd", "db_username")]
  rm(list = obj)
  rm(obj)
}

```


## 0b. Year
```{r}

selected_year <- 2019

```

## 0c. Lookup tables  

### Tissue from medium lookup  
```{r}
lookup_tissue <- read.csv(textConnection("
Medium_id,TISSUE_NAME
BA,Galle
BB,Whole soft body
BE,Egg
BL,Lever
BM,Muskel
BO,Blod
VS,Saltvann
"), header = TRUE)


```

### Unit from Enhet_id lookup  
```{r}

lookup_unit <- read.csv(textConnection("
Enhet_id,Unit
2,% v.v.
3,None
7,µg/kg v.v.
30,g/kg
42,mg/kg v.v.
93,pmol/min/mg protein
94,ng/min/mg protein
"), header = TRUE, stringsAsFactors = FALSE)

```




## 1. Read data   

### a. Read file made by NIVA, used by Miljødir 
```{r}

fn_read <-  "Eksport Vannmiljø/2019_data/NIVA Milkys 2019-data 2020-09-04 versjon 03.xlsx"
fn_write <-  "Eksport Vannmiljø/2019_data/NIVA Milkys 2019-data 2020-09-04.xlsx"  # used in the end, 

dat_stations_original <- read_excel(fn_read, sheet = "Vannlokaliteter")
dat_export_original <- read_excel(fn_read, sheet = "Vannregistreringer")

dat_stations <- dat_stations_original %>%
  select(Kilde_id, Navn)

dat_export <- dat_export_original %>%
  mutate(Enhet_id = as.numeric(Enhet_id)) %>%
  left_join(dat_stations, by = c("Vannlok_kode" = "Kilde_id")) %>%
  left_join(lookup_tissue, by = "Medium_id") %>%
  left_join(lookup_unit, by = "Enhet_id")

cat("Number of observations per medium:\n")
xtabs(~Medium_id, dat_export)

cat("\n")
cat("Number of observations per medium:\n")
xtabs(~TISSUE_NAME, dat_export)

cat("\n")
cat("Number of parameters per medium:\n")
tab <- xtabs(~TISSUE_NAME + Parameter_id, dat_export)
apply(tab > 0, 1, sum)

cat("\n")
cat("Number of stations per medium:\n")
tab <- xtabs(~TISSUE_NAME + Vannlok_kode, dat_export)
apply(tab > 0, 1, sum)

# xtabs(~Parameter_id, dat_export)

# head(dat_export)

if (FALSE){
  # Should be number of samples per station
  # Can read sheet "Vannlokaliteter" to get actual station names 
  #   (couple using 'Vannlok_kode')  
  dat_export %>%
    count(Vannlok_kode, Medium_id, LatinskNavn_id)
}


```


### b. Data used in Milkys  
```{r}

dat <- readRDS("Input_data_2019/101_data_updated_2020-08-05.rds") %>%
  filter(MYEAR %in% 2019)

cat("\nNumber of parameters per tissue:\n")
tab <- xtabs(~TISSUE_NAME + PARAM, dat)
apply(tab > 0, 1, sum)

cat("\nNumber of stations per tissue:\n")
tab <- xtabs(~TISSUE_NAME + STATION_CODE, dat)
apply(tab > 0, 1, sum)

```
### c. Station data from Excel  
```{r}

data_stations <- readxl::read_excel("Input_data/Kartbase.xlsx")

# Note that 
#     the current station code is 'stasjonskode', not "STATION_CODE"
#     `catch LAT__1` and `catch LONG__1` are currect (planned?) positions
#     the current station name is 'stasjonsnavn', not "STATION_CODE"
#     but station name for report is `Til Rapport`
#         station name for maps are `Til Kart`

# Check the two station code columns
# data_stations %>%
#   filter(STATION_CODE != stasjonskode)    # zero rows

lookup_stations <- data_stations %>%
  select(STATION_CODE, STATION_NAME) %>%
  filter(!is.na(STATION_CODE) & !is.na(STATION_NAME))

# Check
check <- lookup_stations %>%
  group_by(STATION_CODE) %>%
  mutate(n = n()) %>%
  filter(n > 1)

if (nrow(check) > 0 )
  stop("Duplicates!")


```

### for reference, previous year's data (2018 data)
```{r}

fn <-  "Eksport Vannmiljø/2018_data/NIVA Milkys 2018-data 2019-09-13_B.xlsx"

dat_stations_2018 <- read_excel(fn, sheet = "Vannlokaliteter") %>%
  select(VannlokalitetId, Navn)

dat_export_2018 <- read_excel(fn, sheet = "Vannregistreringer") %>%
  left_join(dat_stations_2018, by = c("Vannlok_kode" = "VannlokalitetId")) %>%
  left_join(lookup_tissue, by = "Medium_id")


```


## 2. Some tables    

### Check dates/years  
```{r}

dat_export %>% 
  xtabs(~ + addNA(TISSUE_NAME), .)

```

### Units  
```{r}

#tab <- xtabs(~Enhet_id + Parameter_id, dat_export)
tab <- xtabs(~addNA(Unit) + Parameter_id, dat_export)
# tab
tab2 <- apply(tab > 1, 2, sum)

if (sum(tab2>1) > 0){
  cat("Some parameters have > 1 unit: \n")
  print(tab2[tab2 > 1])
}


cat("\n")
cat("Number of parameters per unit: \n")
tab3 <- apply(tab > 1, 1, sum)
tab3


```


### All metals in cod sample station  
```{r}
# All parameters in cod in Kristiansand harbour, Vannmiljø file
cat("All metals in cod in Kristiansand harbour, Vannmiljø file \n")
dat_export %>%
  filter(Navn %in% c("Kristiansand harbour")) %>%
  filter(nchar(Parameter_id) == 2 & Parameter_id != "TS") %>%    # TS = % dry weight
  xtabs(~TISSUE_NAME + Parameter_id, .)

# All parameters in cod in Kristiansand harbour, actual
cat("\n")
cat("All metals in cod in Kristiansand harbour, actual data \n")
dat %>%
  left_join(lookup_stations,  by = "STATION_CODE") %>%
  filter(STATION_NAME %in% c("Kristiansand harbour")) %>%
  filter(nchar(PARAM) == 2) %>%
  xtabs(~TISSUE_NAME + PARAM, .)

# All parameters in cod in Kristiansand harbour, Vannmiljø file
cat("\n")
cat("All metals in cod in Kristiansand harbour, Vannmiljø file in 2018 \n")
dat_export_2018 %>%
  filter(Navn %in% c("Kristiansand harbour")) %>%
  filter(nchar(Parameter_id) == 2 & Parameter_id != "TS") %>%   # TS = % dry weight
  xtabs(~TISSUE_NAME + Parameter_id, .) 

if (FALSE){
  dat_export %>%
  filter(Navn %in% c("Kristiansand harbour")) %>%
  filter(nchar(Parameter_id) == 2 & Parameter_id != "TS") %>%    # TS = % dry weight
  xtabs(~Enhet_id + Parameter_id, .)
  }

```


### All PCBs   
```{r}
# All parameters in cod in Kristiansand harbour, Vannmiljø file
cat("All metals in cod in Kristiansand harbour, Vannmiljø file \n")
dat_export %>%
  filter(Navn %in% c("Kristiansand harbour")) %>%
  filter(grepl("CB", Parameter_id)) %>%   
  xtabs(~TISSUE_NAME + Parameter_id, .)

# All parameters in cod in Kristiansand harbour, actual
cat("\n")
cat("All metals in cod in Kristiansand harbour, actual data \n")
dat %>%
  left_join(lookup_stations,  by = "STATION_CODE") %>%
  filter(STATION_NAME %in% c("Kristiansand harbour")) %>%
  filter(grepl("CB", PARAM)) %>%   
  xtabs(~TISSUE_NAME + PARAM, .)

# All parameters in cod in Kristiansand harbour, Vannmiljø file
cat("\n")
cat("All metals in cod in Kristiansand harbour, Vannmiljø file in 2018 \n")
dat_export_2018 %>%
  filter(Navn %in% c("Kristiansand harbour")) %>%
  filter(grepl("CB", Parameter_id)) %>%   
  xtabs(~TISSUE_NAME + Parameter_id, .) 

```
### All BDEs   
```{r}
# All parameters in cod in Kristiansand harbour, Vannmiljø file
cat("All metals in cod in Kristiansand harbour, Vannmiljø file \n")
dat_export %>%
  filter(Navn %in% c("Kristiansand harbour")) %>%
  filter(grepl("BDE", Parameter_id)) %>%   
  xtabs(~TISSUE_NAME + Parameter_id, .)

# All parameters in cod in Kristiansand harbour, actual
cat("\n")
cat("All metals in cod in Kristiansand harbour, actual data \n")
dat %>%
  left_join(lookup_stations,  by = "STATION_CODE") %>%
  filter(STATION_NAME %in% c("Kristiansand harbour")) %>%
  filter(grepl("BDE", PARAM)) %>%   
  xtabs(~TISSUE_NAME + PARAM, .)

# All parameters in cod in Kristiansand harbour, Vannmiljø file
cat("\n")
cat("All metals in cod in Kristiansand harbour, Vannmiljø file in 2018 \n")
dat_export_2018 %>%
  filter(Navn %in% c("Kristiansand harbour")) %>%
  filter(grepl("BDE", Parameter_id)) %>%   
  xtabs(~TISSUE_NAME + Parameter_id, .) 

```
### All HBCDs   
```{r}
# All parameters in cod in Kristiansand harbour, Vannmiljø file
cat("All metals in cod in Kristiansand harbour, Vannmiljø file \n")
dat_export %>%
  filter(Navn %in% c("Kristiansand harbour")) %>%
  filter(grepl("HBCD", Parameter_id)) %>%   
  xtabs(~TISSUE_NAME + Parameter_id, .)

# All parameters in cod in Kristiansand harbour, actual
cat("\n")
cat("All metals in cod in Kristiansand harbour, actual data \n")
dat %>%
  left_join(lookup_stations,  by = "STATION_CODE") %>%
  filter(STATION_NAME %in% c("Kristiansand harbour")) %>%
  filter(grepl("HBCD", PARAM)) %>%   
  xtabs(~TISSUE_NAME + PARAM, .)

# All parameters in cod in Kristiansand harbour, Vannmiljø file
cat("\n")
cat("All metals in cod in Kristiansand harbour, Vannmiljø file in 2018 \n")
dat_export_2018 %>%
  filter(Navn %in% c("Kristiansand harbour")) %>%
  filter(grepl("HBCD", Parameter_id)) %>%   
  xtabs(~TISSUE_NAME + Parameter_id, .) 

```
### All PFAS   
```{r}
# All parameters in cod in Kristiansand harbour, Vannmiljø file
cat("All metals in cod in Kristiansand harbour, Vannmiljø file \n")
dat_export %>%
  filter(Navn %in% c("Kristiansand harbour")) %>%
  filter(substr(Parameter_id, 1, 2) == "PF") %>%   
  xtabs(~TISSUE_NAME + Parameter_id, .)

# All parameters in cod in Kristiansand harbour, actual
cat("\n")
cat("All metals in cod in Kristiansand harbour, actual data \n")
dat %>%
  left_join(lookup_stations,  by = "STATION_CODE") %>%
  filter(STATION_NAME %in% c("Kristiansand harbour")) %>%
  filter(substr(PARAM, 1, 2) == "PF") %>%   
  xtabs(~TISSUE_NAME + PARAM, .)

# All parameters in cod in Kristiansand harbour, Vannmiljø file
cat("\n")
cat("All metals in cod in Kristiansand harbour, Vannmiljø file in 2018 \n")
dat_export_2018 %>%
  filter(Navn %in% c("Kristiansand harbour")) %>%
  filter(substr(Parameter_id, 1, 2) == "PF") %>%   
  xtabs(~TISSUE_NAME + Parameter_id, .) 

```


### Everything else     
```{r}
# All parameters in cod in Kristiansand harbour, Vannmiljø file
cat("All metals in cod in Kristiansand harbour, Vannmiljø file \n")
dat_export %>%
  filter(Navn %in% c("Kristiansand harbour")) %>%
  filter((!grepl("CB", Parameter_id) & 
           !grepl("BDE", Parameter_id) &
           !grepl("HBCD", Parameter_id) &
           substr(Parameter_id, 1, 2) != "PF" &
           nchar(Parameter_id) != 2) | 
           Parameter_id == "TS") %>%      # need to add "or TS" as TS is dry weight and not a metal
  xtabs(~TISSUE_NAME + Parameter_id, .)

# All parameters in cod in Kristiansand harbour, actual
cat("\n")
cat("All metals in cod in Kristiansand harbour, actual data \n")
dat %>%
  left_join(lookup_stations,  by = "STATION_CODE") %>%
  filter(STATION_NAME %in% c("Kristiansand harbour")) %>%
  filter(!grepl("CB", PARAM) & 
           !grepl("BDE", PARAM) &
           !grepl("HBCD", PARAM) &
           substr(PARAM, 1, 2) != "PF" &
           nchar(PARAM) != 2) %>%   
  xtabs(~TISSUE_NAME + PARAM, .)

# All parameters in cod in Kristiansand harbour, Vannmiljø file
cat("\n")
cat("All metals in cod in Kristiansand harbour, Vannmiljø file in 2018 \n")
dat_export_2018 %>%
  filter(Navn %in% c("Kristiansand harbour")) %>%
  filter((!grepl("CB", Parameter_id) & 
           !grepl("BDE", Parameter_id) &
           !grepl("HBCD", Parameter_id) &
           substr(Parameter_id, 1, 2) != "PF" &
           nchar(Parameter_id) != 2) | 
           Parameter_id == "TS") %>%   
  xtabs(~TISSUE_NAME + Parameter_id, .)

```

## 3. Show some data  

### All metals in liver  
```{r}

# All parameters in cod in Kristiansand harbour, Vannmiljø file
dat_export %>%
  filter(TISSUE_NAME == "Lever") %>%
  filter(nchar(Parameter_id) == 2 & Parameter_id != "TS") %>%    # TS = % dry weight
  mutate(
    Parameter_id = paste0(Parameter_id, " (", Unit, ")"),
    Navn = substr(Navn, 1, 20)
    ) %>%
  ggplot(aes(Navn, Verdi, color = Operator)) +
  geom_point() + 
  facet_wrap(vars(Parameter_id), scales = "free_y") +
  theme(axis.text.x.bottom = element_text(angle = -45, hjust = 0)) +
  labs(title = "Vannmiljø file")



# All parameters in cod in Kristiansand harbour, actual
dat %>%
  filter(TISSUE_NAME == "Lever") %>%
  left_join(lookup_stations,  by = "STATION_CODE") %>%
  filter(nchar(PARAM) == 2 & !PARAM %in% c("D4","D5","D6")) %>%
  mutate(
    PARAM = paste0(PARAM, " (", UNIT, ")"),
    STATION_NAME = substr(STATION_NAME, 1, 20)
    ) %>%
  ggplot(aes(STATION_NAME, VALUE_WW, color = is.na(FLAG1))) +
  geom_point() + 
  facet_wrap(vars(PARAM), scales = "free_y") +
  theme(axis.text.x.bottom = element_text(angle = -45, hjust = 0)) +
  labs(title = "Data file") 


```


## 4. Errors    

### Error 1. Station 26A2 with year 2020 instead of 2019    
NOTE! Check that date actualy are erronous. Cod may be sampled in January-February the year after  
```{r}

sel <- with(dat_export, grepl("2020", Tid_provetak))
xtabs(~ Tid_provetak + Navn, dat_export[sel,])

# Change year 2020 to 2019:
dat_export$Tid_provetak[sel] <- sub("2020", "2019", dat_export$Tid_provetak[sel])


```

#### Check again  

```{r}

sel <- with(dat_export, grepl("2020", Tid_provetak))
xtabs(~ Tid_provetak + Navn, dat_export[sel,])

```


### Error 2. Records with PARAM = CA should be changed to PARAM = CD   
Was correctly labeled only for industry stations I965 and I969. But now this seems to have been corrected    

```{r}
dat_export %>%
  filter(Parameter_id %in% c("CD", "CA")) %>%
  xtabs(~TISSUE_NAME + Parameter_id, .)

```


### Error 3. Station 80B: One AG lacking (SAMPLE_NO2 = 3), values shoud be <0.05 MG_P_KG    
NOT DONE NOW. SHOULD BE ADDED TO Labware   
```{r}

lookup_stations %>%
  filter(STATION_CODE == "80B") %>%
  pull(STATION_NAME)

# Still lacking  
dat_export %>%
  filter(Navn %in% c("Munkholmen") & nchar(Parameter_id) == 2) %>%
  xtabs(~TISSUE_NAME + Parameter_id, .) 

dat_export %>%
  filter(Navn %in% c("Munkholmen") & nchar(Parameter_id) == 2) %>%
  xtabs(~TISSUE_NAME + Provenr, .) 

# we must add one line with Provenr = 244015 and and AG measurement  


# We must also add a new ID_lokal. As shown below they are unique  
tab <- table(dat_export$ID_lokal)
table(tab)
i <- dat_export$ID_lokal %>% stringr::str_extract("[0-9]+") %>% as.numeric() %>%range()
range(i)
diff(range(i))

# Pick another AG measurement and change it?
data_to_add <- dat_export %>%
  filter(Navn %in% c("Munkholmen") & Parameter_id == "AG")
data_to_add <- data_to_add[1,]
data_to_add


```

### Error 4. st. 98B1: Zn in liver is 1000x to high    
For 2 samples (12000 and 11000 instead of 12 and 11 MG_P_KG)   
```{r}

lookup_stations %>%
  filter(STATION_CODE == "98B1") %>%
  pull(STATION_NAME)

# Check
dat_export %>%
  filter(Navn %in% c("Lofoten, Skrova") & Parameter_id == "ZN") %>%
  select(Provenr, Verdi) %>%
  arrange(desc(Verdi))


# MAke the correction
sel <- with(dat_export,
            Navn %in% c("Lofoten, Skrova") & Parameter_id == "ZN" & Verdi > 1000)
sum(sel)
dat_export$Verdi[sel] <- dat_export$Verdi[sel]/1000


```

#### Check again  
```{r}

dat_export %>%
  filter(Navn %in% c("Lofoten, Skrova") & Parameter_id == "ZN") %>%
  select(Provenr, Verdi) %>%
  arrange(desc(Verdi))

```

## 5. Industry stations   
NIVA@47181    Bjørnebærviken
NIVA@47184    Moholmen
NIVA@47186    Toraneskaien


### Check  
```{r}

check <- dat_export %>%
  filter(Vannlok_kode %in% c("NIVA@47181", "NIVA@47184", "NIVA@47186"))

nrow(check)

```

### Remove from both data sheets
```{r}

sel <-  dat_export$Vannlok_kode %in% c("NIVA@47181", "NIVA@47184", "NIVA@47186")
dat_export <- dat_export[!sel,]
cat(sum(sel), "stations were removed from 'dat_export' \n")

sel <-  dat_stations_original$Kilde_id %in% c("NIVA@47181", "NIVA@47184", "NIVA@47186")
dat_stations_original <- dat_stations_original[!sel,]
cat(sum(sel), "stations were removed from 'dat_stations_original' \n")



```


## 6. Write fixed file to excel  
```{r}

data_to_write_1 <- dat_stations_original
data_to_write_2 <- dat_export[names(dat_export_original)]

data_to_write <- list(
  Vannlokaliteter = data_to_write_1,
  Vannregistreringer = data_to_write_2
)

writexl::write_xlsx(data_to_write,fn_write)

```


## 5. Check resulting data    
Called 'versjon 2'   
All 'n' should be 1 - and they are  
```{r}

dat_export <- read_excel(
  "Eksport Vannmiljø/2015_data/NIVA Milkys 2015-data MCCP unntatt Bergen havn 67807 versjon 2.xlsx",
  sheet = "Vannregistreringer")

dat_export %>%
  count(Vannlok_kode, Medium_id, LatinskNavn_id, Provenr, Tid_provetak, Parameter_id) %>%
  xtabs(~n, .)

```

