---
title: "34_Add_uncertainty"
author: "DHJ"
date: "29 8 2019"
output: 
  html_document:
    keep_md: true
---

**UNCRT can be given in three ways:**
- as standard deviation `u` (set METCU = SD)   
- as "expanded uncertainty" = 2 x standard deviation = 2u (set METCU = U2)   
- as percentage = 100u/VALUE (set METCU = %)   
NOTE: The formula used in part 6 and 7, sqrt(s^2 + (Value^2)*(v^2)), is taken from Rob Fryer and is given both in Annex 5 of the 2018/19 MIME report as well as in ICES DOME 'Frequently Asked Questions'    

- This year (part 8), we use percentage (and METCU = %) as given by the labs.   




## 1. Libraries
```{r, results=}

library(dplyr)
library(purrr)
library(readxl)
library(safejoin)  # for safe_left_join() with check = "V" 
                   # (check if no set of values of join columns is duplicated in y)
library(ggplot2)

library(here)
# setwd("C:/Data/seksjon 212")
# set_here("C:/Data/seksjon 212")

# here()
# ?set_here
# set_here("C:/Data/seksjon 212/Milkys")

```


### Selected year  
```{r}

selected_year <- 2020

```


## 2. Chemical data    
We only keep data from the last year  
```{r}

# dir("Data")
# Data at sample level (chemical data, biological efect parameters and VDSI)
# Also in "../Milkys2/Data/101_data_updated_2020-08-05.rds"
 
fn <- "Files_from_Jupyterhub_2020/Raw_data/101_dat_new_2021-09-01.rds"     # data FROM Milkys2 on Jupyterhub
fn <- "../Milkys2/Data/101_dat_new_2021-12-30.rds"                         # data FROM Milkys2 on PC 

dat_all_allyears <- readRDS(fn)

# Remember: for sum variables, only SAMPLE_NO2 has been set (SAMPLE_NO is NA)
check <- dat_all_allyears %>%
  filter(SAMPLE_NO != SAMPLE_NO2)

if (nrow(check) > 0)
  stop("Check cases where SAMPLE_NO != SAMPLE_NO2")

dat_all_allyears <- dat_all_allyears %>%
  mutate(
    TISSUE_NAME = case_when(
      TISSUE_NAME %in% "Egg homogenate of yolk and albumin" ~ "Egg",
      TRUE ~ TISSUE_NAME)
  )


```


### Check if there are unique values 
```{r}

check <- dat_all_allyears %>%
  ungroup() %>%
  group_by(STATION_CODE, LATIN_NAME, TISSUE_NAME, MYEAR, SAMPLE_NO2, 
           UNIT, PARAM) %>%
  mutate(n = n()) %>%
  filter(n > 1)

cat("The number of duplicate values in 'dat_all_allyears' is: ", nrow(check))  # 0


if (FALSE){
  check %>%
    filter(TISSUE_NAME == "Egg" & SAMPLE_NO == 1)
    
}

```
### Plot raw data   
- In earlier veriosn of the data, some PFDcA values from Sørfjorden are very low  
- the same for some PFUdA data 
- PCBs are also very low for 19B, but this is correct (methods are adjusted for this station)   

```{r}

plot_raw_data_liver_highlight <- function(param, 
                                          highlight = "53B", 
                                          highlight_text = "Inner Sørfjord", 
                                          log = FALSE){
  df <- dat_all_allyears %>% 
    filter(LATIN_NAME == "Gadus morhua" & MYEAR == 2020 & PARAM == param) %>% 
    arrange(VALUE_WW) %>%
    mutate(
      Flag1 = ifelse(is.na(FLAG1), "Over LOQ", "Under LOQ"),
      `Station color` = ifelse(STATION_CODE == highlight, highlight_text, "Other stations")
      )
  gg <- ggplot(df, aes(STATION_CODE, VALUE_WW, color = `Station color`, shape = Flag1)) +
    geom_jitter(width = 0.25, height = 0) +
    labs(title = paste(param, "in cod liver"),
         subtitle = "Dashed line = Lowest 'under LOQ-value'")
  if (sum(df$Flag1 == "Under LOQ") > 0)
    gg <- gg + geom_hline(yintercept = min(df$VALUE_WW[df$Flag1 == "Under LOQ"]), 
                          linetype = "dashed")
  if (log)
    gg <- gg + scale_y_log10()
  print(gg)
}

dat_all_allyears %>% 
    filter(LATIN_NAME == "Gadus morhua" & MYEAR == 2020 & substr(PARAM, 1, 2) == "PF") %>%
  group_by(PARAM) %>%
  summarise(median(VALUE_WW), mean(is.na(FLAG1)))

plot_raw_data_liver_highlight("PFDcA") # strange values  
plot_raw_data_liver_highlight("PFOS")
plot_raw_data_liver_highlight("PFOSA")
plot_raw_data_liver_highlight("PFUdA")
plot_raw_data_liver_highlight("PFNA")

plot_raw_data_liver_highlight("CB180", highlight = "19B", highlight_text = "Svalbard", log = TRUE)
plot_raw_data_liver_highlight("CB101", highlight = "19B", highlight_text = "Svalbard", log = TRUE)

```
### Fix PFDcA and PFUdA in cod liver    
- Set values under 0.5 to "< 0.5"
```{r}

sel <- with(dat_all_allyears, 
            LATIN_NAME == "Gadus morhua" & PARAM == "PFDcA" & MYEAR == 2020 & VALUE_WW < 0.5)
message("Change ", sum(sel), " PFDcA records")

dat_all_allyears$VALUE_WW[sel] <- 0.5 
dat_all_allyears$FLAG1[sel] <- "<" 
dat_all_allyears$VALUE_DW[sel] <- NA 
dat_all_allyears$VALUE_FB[sel] <- NA 

# PLot again
# plot_raw_data_liver_highlight("PFDcA") # strange values  

```
### Fix concentrations below zero  
```{r}

if (FALSE){
  
  dat_all_allyears %>%
    filter() %>%
    xtabs(~LATIN_NAME + PARAM + MYEAR, .)
  
}

sel <- with(
  dat_all_allyears, 
  VALUE_WW < 0 & !PARAM %in% c("Delta13C") & LATIN_NAME == "Somateria mollissima")
message("Fix ", sum(sel), " eider duck PCBs with conc. < 0, these are below LOQ")

dat_all_allyears$VALUE_WW[sel] <- -dat_all_allyears$VALUE_WW[sel]
dat_all_allyears$VALUE_DW[sel] <- -dat_all_allyears$VALUE_DW[sel]
dat_all_allyears$VALUE_FB[sel] <- -dat_all_allyears$VALUE_WW[sel]
dat_all_allyears$FLAG1[sel] <- "<"

```


### Create `dat_all`  
- Data for the selected year   

```{r}

dat_all <- dat_all_allyears %>%
  filter(MYEAR %in% selected_year)

xtabs(~ addNA(LATIN_NAME), dat_all)

```


### Fix QUANTIFICATION_LIMIT   
- Given for metals    

```{r}

sel <- !is.na(dat_all$QUANTIFICATION_LIMIT)
cat("Has QUANTIFICATION_LIMIT: \n")
table(sel)

cat("\nHas VALUE_WW > QUANTIFICATION_LIMIT: \n")
xtabs(~(VALUE_WW > QUANTIFICATION_LIMIT), dat_all[sel,])

sel2 <- 
  !is.na(dat_all$QUANTIFICATION_LIMIT) & 
  !is.na(dat_all$VALUE_WW) & 
  with(dat_all, VALUE_WW < QUANTIFICATION_LIMIT)
cat("\nHas VALUE_WW < QUANTIFICATION_LIMIT: \n")
sum(sel2)

# Check these data
# dat_all[sel2,]

# Quantification limits of metals is 1-20; this must be ug/kg, not mg/kg
dat_all[sel2,] %>%
  select(PARAM, VALUE_WW, QUANTIFICATION_LIMIT) %>%
  group_by(PARAM) %>%
  summarise(across(.fns = list(mean=mean, min=min, sd=sd)))

# If not yet fixed (all QUANTIFICATION_LIMIT are above 0.5)
if (min(dat_all$QUANTIFICATION_LIMIT[sel2]) >= 0.5){   
  # fix by dividing values by 1000
  dat_all$QUANTIFICATION_LIMIT[sel2] <- dat_all$QUANTIFICATION_LIMIT[sel2]/1000
}
# after fixing, the lowest QUANTIFICATION_LIMIT is 0.001 (so this fixing can only be done once)

```

### Extra: View the most important data columns     
- For interactive use  
```{r, echo=FALSE}
#
# SHOW RAW DATA
# data set with selected columns, just for View
#
dat_all_allyears_view <- dat_all_allyears %>% 
  select(
    STATION_CODE, STATION_NAME, SAMPLE_DATE, LATIN_NAME, TISSUE_NAME, PARAM, MYEAR, 
    SAMPLE_NO2, VALUE_WW, FLAG1, QUANTIFICATION_LIMIT, UNCERTAINTY
  )

if (FALSE){
  View(dat_all_allyears_view %>% 
         filter(LATIN_NAME == "Gadus morhua" & PARAM == "CB180" & MYEAR == 2020) %>% 
         arrange(VALUE_WW), title = "CB180")
  View(dat_all_allyears_view %>% 
         filter(LATIN_NAME == "Gadus morhua" & PARAM == "PFUdA" & MYEAR == 2020) %>% 
         arrange(VALUE_WW), title = "PFUdA")
  View(dat_all_allyears_view %>% 
         filter(LATIN_NAME == "Gadus morhua" & PARAM == "PFDcA" & MYEAR == 2020) %>% 
         arrange(VALUE_WW), title = "PFDcA")
  View(dat_all_allyears_view %>% 
         filter(LATIN_NAME == "Gadus morhua" & PARAM == "PFOS" & MYEAR == 2020) %>% 
         arrange(VALUE_WW), title = "PFOS")
}

```

## 3a. Uncertainty data   
### Eurofins   
- NOTE: Uncertainty given in data sheet is 2*SE ("expanded uncertainty")   
- `Uncertainty_expand` is `Uncert_EF` if it has data, otherwise it is `Uncert_DHJ`
```{r}

# Uncertainty data  
fn <- "Input_files_2020/Uncertainty/2018-2019_MILKYS_QAdata_Eurofins_for2019data.xlsx"

df_uncert1 <- read_excel(fn, sheet = "QA-data", skip = 5) %>%
  filter(!is.na(`Parameter code`)) %>%
  rename(Uncert_EF = `Estimation of uncertainty of measurements (if available)`) %>%
  mutate(Uncert_EF = as.numeric(Uncert_EF))
# colnames(df_uncert1)

# Sheet for translating to standard name
df_uncert1_NAMES <- read_excel(fn, sheet = "Names")

# Add standard name
df_uncert1 <- df_uncert1 %>%
  safe_left_join(df_uncert1_NAMES, na_matches = "never", check = "V") %>%
  mutate(Uncertainty_expand = 
           case_when(!is.na(Uncert_EF) ~ Uncert_EF,
                     is.na(Uncert_EF) ~ Uncert_DHJ)
         ) %>%
  select(PARAM, Uncertainty_expand, Uncert_EF, Uncert_DHJ)
df_uncert1

```

## . Check and fix uniqueness   
BPA occurs 3 times - we use the mean 
```{r}

# Check 
check <- df_uncert1 %>%
  filter(!is.na(PARAM)) %>%
  group_by(PARAM) %>%
  mutate(n = n()) 

check %>% filter(n > 1)

df_uncert1 <- df_uncert1 %>%
  filter(!is.na(PARAM)) %>%
  group_by(PARAM) %>%
  summarise_all(list(mean) )

# df_uncert1
```

### NILU
```{r}
# This part is used only one time, in order to make the start of an Excel sheet 
#   that is filled out manually

fn <- "Input_files_2020/Uncertainty/NILU_draft.csv"

make_manual_excel_sheet <- FALSE
if (make_manual_excel_sheet){
  pars <- dat_all %>%
    filter(LATIN_NAME %in% "Somateria mollissima") %>%
    count(PARAM) %>%
    pull(PARAM)
  write.csv2(tibble(PARAM = pars), fn)
}
#  end of 'make_manual_excel_sheet' part

fn <- "Input_files_2020/Uncertainty/NILU uncertainty.xlsx"
df_uncert2 <- read_excel(fn, sheet = "For reading")
df_uncert2

```

### Extra values
```{r}
# dat_all %>% filter(PARAM %in% "PFBS" & MYEAR == "2019")

df_uncert_extra <- read.csv(textConnection("
PARAM,Uncertainty_expand,Lab
BDE126,30,Eurofins
BDE183,30,Eurofins
BDE6S,30,Eurofins
BDESS,30,Eurofins
CB_S7,30,Eurofins
DBA3A,30,Eurofins
PFDcA,35,Eurofins
PFDA,35,Eurofins
PFHpA,35,Eurofins
PFHxA,35,Eurofins
PFNA,35,Eurofins
PFOA,35,Eurofins
PFOS,35,Eurofins
PFOSA,35,Eurofins
PFUdA,35,Eurofins
PFUnda,35,Eurofins
P_S,35,Eurofins
PFAS,35,Eurofins
HBCDD,30,Eurofins
PAH16,30,Eurofins
KPAH,30,Eurofins
TDEPP,20,Eurofins
PFBS,35,Eurofins
"), stringsAsFactors = FALSE)


```


### Combine
```{r}

param_pfas <- c("PFBS", "PFDcA", "PFHpA", "PFHxA", "PFHxS", "PFNA", "PFOA", "PFOS", "PFOSA", "PFUdA")

df_uncert <- bind_rows(
  df_uncert1 %>% select(PARAM, Uncertainty_expand) %>% 
    mutate(Lab = "Eurofins"),
  df_uncert2 %>% select(PARAM, Uncertainty_expand) %>% 
    mutate(Lab =case_when(
      PARAM %in% param_pfas ~ "NIVA",
      TRUE ~ "NILU")
      )
) %>%
  filter(!is.na(PARAM)) %>%
  mutate(MYEAR = selected_year)
# df_uncert %>%
#   arrange(PARAM)

```

```{r}
df_uncert1 %>% filter(PARAM == "BDESS")
df_uncert2 %>% filter(PARAM == "BDESS")
df_uncert %>% filter(PARAM == "BDESS")
df_uncert %>% filter(PARAM == "BDESS")
```



## 3b. Add 'Uncertainty_expand' and 'Lab' to data   
- Creates `dat_all_2` from `dat_all` 
```{r}

param_bioleffects <- c("PA1OH", "PA1O", "1-OH-fenantren", "PYR1OH",  "PYR1O", "1-OH-pyren", 
      "BAP3OH", "BAP3O", "3-OH-benzo[a]pyren", 
      "EROD", "ALAD", 
      "VDSI", "Intersex", "INTF%")     # INTF% = percentage intersex, see vocab.ices.dk

dat_all_2 <- dat_all %>%
  mutate(Lab =
           case_when(LATIN_NAME %in% "Somateria mollissima" ~ "NILU",
                     PARAM %in% c("D4", "D5", "D6") ~ "NILU",
                     PARAM %in% c(param_pfas, param_bioleffects) ~ "NIVA",
                     TRUE ~ "Eurofins")
  ) %>% # filter(PARAM == "ALAD") %>% select(PARAM, Lab) %>% View()
  left_join(df_uncert,
                 na_matches = "never",
            by = c("MYEAR", "Lab", "PARAM")
            )


if (nrow(dat_all) != nrow(dat_all_2)){
  stop("Duplicated rows in 'PARAM'!")

  df_uncert %>%
    add_count(MYEAR, Lab, PARAM) %>%
    filter(n > 1)  
  
    }

```


### Values lacking uncertainty
```{r}

xtabs(~is.na(Uncertainty_expand), dat_all_2 %>% filter(MYEAR == selected_year))

if (FALSE){
  xtabs(~PARAM + Lab, dat_all_2 %>% filter(MYEAR == selected_year & is.na(Uncertainty_expand)))
  xtabs(~PARAM + Lab, dat_all_2 %>% filter(MYEAR == selected_year & !is.na(Uncertainty_expand)))
}

```

### Check example  
```{r}

if (FALSE) {
  
  dat_all_2 %>%
    filter(PARAM %in% "BDE126" & MYEAR == selected_year) %>%
    select(TISSUE_NAME, PARAM, Lab, VALUE_WW, Uncertainty_expand)

}

```


## 4. LOQ values     

### `tab_loq1` created  
- Compare QUANTIFICATION_LIMIT and LOQ from less-than   
    - `LOQ_lt_min` (minimum value of less-than's) will be used where QUANTIFICATION_LIMIT is lacking  
    - This is a change since 2019 (and before), when we used LOQ_lt_med (median less-than value). This lead to very high errors in some parameters with large variation in less-than values, such as CB180 in cod. See below!
    - In 2019: QUANTIFICATION_LIMIT given by Eurofins, not given in 2020 except for metals  
```{r, fig.width=5, fig.height=10, warning=FALSE}

tab_loq1 <- dat_all_2 %>%
  filter(MYEAR == selected_year & FLAG1 %in% "<") %>%
  group_by(MYEAR, Lab, LATIN_NAME, TISSUE_NAME, PARAM) %>%
  summarise(LOQ_gi_min = min(QUANTIFICATION_LIMIT, na.rm = TRUE),
            LOQ_given = median(QUANTIFICATION_LIMIT, na.rm = TRUE),
            LOQ_lt_min = min(VALUE_WW, na.rm = TRUE) %>% round(4),
            LOQ_lt_med = median(VALUE_WW, na.rm = TRUE) %>% round(4),
            LOQ_lt_max = max(VALUE_WW, na.rm = TRUE) %>% round(4),
            .groups = "drop"
            ) %>%
  ungroup()

# Show it (try to make it fit in the width)
df <- tab_loq1 %>%
  filter(Lab %in% "Eurofins") %>%
  select(-MYEAR, -Lab) %>%
  mutate(LATIN_NAME = substr(LATIN_NAME, 1, 10),
         TISSUE_NAME = substr(TISSUE_NAME, 1, 10),
         PARAM = substr(PARAM, 1, 10))

df

# Parameters/tissues with given QUANTIFICATION_LIMIT and at least one less-than value
# None in 2020  

# If any data (with less-thans) are given with QUANTIFICATION_LIMIT, make plot:
if (sum(!is.na(df$LOQ_given)) > 0){
  
  df %>%
    mutate(LOQ_lt_min = LOQ_lt_min/LOQ_given,
           LOQ_lt_med = LOQ_lt_med/LOQ_given,
           LOQ_lt_max = LOQ_lt_max/LOQ_given) %>%
    pivot_longer(LOQ_gi_min:LOQ_lt_max, names_to = "Type", values_to = "LOQ") %>%
    ggplot(aes(x = Type, y = paste(PARAM, TISSUE_NAME), fill = LOQ)) +
    geom_tile()
  
}


```

### `tab_loq2`, overall minimum value   
- this is created for all records   
- to use only for those records where there are neither QUANTIFICATION_LIMIT nor less-thans (part 5)    
- We make and exception for CB180 in cod, where we will use the lowest observed value for 19B, or the lowest observed value for all the rest   
```{r}

tab_loq2 <- dat_all_2 %>%
  filter(MYEAR %in% selected_year & !is.na(VALUE_WW)) %>%
  group_by(MYEAR, Lab, LATIN_NAME, TISSUE_NAME, PARAM) %>%
  summarise(Min_value = min(VALUE_WW) %>% round(4),
            .groups = "drop")

#
# Special case!
# For CB180 in cod liver, we use a different minimum value for 19B (Svalbard) and the rest, see graph in section 2
# Will be used in section 5
#
tab_loq_CB180_cod <- dat_all_2 %>%
  filter(MYEAR %in% selected_year & !is.na(VALUE_WW) & LATIN_NAME == "Gadus morhua" & PARAM == "CB180") %>%
  mutate(
    Station19B = (STATION_CODE %in% "19B")
    ) %>%
  group_by(MYEAR, Lab, LATIN_NAME, TISSUE_NAME, PARAM, Station19B) %>%
  summarise(Min_value = min(VALUE_WW) %>% round(4),
            .groups = "drop")



```

### `tab_loq` creation  
- from tab_loq1 + tab_loq2    
```{r}

tab_loq <- safejoin::safe_full_join(
  tab_loq1, tab_loq2, 
  na_matches = "never", 
  check = "uv") %>%
  mutate(
    Ratio_min_LOQ = Min_value/LOQ_lt_min
    )


```

### Minimum under min LOQ, table  
- Overall minimum value less than lowest less-than?   
- i.e., Ratio_min_LOQ significantly less than 1  
- 3 cases: PFDcA, CB180 and PFUdA   
- Lowest PFDcA concentration is zero!  
```{r}

cat("Ratio_min_LOQ less than 1: ")
cat(
  mean(tab_loq$Ratio_min_LOQ < 1, na.rm = TRUE)*100, "%, ",
  sum(tab_loq$Ratio_min_LOQ < 1, na.rm = TRUE), "cases \n\n"  )

cat("Ratio_min_LOQ significantly less than 1: ")
cat(
  mean(tab_loq$Ratio_min_LOQ < 0.95, na.rm = TRUE)*100, "%, ",
  sum(tab_loq$Ratio_min_LOQ < 0.95, na.rm = TRUE), "cases \n\n"  )

tab_loq %>% 
  filter(Ratio_min_LOQ < 1) %>%
  arrange(Ratio_min_LOQ)
        
```

### Minimum under min LOQ, plot    
```{r}

ggplot(tab_loq, aes(LOQ_lt_min, Min_value)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_point(data = subset(tab_loq, Ratio_min_LOQ < 1), color = "red") +
  geom_text(data = subset(tab_loq, Ratio_min_LOQ < 1), 
            aes(label = paste(PARAM, TISSUE_NAME)), color = "red", hjust = -0.1, size = rel(3))
  

```


## 5. Add LOQ where QUANTIFICATION_LIMIT not given  

### Create `dat_all_updated`   
- Creates `dat_all_updated` by adding `tab_loq` columns to `dat_all_2`   
- Creates new values of `QUANTIFICATION_LIMIT`  
If no less-thans exist, we use half of the minimum value!
Creating new data set 'dat_all_updated'  
```{r}

dat_all_updated <- dat_all_2 %>%
  # Add LOQ_lt_med variable
  safe_left_join(tab_loq %>% select(MYEAR, Lab, LATIN_NAME, TISSUE_NAME, PARAM, 
                                    LOQ_lt_min, LOQ_lt_med, Min_value),
                 na_matches = "never", 
                 check = "v") %>%
  mutate(LOQ_source = case_when(
    !is.na(QUANTIFICATION_LIMIT) ~ "Given",
    !is.na(LOQ_lt_min) ~ "From LT",
    is.na(LOQ_lt_min) ~ "From minvalue"),
  # Where we lack QUANTIFICATION_LIMIT, we set it to LOQ_lt_med variable
         QUANTIFICATION_LIMIT = 
           case_when(LOQ_source %in% "Given" ~ QUANTIFICATION_LIMIT,
                     LOQ_source %in% "From LT" ~ LOQ_lt_min,          # if QUANTIFICATION_LIMIT not given 
                     LOQ_source %in% "From minvalue" ~ Min_value/2)   # if also no less-thans exist
  )
# We keep Lab, LOQ_lt_min, LOQ_lt_med and LOQ_given for later!

# dat_all_updated 

xtabs(~is.na(QUANTIFICATION_LIMIT) + LOQ_source, dat_all_updated %>% filter(MYEAR == selected_year))

```


### Make exeption for CB180 values   
```{r}

#
# 'Min_value' has been set from the very low 19B values
# For stations except 19B, we set it higher
#
sel <- with(dat_all_updated, 
            LATIN_NAME == "Gadus morhua" & PARAM == "CB180" & MYEAR == 2020 & !STATION_CODE %in% "19B")
message("Min_value changed for ", sum(sel), " records")
dat_all_updated$Min_value[sel] <- subset(tab_loq_CB180_cod, !Station19B)$Min_value

#
# For all CB180 in liver, we use Min_value although LOQ exists
#
sel <- with(dat_all_updated, 
            LATIN_NAME == "Gadus morhua" & PARAM == "CB180" & MYEAR == 2020)
message("QUANTIFICATION_LIMIT and LOQ_source changed for ", sum(sel), " records")
dat_all_updated$QUANTIFICATION_LIMIT[sel] <- dat_all_updated$Min_value[sel]
dat_all_updated$LOQ_source[sel] <- "From minvalue"


#
# For check
#
if (FALSE){
  
  dat_all_updated_view <- dat_all_updated %>% 
    select(
      STATION_CODE, TISSUE_NAME, PARAM, MYEAR, 
      SAMPLE_NO2, VALUE_WW, FLAG1, QUANTIFICATION_LIMIT, UNCERTAINTY,
      QUANTIFICATION_LIMIT, LOQ_source
    )
  
  dat_all_updated_view %>% 
    filter(TISSUE_NAME == "Lever" & PARAM == "CB180") %>%
    View()

}

```




#### Checks (not necessary)
```{r}
# xxx
dat_all_updated %>% 
  filter(MYEAR == selected_year & is.na(FLAG1)) %>%
  xtabs(~PARAM + is.na(QUANTIFICATION_LIMIT), .)

dat_all_updated %>% 
  filter(MYEAR == selected_year & is.na(FLAG1) & PARAM == "BDE47") %>%
  xtabs(~Lab + is.na(QUANTIFICATION_LIMIT), .)

dat_all_updated %>% 
  filter(MYEAR == selected_year & is.na(FLAG1) & PARAM == "CB118") %>%
  xtabs(~Lab + is.na(QUANTIFICATION_LIMIT), .)

dat_all_updated %>% 
  filter(MYEAR == selected_year & is.na(FLAG1) & PARAM == "CB118" & Lab %in% "NILU") 
  
tab_loq2 %>%
    filter(MYEAR == selected_year & PARAM == "CB118" & Lab %in% "NILU")
tab_loq %>%
    filter(MYEAR == selected_year & PARAM == "CB118" & Lab %in% "NILU")
```



## 6. Make UNCRT - example     
- UNCRT can be given in three ways:
    - as standard deviation `u` (set METCU = SD)   
    - as "expanded uncertainty" = 2 x standard deviation = 2u (set METCU = U2)   
    - as percentage = 100u/VALUE (set METCU = %)   
NOTE: The formula sqrt(s^2 + (Value^2)*(v^2)) is taken from Rob Fryer and is given both in Annex 5 of the 2018/19 MIME report as well as in ICES DOME 'Frequently Asked Questions'  
  
In 2016 and 2017, we converted from percent to standard deviation (METCU = SD), using an assumed value of LOD:   
Based on Annex 5 2018 ('Annex05_reporting_less-thans_and_uncertainties_v3.docx')   
  - u = standard deviation   
  - u = sqrt(s^2 + (Value^2)*(v^2)) where
      - v = relative uncertainty (percentage divided by 100)
      - s = SD of replicate measurements of the blank
      - s is assumed to be LOD/3
      - LOD may be assumed to be LOQ/x
      
### Example   
- assuming (quite conservatively) that LOD = LOQ/1.33, or (even more consrvatively) that LOD = LOQ  
```{r,fig.width=8, fig.height=4}

v <- 0.3          # relative uncertainty (percentage divided by 100)
LOQ <- 0.2

par(mfrow = c(2,2), mar = c(4,4,2,1))

x <- 1.333      # assume (quite conservatively) LOD = LOQ/1.33 s = LOD/3 = (LOQ/1.33)/3 = 4
LOD <- LOQ/x
s <- LOD/3      # SD of replicate measurements of the blank - assumed to be LOD/3
Value <- seq(0.2, 4, 0.1)
u <- sqrt(s^2 + (Value^2)*(v^2))
plot(Value, u, main = paste("x =", x, "(LOD = LOQ/x)"))
plot(Value, 100*u/Value, ylab = "Percent", main = paste("x =", x))
abline(v = LOQ, lty = "dashed")

x <- 1.000        # assume (very conservatively) LOD = LOQ
LOD <- LOQ/x
s <- LOD/3      # SD of replicate measurements of the blank - assumed to be LOD/3
Value <- seq(0.2, 4, 0.1)
u <- sqrt(s^2 + (Value^2)*(v^2))
plot(Value, u, main = paste("x =", x, "(LOD = LOQ/x)"))
plot(Value, 100*u/Value, ylab = "Percent", main = paste("x =", x))
abline(v = LOQ, lty = "dashed")

```

## 7. Using (as test) median less-than value as LOQ  
- I.e. LOQ_lt_med   
- Creates 'dat_all_updated_test'  
- Based on the example in part 6  
- We ended up using *not* this approach, but just the given percentages, see part 8

### Add LOD, UNCRT and METCU  
We here use LOQ from less-thans as LOQ, if possible  
    - i.e. LOQ_lt_med, or QUANTIFICATION_LIMIT if LOQ_lt_med is not given  
    - we here assume that LOD = LOQ (so we use 'LOQ_lt_med/3')
```{r}

# dat_all_updated_test <- dat_all_updated %>%
#   mutate(
#     LOD = case_when(
#       !is.na(LOQ_lt_med) ~ LOQ_lt_med,
#       is.na(LOQ_lt_med) ~ QUANTIFICATION_LIMIT),
#     s = LOD/3,
#     UNCRT = case_when(
#       !is.na(s) ~ sqrt(s^2 + (VALUE_WW^2)*((Uncertainty/100)^2)),
#       is.na(s) ~ Uncertainty),   # should be zero cases of this
#     METCU = "SD"
#   )

```

### Check 1 - table       
- Summary table  
```{r}

# mean_uncrt <- dat_all_updated_test %>%
#   filter(MYEAR == selected_year & !is.na(UNCRT)) %>%
#   group_by(MYEAR, Lab, LATIN_NAME, TISSUE_NAME, PARAM) %>%
#   summarise(
#     Uncertainty = mean(Uncertainty, na.rm = TRUE),
#     QUANTIFICATION_LIMIT = mean(QUANTIFICATION_LIMIT, na.rm = TRUE),
#     LOQ_source = paste(unique(LOQ_source), collapse = ","),
#     VALUE_mean = mean(VALUE_WW, na.rm = TRUE) %>% round(3),
#     VALUE_min = min(VALUE_WW, na.rm = TRUE) %>% round(3),
#     UNCRT_mean = mean(UNCRT, na.rm = TRUE) %>% round(3),
#     UNCRT_max = max(UNCRT, na.rm = TRUE) %>% round(3),
#     percerror_mean = mean(100*UNCRT/VALUE_WW, na.rm = TRUE) %>% round(3),
#     percerror_max = max(100*UNCRT/VALUE_WW, na.rm = TRUE) %>% round(3)
#   )
# 
# # View(mean_uncrt)
# 
# # For comparison with minimum less-than value, 
# mean_uncrt_1 <- mean_uncrt

```


### Check 2 - plot    
- Plot one parameter/tissue
```{r}

# plot_error <- function(lab, param, species, tissue, data = dat_all_updated) {
#   
#   df_plot <- data %>%
#     filter(MYEAR == selected_year & !is.na(UNCRT)) %>%
#     filter(Lab == lab & PARAM == param & LATIN_NAME == species & TISSUE_NAME == tissue) %>%
#     mutate(Percent_error = 100*UNCRT/VALUE_WW)
#   
#   gg <- ggplot(df_plot, aes(VALUE_WW, Percent_error)) +
#     geom_point() +
#     geom_vline(xintercept = df_plot$s[1]*3, linetype = "dashed") +
#     xlim(0,50) +
#     labs(title = paste0(param, " (", lab, ") in ", species, ", ", tissue),
#          subtitle = "dotted line = assumed LOD")
#   
#   print(gg)
#   
# }
# 
# plot_error("Eurofins", "CB180", "Gadus morhua", "Lever", dat_all_updated_test)

```

## 8. Just use 'Uncertainty_expand' as given by the labs     
-  METCU = %
```{r}

dat_all_updated_2 <- dat_all_updated %>%
  mutate(
    UNCRT = Uncertainty_expand,   # Just use as is
    METCU = "%"
  )

cat("Number of 'UNCRT' lacking: \n")
xtabs(~is.na(UNCRT), dat_all_updated_2 %>% filter(MYEAR == selected_year))
cat("\n")
cat("Number of 's' lacking: \n")
xtabs(~is.na(s), dat_all_updated_2 %>% filter(MYEAR == selected_year))



```


### Show raw data  
- selected columns 
```{r}
#
# data set with selected columns, just for View
#
dat_all_updated_view <- dat_all_updated_2 %>% 
  select(
    STATION_CODE, STATION_NAME, SAMPLE_DATE, LATIN_NAME, TISSUE_NAME, PARAM, MYEAR, 
    SAMPLE_NO2, VALUE_WW, FLAG1, Lab, QUANTIFICATION_LIMIT,  
    LOQ_lt_min, LOQ_lt_med, Min_value, LOQ_source, UNCRT, METCU
  )

if (FALSE){
  
  View(dat_all_updated_view %>% 
         filter(LATIN_NAME == "Gadus morhua" & PARAM == "CB180") %>% 
         arrange(VALUE_WW), title = "CB180")
  View(dat_all_updated_view %>% 
         filter(LATIN_NAME == "Gadus morhua" & PARAM == "PFUdA") %>% 
         arrange(VALUE_WW), title = "PFUdA")
  View(dat_all_updated_view %>% 
         filter(LATIN_NAME == "Gadus morhua" & PARAM == "PFDcA") %>% 
         arrange(VALUE_WW), title = "PFDcA")
  View(dat_all_updated_view %>% 
         filter(LATIN_NAME == "Gadus morhua" & PARAM == "PFOS") %>% 
         arrange(VALUE_WW), title = "PFOS")
}

```




### Check if there are unique values 
```{r}

check <- dat_all_updated_2 %>%
  ungroup() %>%
  group_by(STATION_CODE, LATIN_NAME, TISSUE_NAME, MYEAR, 
           SAMPLE_NO2, UNIT, PARAM) %>%
  mutate(n = n()) %>%
  filter(n > 1)

if (nrow(check) > 0){
  warning("There are ", nrow(check), " duplicates in the data!")
  # xtabs(~STATION_CODE + MYEAR, check)
  xtabs(~PARAM + MYEAR, check)
} else {
  message("There are no duplicates in the data.")
}


```

## 9. Save  
**This is input for 842_ICES_submission_xxxxdata.Rmd**
```{r}

fn <- paste0("Data/841_dat_all_", selected_year) 
saveRDS(dat_all_updated_2, fn)

message("Data written to ", fn)

#
# This is input for script 842 ICES submission
#

#
# Read back, if needed:
# 2019 data
# dat_all_updated_2 <- readRDS("../Milkys/Data/34_dat_all.rds")
# 2020 data
# dat_all_updated_2 <- readRDS(fn)

```



#### Final check
```{r}

# xxx
dat_all_updated_2 %>% 
  filter(MYEAR == selected_year & is.na(FLAG1)) %>%
  xtabs(~PARAM + is.na(UNCRT), .)

dat_all_updated_2 %>% 
  filter(MYEAR == selected_year & is.na(FLAG1) & PARAM == "BDE47") %>%
  xtabs(~Lab + is.na(UNCRT), .)

dat_all_updated_2 %>% 
  filter(MYEAR == selected_year & is.na(FLAG1) & PARAM == "BDE47" & Lab %in% "NILU")

dat_all_updated_2 %>% 
  filter(MYEAR == selected_year & is.na(FLAG1) & PARAM == "BDE126" & Lab %in% "NILU")

```

#### Check in MIME   
```{r}

df <- dat_all_updated_2 %>% 
  filter(MYEAR == selected_year & is.na(FLAG1) & 
           PARAM == "HG" & STATION_CODE == "30B" &
           TISSUE_NAME == "Muskel")
df %>%
  select(VALUE_WW, Uncertainty_expand, LOQ_lt_min, Min_value, UNCRT, METCU)

```






### Comparison   
- with stuff in 7, but now commented out  
```{r}


# mean_uncrt_compare <- function(lab, param, species, tissue){
#   bind_rows(
#     mean_uncrt_1 %>%
#       filter(Lab == lab & PARAM == param & LATIN_NAME == species, TISSUE_NAME == tissue),
#     mean_uncrt_2 %>%
#       filter(Lab == lab & PARAM == param & LATIN_NAME == species, TISSUE_NAME == tissue)
#   )
# }
# 
# mean_uncrt_compare("Eurofins", "CB180", "Gadus morhua", "Lever")



```

### Check 3   
- Kind of table/plot  
```{r}

param <- "CB118"

df1 <- dat_all_updated_2 %>% 
  filter(MYEAR == selected_year & is.na(FLAG1))

df2 <- dat_all_updated_2 %>% 
  filter(MYEAR == selected_year & is.na(FLAG1) & PARAM == param)

cat("\n************************************************************\n")
cat("Check error/value ratio for 1) all data and 2) CB118\n\n")

df_list = list(df1,df2)
names(df_list) <- c("All data", param)

for (i in 1:2){
  
  df <- df_list[[i]]
  
  with(df, print(quantile(UNCRT/VALUE_WW, na.rm = TRUE)))

  gg <- ggplot(df, aes(y = UNCRT/VALUE_WW)) + 
    geom_boxplot() +
    scale_y_log10() +
    labs(title = names(df_list)[i])
  print(gg)
  
  }

  

# xtabs(~PARAM + Lab, dat_all_updated_2 %>% filter(MYEAR == selected_year & is.na(UNCRT)))
# dat_all_updated_2 %>% filter(PARAM %in% "BDE47" & STATION_CODE == "19N") %>% View()

```




