---
title: "22 Submitting data to ICES"
output: 
  html_document:
    keep_md: true
---

For submission of 2021 data in autumn 2022    
  
Soure of main data (data_ind2):
- prepared in Jupyterhub script 101 (where it also gets input from scr. 101 in this folder)    
- exported from Jupyterhub to Files_from_Jupyterhub_2020 (in this folder) + to the Milkys2 folder  
- used in script 34 (2019 version) where uncertainty (UNCRT) is added  

Some important things to note:  
- There is an error in Labware for 53B - muscle data ar assigne to the wrong individuals  
- "HCHA", "HCHG", and "HCB" has not been submitted to ICES (see 'Table 10 a1') - should be resumbitted for previous years (HCHA and HCHG are lacking from OSPAR's app, for HCB there seems to be data up to 2014 only)  


_NOTE_: METOA should be brought into the 'data_ind2' data set on a much former stages. METOA is in the methods table of NIVAbasen - See '03extra01'.   
_however:_ cannot find it in this base? Seems to be in 'ICES_EXPORT.ANALYTICAL_METHODS_RECORD':
get_nivabase_data("select OWNER,TABLE_NAME from ALL_TAB_COLUMNS where column_name = 'METOA'")   
         OWNER                TABLE_NAME  
 1 ICES_EXPORT ANALYTICAL_METHODS_RECORD  
  
... but cannot access these data:   
get_nivabase_data(str)   
(...times out...)  


## Where files will be saved  
```{r}

save_folder_txt <- "Files_to_ICES/2021/Test"
save_folder_rds <- "Files_to_ICES/2021/Rdata"
save_folder_excel <- "Files_to_ICES/2021/Excel"

```

### Selected year  
```{r}

selected_year <- 2021

```


## 0. Libraries and functions
```{r}

library(tidyverse)
library(lubridate)

# library(safejoin) # https://github.com/moodymudskipper/safejoin
# - results in error when used. Tried to reinstall:
# devtools::install_github("moodymudskipper/safejoin")
# -  HTTP error 401. Bad credentials
# Instead, defined my own safe_left_join (see 'Functions')  

# source("01_Get_chemical_data_NIVAbasen_functions.R")  # for get_standard_parametername()

source("844_ICES_submission_check_functions.R")

```



### Functions  
Between decimal and minute formats. Note the fixed format (2 signs for degrees, 6 signs for minutes) 
```{r}

decimal2minute <- function(x) {
  deg <- floor(x) %>% sprintf("%02.0f", .)
  min <- round(60*(x - floor(x)), 3) %>% sprintf("%02.3f", .)
  paste0("+", deg, " ", min) 
  }

minute2decimal <- function(x) {
  deg <- substr(x,2,3) %>% as.numeric()
  min <- as.numeric(substr(x,5,10) %>% as.numeric())/60
  deg + min
}

# Test
# x <- df_stat2$long_2018faktisk
# x <- subset(df_stat2, stasjon == "19B")$long_2018faktisk
# x <- xsel <- [!is.na(x%in%
# x %>% head()
# x %>% decimal2minute() %>% head()
# x %>% decimal2minute() %>% minute2decimal() %>% head()



#
# left-join that fails if there are several rows in the 'right' table that fits one row in the 'left' table 
#
# This is used instead of safe_left_join in the safejoin package (https://github.com/moodymudskipper/safejoin)
#
safe_left_join <- function(x, ...){
  n1 <- nrow(x)
  result <- left_join(x=x, ...)
  n2 <- nrow(result)
  if (n2 > n1){
    stop("Number of rows in 'left' table increased by ", n2-n1, ". In at least one case, there are several rows in the 'right' table that fits one row in the 'left' table.")
  } else {
    message("Left join performed. Rows in the 'left' table are unchanged.")
  }
  result
}

if (FALSE){
  # Test  
  df1 <- tibble(x = 1:3)
  df2a <- tibble(x = c(1, 1, 2), y = c("first", "second", "third"))
  df2b <- tibble(x = c(1,  2), y = c("first", "third"))
  df1 %>% left_join(df2a)        # does not fail, gives 4 rows   
  df1 %>% safe_left_join(df2a)   # fails
  df1 %>% safe_left_join(df2b)   # does not fail
}




```



## 1. Read data


### a1. Get data produced by script 105 in Jupyterhub  

- used to be based on data produced by script 841 (see scripts from previous years)    

```{r}

folder <- "Files_from_Jupyterhub_2021/Raw_data"
dir(folder, "105_*")

data_all <- readRDS(paste0(folder, "/", "105_data_with_uncertainty_2022-09-23.rds")) %>%
  # Set 36A to 36A1
  mutate(
    STATION_CODE = ifelse(STATION_CODE == "36A", "36A1", STATION_CODE)
  )

    
#
# Data in the selected year
#
data_ind2 <- data_all %>%
  filter(MYEAR %in% selected_year) %>%
  mutate(
    SAMPLE_NO = SAMPLE_NO2,
    PARAM = case_when(
      PARAM %in% "1-OH-fenantren" ~ "PA1OH",
      PARAM %in% "1-OH-pyren" ~ "PYR1OH",
      PARAM %in% "3-OH-benzo[a]pyren" ~ "BAP3OH",
      TRUE ~ PARAM),
    UNCRT = round(UNCRT, 3)
  )

#
# Couple of tables
#

table(data_ind2$TISSUE_NAME)

xtabs(~ PARAM, data_ind2 %>% filter(grepl("OH", PARAM)))

xtabs(~ is.na(SAMPLE_NO), data_ind2)

xtabs(~ addNA(LATIN_NAME), data_ind2)

#
# Data on individual level of fish (lengt, weight, gonad weight)
#
# Must contain these exact column names:
# STATION_CODE, Fish_no, Length, Weight, Gonad_weight
#
# data_ind_fish_old <- readRDS("Data/03_Fish_individual_data.rds")

dir(folder, "150_*")
data_ind_fish <- readRDS(paste0(folder, "/", "150_Fish_individual_data_2021.rds"))

```

#### Check PAH metabolites
```{r}

if (FALSE){
  
  data_ind2 %>%
    filter(PARAM %in% c(
      "PA1OH", "PA1O", "1-OH-fenantren", "PYR1OH",  "PYR1O", "1-OH-pyren", 
      "BAP3OH", "BAP3O", "3-OH-benzo[a]pyren")) %>%
    xtabs(~PARAM, .)

  data_ind2 %>%
    filter(PARAM %in% c("VDSI", "Intersex")) %>%
    xtabs(~PARAM + STATION_CODE, .)
  
}


```

### a2-0. Change Intersex to code `INTF%`  
- https://vocab.ices.dk/?CodeID=53682  
- VDSI is OK   

```{r}

sel <- data_ind2$PARAM %in% "Intersex"
data_ind2$PARAM[sel] <- "INTF%"

cat("Intersex changed to code INTF% for", sum(sel), "records")


```


### a2-1. Get VDSI + intersex data   
- Intersex should have code `INTF%` (https://vocab.ices.dk/?CodeID=53682)   
- NOT NEEDED anymore (see scripts from previous years)   

### a2-2. Add VDSI + intersex data   
- NOT NEEDED anymore (see scripts from previous years)   


### a3. For checking single stations
```{r}

# debugonce(report_station)
data_all %>% report_station("43B2", "Lever", year = 2020, threshold_samples = 7)
# report_station("30B") 


```


### a4. Data in Aquamonitor format, for dates only   
We us these data to create ´df_dates´ containing ´SAMPLE_DATE´    
- NOT NEEDED (see older scripts if needed)   

### a5. Add SAMPLE_DATE to data
- NOT NEEDED (see older scripts if needed)   

### a6. Check 227G2  
- NOT NEEDED (see older scripts if needed)   

### b. Checks  

#### Years
```{r}
table(data_ind2$MYEAR)
```

SAMPLE_NO 1 

- Excluding siloxanes, which have sample numbers >100

```{r}

data_ind2 %>%
  filter(MYEAR %in% selected_year & !PARAM %in% c("D4","D5","D6")) %>%
  group_by(STATION_CODE, TISSUE_NAME) %>% 
  summarise(Sample_no = paste(min(SAMPLE_NO), "-", max(SAMPLE_NO)),
            .groups = "drop") %>%
  spread(TISSUE_NAME, Sample_no)

```
- Special case: siloxanes - sample numbers >100 for liver  
- Note 24B, which has SAMPLE_NO = 7 + all numbers 101-115    

```{r}

data_ind2 %>%
  filter(MYEAR %in% selected_year & PARAM %in% c("D4","D5","D6")) %>% # View()
  group_by(STATION_CODE, TISSUE_NAME) %>% 
  summarise(Sample_no = paste(min(SAMPLE_NO), "-", max(SAMPLE_NO)),
            .groups = "drop") %>%
  spread(TISSUE_NAME, Sample_no)

```


SAMPLE DATE (cod)
```{r}

data_ind2 %>%
  filter(TISSUE_NAME %in% c("Lever", "Muskel")) %>%
  group_by(STATION_CODE) %>% 
  summarise(Sample_date = paste(sort(unique(as.character(SAMPLE_DATE))), collapse = ", "),
            .groups = "drop")

```
  
#### Drop I964  
We drop I964 - Toraneskaien (B4), Mo i Rana (industry station)
```{r}

data_ind2 <- data_ind2 %>% filter(!STATION_CODE %in% "I964")

```

#### Drop new stations, OR NOT
we may or may not drop the following stations (as they are new and not in the station dictionary)
    + 19B (Svalbard)  78.33153483	15.14101617 (mail fra Norman torsdag 23. august 2019 10.15)
    + 28A2 Ålesund havn  
    + 97A3 Bodø havn  
```{r}

DROP_NEW_STATIONS <- FALSE

if (DROP_NEW_STATIONS)
  data_ind2 <- data_ind2 %>% filter(!STATION_CODE %in% c("19B", "28A2", "97A3"))

```

#### Drop NA stations  
```{r}

sel <- is.na(data_ind2$STATION_CODE)
# data_ind2[sel,]

# Delete these 
data_ind2 <- data_ind2[!sel,]

cat("Removed", sum(sel), "recods that had no station code\n")

```

#### Parameter list
```{r}

# xtabs(~PARAM, data_ind2 %>% filter(MYEAR == selected_year))

```

#### Check some perfluor compounds
```{r}
# grep("BD", unique(data_ind2$PARAM), value = TRUE) 
grep("PFH", unique(data_ind2$PARAM), value = TRUE) 

# None of these:
data_ind2 %>%
  filter(PARAM == "Perfluorheksansulfonat (PFHxS)") %>%
  select(MYEAR, PARAM, STATION_CODE, LATIN_NAME, TISSUE_NAME, 
         VALUE_WW, FLAG1, UNIT, DRYWT, FAT_PERC) %>%
  group_by(MYEAR, PARAM, STATION_CODE, LATIN_NAME, TISSUE_NAME) %>%
  summarise(N = n(), Median_ww = median(VALUE_WW), 
            Min_ww = min(VALUE_WW), max_ww = max(VALUE_WW), 
            Q_limit = median(VALUE_WW[!is.na(FLAG1)])) %>%
  select(-PARAM)

# None of these:
data_ind2 %>%
  filter(PARAM == "Perfluorheksansulfonat (PFHxS)" & LATIN_NAME %in% "Somateria mollissima") %>%
  select(MYEAR, PARAM, STATION_CODE, LATIN_NAME, TISSUE_NAME, VALUE_WW, FLAG1, UNIT, DRYWT, FAT_PERC) %>%
  arrange(TISSUE_NAME, VALUE_WW)

```

#### Check uncertainty (UNCRT)
```{r}

# str(data_ind2)
xtabs(~ is.na(UNCRT), data_ind2)

df1 <- data_ind2 %>%
  mutate(UNCRT_lacking = is.na(UNCRT)) %>%
  group_by(PARAM) %>%
  summarise(UNCRT_lacking = mean(UNCRT_lacking)*100 %>% round(1),
            .groups = "drop")
df2 <- data_ind2 %>%
  mutate(UNCRT_lacking = is.na(UNCRT)) %>%
  group_by(PARAM, Lab) %>%
  summarise(UNCRT_lacking = mean(UNCRT_lacking)*100 %>% round(1),
            .groups = "drop") %>%
  pivot_wider(names_from = "Lab", values_from = "UNCRT_lacking")

df <- left_join(df1, df2)

df %>%
  filter(UNCRT_lacking == 0)


```


### b1. Get tables submitted to ICES in 2016
For comparison and getting the PARAM we want
```{r}
load(file = "../Milkys_2018/Input_data/22_ICES_tables_2016-10-06.RData")     # data file from script 22 used in 2017

data_03_2016 <- data_03
data_10_2016 <- data_10
data_21_2016 <- data_21
data_91_2016 <- data_91

```

### b2. Get tables submitted to ICES in 2019 (2018 data)
```{r}

folder_data <- "../Milkys/ICES/delivered to Mildir"
# dir(folder_data)
# fns <- dir(folder_data, ".NO", full.names = TRUE)
fn <- paste0(folder_data, "/NIVA2018CF_2nd_version.NO")
data_2018 <- read_ices_file(fn)  
data_2018 <- add_field_codes(data_2018)  

if (FALSE){
  
  names(data_2018)
  data_2018[["10"]] %>% xtabs(~PARAM, .)
  
  data_2018[["91"]] %>% xtabs(~STNNO, .)
  data_2018[["91"]] %>% xtabs(~STATN, .)

}

```




### c. Also get the newest 2017 data produced by script 10  
We call it 'data_ind2_10'  
We do this only to add uncertainty and quantification limit
```{r}

# NOT NEEDED - WE SHOULD ALREADY HAVE uncertainty and quantification limit
# str(data_ind2)

# Last year's statistics
# xtabs(~is.na(UNCRT), data_10_2016)   # UNCERTAINTY
# xtabs(~is.na(DETLI), data_10_2016)    # QUANTIFICATION_LIMIT

# xtabs(~is.na(QUANTIFICATION_LIMIT), data_ind2_10 %>% filter(PARAM %in% "EROD"))
# xtabs(~is.na(DETLI), data_10_2016 %>% filter(PARAM %in% "EROD"))
# data_10_2016 %>% filter(PARAM %in% "EROD") %>% pull(DETLI)

```

### d. Data used for submitting 2017 data
```{r}

df_ind_old <- read.csv("../Milkys_2018/Data/09_df_ind.csv", stringsAsFactors = FALSE)
df_liver_old <- read.csv("../Milkys_2018/Data/09_df_liver.csv", stringsAsFactors = FALSE)
df_muscle_old <- read.csv("../Milkys_2018/Data/09_df_muscle.csv", stringsAsFactors = FALSE)
df_liver_33F_old <- readRDS("../Milkys_2018/Data/09_df_liver_33F.RData")
df_muscle_33F_old <- readRDS("../Milkys_2018/Data/09_df_muscle_33F.RData")

```


### e1. Fish individual data
```{r}

# df_samples <- readRDS(file = "Data/01_df_samples_2019.rds")
# df_samples <- readRDS("Files_to_Jupyterhub_2020/Labware_samples_2020_2021-07-02.rds")  

folder <- paste0("Files_to_Jupyterhub_", selected_year)
dir(folder, pattern = "Labware_samples_*")

df_samples <- readRDS(paste0(folder, "/Labware_samples_2021_2022-09-21.rds")) %>%
  filter(STATUS %in% "Ferdig") %>%
  # Set 36A to 36A1
  mutate(
    AQUAMONITOR_CODE = ifelse(AQUAMONITOR_CODE == "36A", "36A1", AQUAMONITOR_CODE)
  ) %>%
  # Remove "SIA" lines 
  filter(!grepl("SIA", DESCRIPTION))


# Nice way of viewing these data  
if (FALSE)
  df_samples %>% 
    select(AQUAMONITOR_CODE, TISSUE, DESCRIPTION, BIOTA_SAMPLENO, X_BULK_BIO) %>% 
    View()

# if (selected_year == 2021){
#   
#   # 36A lacking for some reason
# 
#   df_samples <- bind_rows(
#     df_samples,
#     data.frame(
#       AQUAMONITOR_CODE = "36A", TISSUE = "SB-Whole soft body", DESCRIPTION = "",
#       BIOTA_SAMPLENO = 1:3, X_BULK_BIO = as.character(NA))
#   )
# 
# }


# Check again
check <- setdiff(unique(data_ind2$STATION_CODE), unique(df_samples$AQUAMONITOR_CODE))
if (length(check) > 0){
  stop("The station(s) ", paste(check, collapse = ", "), " are lacking in 'df_samples'.")
}

```
#### 

#### Check data  
```{r}

check <- df_samples %>%
  filter(TISSUE %in% "LI-Lever") %>%
  add_count(AQUAMONITOR_CODE, TISSUE, BIOTA_SAMPLENO) %>%
  filter(n > 1)
if (nrow(check) > 0){
  stop("Duplicates in liver data (error in section 1.e1)")
  # If error, check using the following lines:
  check %>% 
    select(AQUAMONITOR_CODE, TISSUE, DESCRIPTION, BIOTA_SAMPLENO, X_BULK_BIO) %>% 
    View("Liver samples")
}

check <- df_samples %>%
  filter(TISSUE %in% "MU-Muskel") %>%
  add_count(AQUAMONITOR_CODE, TISSUE, BIOTA_SAMPLENO) %>%
  filter(n > 1)
if (nrow(check) > 0){
  stop("Duplicates in muscle data (error in section 1.e1)")
  # If error, check using the following lines:
  check %>% 
    select(AQUAMONITOR_CODE, TISSUE, DESCRIPTION, BIOTA_SAMPLENO, X_BULK_BIO) %>% 
    View("Muscle samples")
}


```

### e2. Check for errors in df_samples    
Lacking X_BULK_BIO where there should have been a number   
- this is the case for 53B liver  
```{r}

check <- df_samples %>%
  # filter(AQUAMONITOR_CODE == "53B")  %>% 
  select(AQUAMONITOR_CODE, TISSUE, DESCRIPTION, BIOTA_SAMPLENO, X_BULK_BIO) %>% 
  mutate(
    DESCRIPTION = gsub("<f8>", "ø", DESCRIPTION, fixed = TRUE),
    DESCRIPTION = gsub("<c5>", "Å", DESCRIPTION, fixed = TRUE),
    DESCRIPTION = gsub("<e5>", "å", DESCRIPTION, fixed = TRUE),
    DESCRIPTION = gsub("<c6>", "Æ", DESCRIPTION, fixed = TRUE),
    DESCRIPTION = gsub("<e6>", "æ", DESCRIPTION, fixed = TRUE),
    DESCRIPTION2 = DESCRIPTION %>% stringr::str_sub(start = nchar(AQUAMONITOR_CODE) + 1),
    Fish_no = stringr::str_extract(DESCRIPTION2, "[0-9]+")) %>%
  filter(is.na(X_BULK_BIO)) %>%
  filter(Fish_no != BIOTA_SAMPLENO) %>%
  select(AQUAMONITOR_CODE, TISSUE, DESCRIPTION2, BIOTA_SAMPLENO, X_BULK_BIO, Fish_no)

check

```

### e3. Fix df_samples  
```{r}

# Check that we are doing the right thing below
if (FALSE){
  df_samples %>%
    mutate(DESCRIPTION = gsub("<f8>", "ø", DESCRIPTION, fixed = TRUE),
           DESCRIPTION = gsub("<c5>", "Å", DESCRIPTION, fixed = TRUE),
           DESCRIPTION = gsub("<e5>", "å", DESCRIPTION, fixed = TRUE),
           DESCRIPTION = gsub("<c6>", "Æ", DESCRIPTION, fixed = TRUE),
           DESCRIPTION = gsub("<e6>", "æ", DESCRIPTION, fixed = TRUE),
           DESCRIPTION2 = DESCRIPTION %>% stringr::str_sub(start = nchar(AQUAMONITOR_CODE) + 1),
           Fish_no = stringr::str_extract(DESCRIPTION2, "[0-9]+")) %>%
    filter(AQUAMONITOR_CODE == "53B" & TISSUE == "MU-Muskel") %>%
    select(DESCRIPTION, BIOTA_SAMPLENO, X_BULK_BIO, Fish_no)
}

df_samples <- df_samples %>%
  mutate(DESCRIPTION = gsub("<f8>", "ø", DESCRIPTION, fixed = TRUE),
         DESCRIPTION = gsub("<c5>", "Å", DESCRIPTION, fixed = TRUE),
         DESCRIPTION = gsub("<e5>", "å", DESCRIPTION, fixed = TRUE),
         DESCRIPTION = gsub("<c6>", "Æ", DESCRIPTION, fixed = TRUE),
         DESCRIPTION = gsub("<e6>", "æ", DESCRIPTION, fixed = TRUE),
         DESCRIPTION2 = DESCRIPTION %>% stringr::str_sub(start = nchar(AQUAMONITOR_CODE) + 1),
         Fish_no = stringr::str_extract(DESCRIPTION2, "[0-9]+")) %>%
  mutate(X_BULK_BIO = case_when(
    AQUAMONITOR_CODE == "53B" & TISSUE == "MU-Muskel" ~ as.character(Fish_no),
    TRUE ~ X_BULK_BIO)
  )

```


## 2. df_liver and df_muscle - for fish samples  
Basis: the file of LABWARE samples (df_samples)   


#### Liver except 33F
```{r}
# colnames(df_liver_old)
df_liver <- df_samples %>%
  filter(TISSUE %in% "LI-Lever") %>%
  # filter(!grepl("silo", DESCRIPTION, ignore.case = TRUE)) %>%
  filter(!AQUAMONITOR_CODE %in% "33F") %>%
  filter(BIOTA_SAMPLENO < 100) %>% # View("liver")
  group_by(AQUAMONITOR_CODE, BIOTA_SAMPLENO, X_BULK_BIO) %>%
  summarise(STATION_CODE = first(AQUAMONITOR_CODE),
            SAMPLE_NO = first(BIOTA_SAMPLENO),
            Pooled_no = first(X_BULK_BIO),
            .groups = "drop") %>%
  select(STATION_CODE, SAMPLE_NO, Pooled_no) %>%
  mutate(Pooled = 
           case_when(is.na(Pooled_no) ~ "enkel",
                     grepl(",", Pooled_no) ~ "bland",
                     TRUE ~ "enkel"))

# Get each fish number 
fish_numbers <- strsplit(df_liver$Pooled_no, ",")
df_liver$FiskNr_1 <- fish_numbers %>% map_chr(~.[1]) %>% as.numeric()
df_liver$FiskNr_2 <- fish_numbers %>% map_chr(~.[2]) %>% as.numeric()
df_liver$FiskNr_3 <- fish_numbers %>% map_chr(~.[3]) %>% as.numeric()

# For unpooled samples, and if nothing is given in X_BULK_BIO,
#   we set FiskNr_1 = SAMPLE_NO
sel <- is.na(df_liver$FiskNr_1)
df_liver$FiskNr_1[sel] <- df_liver$SAMPLE_NO[sel]

# Fake it till you make it!
# Add some fake weight data (i.e. how the sample is combined from different fish)
df_liver$Vekt1 <- rnorm(nrow(df_liver), 12, 0.5)
df_liver$Vekt2 <- rnorm(nrow(df_liver), 12, 0.5)
df_liver$Vekt3 <- rnorm(nrow(df_liver), 12, 0.5)
df_liver$Vekt1[is.na(df_liver$FiskNr_1)] <- NA
df_liver$Vekt2[is.na(df_liver$FiskNr_2)] <- NA
df_liver$Vekt3[is.na(df_liver$FiskNr_3)] <- NA

df_liver$sum <- apply(df_liver[,c("Vekt1", "Vekt2", "Vekt3")], 1, sum, na.rm = TRUE)

df_liver <- df_liver %>%
  select(-Pooled_no)

check <- df_liver %>%
  add_count(STATION_CODE, SAMPLE_NO) %>%
  filter(n > 1)
if (nrow(check) > 0){
  stop("Duplicates!")
}

```

```{r}
df_muscle <- df_samples %>%
  filter(TISSUE %in% "MU-Muskel" & AQUAMONITOR_CODE != "33F")
```

```{r}
df_muscle_old <- df_muscle
```



#### Muscle (except 33F)
```{r}

# colnames(df_liver_old)
df_muscle <- df_samples %>%
  filter(TISSUE %in% "MU-Muskel" & AQUAMONITOR_CODE != "33F") %>% # nrow()
  # filter(!AQUAMONITOR_CODE %in% "33F") %>%
  group_by(AQUAMONITOR_CODE, BIOTA_SAMPLENO, X_BULK_BIO) %>%
  summarise(STATION_CODE = first(AQUAMONITOR_CODE),
            SAMPLE_NO = first(BIOTA_SAMPLENO),
            Pooled_no = first(X_BULK_BIO),
            .groups = "drop") %>%
  select(STATION_CODE, SAMPLE_NO, Pooled_no) %>%
  mutate(Pooled = 
           case_when(is.na(Pooled_no) ~ "enkel",
                     grepl(",", Pooled_no) ~ "bland",
                     TRUE ~ "enkel"))

if (mean(df_muscle$Pooled == "enkel") == 1){
  cat("No pooled samples of muscle - OK to continue \n")
} else {
  cat(sum(df_muscle$Pooled == "bland"), "pooled samples of muscle! Check, and change code if necessary! \n")
}

if (FALSE)
  df_muscle %>%
    filter(Pooled == "bland")

# Use this code only if no pooled samples of muscle
# Otherwise d as for liver
df_muscle <- df_muscle %>%
  mutate(Fish_no = 
           case_when(is.na(Pooled_no) ~ SAMPLE_NO,
                     !is.na(Pooled_no) ~ as.numeric(Pooled_no))
  )

check <- df_muscle %>%
  add_count(STATION_CODE, SAMPLE_NO) %>%
  filter(n > 1)
if (nrow(check) > 0){
  stop("Duplicates!")
}

```


#### Liver 33F   
#### Muscle 33F

- No flatfish data in 2021! See old code if you need it  



## 3.df_ind - length and weight data  
Basis: the file of LABWARE samples (df_samples) and weight/length data (data_ind_fish)   
Fish individual Length/weight are added to df_ind  
```{r}

df_ind <- bind_rows(
  # Get STATION_CODE + Fish_no for each part
  df_liver %>% 
    select(STATION_CODE, FiskNr_1, FiskNr_2, FiskNr_3) %>%
    gather(Sampleno, Fish_no, -STATION_CODE) %>%
    mutate(Fish_no = as.numeric(Fish_no)) %>%
    select(-Sampleno),
  df_muscle %>% 
    count(STATION_CODE, Fish_no) %>%
    select(-n)
  # The next two is not needed in 2021 - no flatfish data  
  # tibble(STATION_CODE = "33F",
  #        Fish_no = df_liver_33F %>% purrr::transpose() %>% .$Fish_no %>% unlist() %>% sort()),
  # tibble(STATION_CODE = "33F",
  #        Fish_no = df_muscle_33F %>% purrr::transpose() %>% .$Fish_no %>% unlist() %>% sort())
  ) %>%
  # Delete duplicates of STATION_CODE + Fish_no
  count(STATION_CODE, Fish_no) %>%
  select(-n) %>%
  filter(!is.na(Fish_no))

nrow(df_ind)
# Add fish length and weight to the data
df_ind <- df_ind %>%
  safe_left_join(    # using safe_lef_join, see https://github.com/moodymudskipper/safejoin 
    data_ind_fish %>% select(STATION_CODE, Fish_no, Length, Weight, Gonad_weight),
    by = c("STATION_CODE", "Fish_no"),
    na_matches = "never"
  )

range(df_ind$Length, na.rm = TRUE)  # must not include zero
range(df_ind$Weight, na.rm = TRUE)  # must not include zero

```

### Check where we lack data  
- Lack some gonad weight data  
- 2019 - also lacking 13B, 19B, 43B2 and 53B
- 2021 - also lacking many 02B, 80B + some 10B, 23B, 24B, 36B  13B, 19B, 43B2 and 53B
```{r}
# df_ind
df_ind %>%
  group_by(STATION_CODE) %>%
  summarise_all(~mean(is.na(.)))

```


## Table 00
```{r}

data_00 <- data.frame(RECID = "00", RLABO = "NIVA", CNTRY = 58, MYEAR = selected_year, RFVER = "3.2.5")

```

## Table 03  
### Define SMPNO and SMLNK for each species   
    + SMPNO (sample number within station, links to table 04)
    + SMLNK (sample method, links to table 20)
    + I don't think these necessarily are linked to species, but in NIVA's case we let them be  
Updated with Eider duck Somateria mollissima in 2019 (must be updated also for 2018)  
    + SMPNO = 9 (More or less by random)  
    + SMLNK = 11 - links to table 20, where it refers to SMTYP = HAN (hand picking)   
```{r}

df_smpno <- read.table(textConnection("
            LATIN_NAME SMPNO  SMLNK
        'Gadus morhua'     1      8
     'Limanda limanda'     8      8
  'Littorina littorea'     3     11
      'Mytilus edulis'     0     11
    'Nucella lapillus'     3     11
  'Platichthys flesus'     6      9
'Somateria mollissima'     9     11
"), header = TRUE, stringsAsFactors = FALSE)

df_smpno <- df_smpno %>%
  filter(LATIN_NAME %in% unique(data_ind2$LATIN_NAME))  

check <- !unique(data_ind2$LATIN_NAME) %in% df_smpno$LATIN_NAME
if (sum(check) > 0){
  stop("Some species missing!")
}

df_smpno

```

### Create data 03 table  
By aggregating the data  
```{r}
df_station <- data_ind2 %>%
  # filter(MYEAR %in% myear & !LATIN_NAME %in% 'Somateria mollissima') %>%  # NOTE: DELETES EIDER DUCK
  # filter(MYEAR %in% myear) %>%  # NOTE: DELETES EIDER DUCK
  group_by(STATION_CODE, LATIN_NAME) %>%
  summarise(n(), .groups = "drop")

####################################################
# Add
#   Sample identification SMPNO (for each species in haul) 
#   Sample method link: SMLNK
####################################################

df_station <- df_station %>% 
  left_join(df_smpno, by = "LATIN_NAME", .groups = "drop")

data_03 <- data.frame(RECID = "03", 
                      CRUIS = "AA711", 
                      STNNO = df_station$STATION_CODE,
                      DTYPE = "CF", 
                      SMPNO = df_station$SMPNO,
                      SMLNK = df_station$SMLNK,
                      ATIME = NA, 
                      NOAGG = NA, 
                      SPECI = df_station$LATIN_NAME,
                      RLIST = "ER", 
                      FINFL = "GDP", 
                      stringsAsFactors = FALSE
                      )
```

### Check that SMPNO has been set
```{r}

xtabs(~SPECI + is.na(SMPNO), data_03)

```


## Table 04
### a. Fish, test
```{r}

ex_station <- "30B"
# ex_station <- "53B"

df1 <- df_muscle %>%
  filter(STATION_CODE %in% ex_station)
df2 <- df_liver %>%
  filter(STATION_CODE %in% ex_station) %>%
  as.data.frame()

# 1. df_muscle
df1 %>% head()
# 2. df_muscle
df2 %>% head()

# 3. Tables
cat("Example tables from raw data: Station 30B \n--------------------------------------------\n\n")
cat("Liver \n")
data_ind2 %>%
  filter(STATION_CODE %in% ex_station & MYEAR %in% selected_year & TISSUE_NAME %in% "Lever") %>%
  xtabs(~SAMPLE_NO2, .)
cat("\nMuscle \n")
data_ind2 %>%
  filter(STATION_CODE %in% ex_station & MYEAR %in% selected_year & TISSUE_NAME %in% "Muskel") %>%
  xtabs(~SAMPLE_NO2, .)

# Muscle
i1 <- df1$Fish_no                                                     # all 'muscle' fish
cat("Muscle fish (specimen numbers): \n")
i1

# Liver
bulkid <- with(df2, paste(FiskNr_1, FiskNr_2, FiskNr_3, sep = "~"))   # collect ind. numbers
bulkid <- gsub("~NA", "", bulkid, fixed = TRUE)                       # remove NA's
pooled <- grepl("~", bulkid)                                          # which are pooled? 
i2 <- df2$FiskNr_1[!pooled]                                           # get all unpooled fish
cat("Fish without pooled samples: \n")
i2

# Muscle samples (assumed to be unpooled)
df <- tibble(SPECI = "Gadus morhua", Tissue = "Muscle", SUBNO = i1, BULKID = "")
cat("\nMake df: \n")
cat("- start with 'muscle' fish - nrow:", nrow(df), "\n")

# Add unpooled liver samples NOT overlapping with muscle samples (quite rare situation)
liver_extra_unpooled_samples <- i2[!i2 %in% i1]
if (length(liver_extra_unpooled_samples) > 0)
  df <- rbind(
    df,
    tibble(SPECI = "Gadus morhua", Tissue = "Liver", SUBNO = liver_extra_unpooled_samples, BULKID = "")
  )
cat("- add unpooled liver samples NOT overlapping with muscle samples - nrow:", nrow(df), "\n")
liver_pooled_samples_n <- sum(pooled)

# Add pooled liver samples 
if (liver_pooled_samples_n > 0){
  subno_liver <- max(i1, liver_extra_unpooled_samples) + seq(1, liver_pooled_samples_n)
  df <- rbind(
    df,
    tibble(SPECI = "Gadus morhua", Tissue = "Liver", SUBNO = subno_liver, BULKID = bulkid[pooled])
  )
}
cat("- pooled samples added - nrow:", nrow(df), "\n")

# 4. Final df
df %>% as.data.frame()

```


### b. Fish, function    
Note:  
    + Muscle samples are assumed to be unpooled (except for 33F, which is treated specially below)  
    + The case of "unpooled liver samples NOT overlapping with muscle samples" has not been tested  
    + See "ICES/ICES - notes on SUBNO, BULKID etc.docx" for explanation  
```{r}

generate_04_fish_station <- function(station_code, latin_name, 
                                     df_individuals_muscle, df_individuals_liver,
                                     lookuptable_smpno = df_smpno){
  df1 <- df_individuals_muscle %>%
    filter(STATION_CODE %in% station_code)
  df2 <- df_individuals_liver %>%
    filter(STATION_CODE %in% station_code)
  fishno_muscle <- df1$Fish_no %>% sort()
  fishno_liver <- with(df2, c(FiskNr_1, FiskNr_2, FiskNr_3)) %>% unique() %>% sort()
  number_of_fish_liver <- apply(!is.na(df2[,c("FiskNr_1", "FiskNr_2", "FiskNr_3")]), 1, sum)
  pooled <- number_of_fish_liver > 1  
  # Start by listing all individual fish                           
  df <- tibble(STATION_CODE = station_code, LATIN_NAME = latin_name, 
               Fish_no = unique(c(fishno_muscle, fishno_liver)), NOIMP = 1)
  # Muscle samples (assumed to be unpooled)
  df <- df %>%
    left_join(df1 %>% select(Fish_no, SAMPLE_NO), by = "Fish_no") %>% 
    rename(SAMPLE_NO_Muscle = SAMPLE_NO)
  # Liver samples - create bulkid by joining Fisknr and removing "~NA"
  bulkid <- with(df2, paste(FiskNr_1, FiskNr_2, FiskNr_3, sep = "~")) %>%
    gsub("~NA", "", ., fixed = TRUE)
  # Unpooled liver samples overlapping with muscle samples
  df <- df %>%
    left_join(df2[!pooled, c("FiskNr_1", "SAMPLE_NO")], 
              by = c("Fish_no" = "FiskNr_1")) %>%
    rename(SAMPLE_NO_Liver = SAMPLE_NO)
  df$SUBNO <- df$Fish_no                         # for these samples SUBNO equals fish number Fish_no
  # BULKID for all unpooled samples is just 
  # Pooled liver samples 
  liver_pooled_samples_n <- sum(pooled)
  if (liver_pooled_samples_n > 0){
    df2p <- df2[pooled,]
    df2p$SUBNO <- max(df$SUBNO) + 1:nrow(df2p)    # make new SUBNO, starting where we left off
    df3 <- df2p %>% 
      select(SAMPLE_NO, SUBNO, FiskNr_1:FiskNr_3) %>% 
      gather("Part", "Fish_no", FiskNr_1:FiskNr_3) %>% 
      filter(!is.na(Fish_no))
    df3summ <- df3 %>% 
      group_by(Fish_no) %>% 
      summarise(BULKID = paste(SUBNO, sep = "~"))
    df_to_add <- tibble(
      STATION_CODE = station_code, LATIN_NAME = latin_name, 
      Fish_no = NA, 
      NOIMP = number_of_fish_liver[pooled],
      SAMPLE_NO_Muscle = NA,
      SAMPLE_NO_Liver = df2p$SAMPLE_NO,
      SUBNO = df2p$SUBNO,
      BULKID = "")
    df <- left_join(df, df3summ, by = "Fish_no")  # adds BULKID to the relevant fish
    df$BULKID[is.na(df$BULKID)] <- ""             # BULKID = NA is changed to BULKID = ""
    df <- rbind(df, df_to_add)
  }
  # Add SMPNO
  df %>% left_join(lookuptable_smpno %>% select(LATIN_NAME, SMPNO), by = "LATIN_NAME")
}
# debugonce(generate_04_fish_station)
# df <- generate_04_fish_station("24B", "Gadus morhua", df_muscle, df_liver)
df <- generate_04_fish_station("71B", "Gadus morhua", df_muscle, df_liver)
df <- generate_04_fish_station("02B", "Gadus morhua", df_muscle, df_liver)
df %>% as.data.frame()

# For further exploration

explore_table_04 <- FALSE
if (explore_table_04){
  df_samples %>% 
    filter(AQUAMONITOR_CODE == "53B" & TISSUE %in% "MU-Muskel") %>% 
    arrange(BIOTA_SAMPLENO) %>%
    select(DESCRIPTION, BIOTA_SAMPLENO, X_BULK_BIO)
  df %>%
    filter(!is.na(SAMPLE_NO_Muscle)) %>%
    arrange(SAMPLE_NO_Liver)
}
```



### c. Create 'data_04_fish_main'  
I.e. data_04 for all fish stations except 33F ('main table')
```{r}

stat_fish <- unique(df_muscle$STATION_CODE)
df_station_species <- data_ind2 %>%
  filter(STATION_CODE %in% stat_fish & MYEAR %in% selected_year) %>%
  group_by(STATION_CODE, LATIN_NAME) %>%
  summarise(N = n())
data_04_fish_main <- 1:nrow(df_station_species) %>%  
  map_df(~generate_04_fish_station(
    df_station_species$STATION_CODE[.x], df_station_species$LATIN_NAME[.x],
    df_muscle, df_liver)
    )

# Check
data_04_fish_main %>% filter(STATION_CODE == "36B") %>% as.data.frame()

# Check for duplicates  
check <- data_04_fish_main %>%
    group_by(STATION_CODE, SMPNO, SUBNO) %>%
    mutate(n = n()) %>%
    filter(n > 1) %>%
    arrange(STATION_CODE, SMPNO, SUBNO)

if (nrow(check) > 0){
  cat(nrow(check), "duplicate data in 'data_04_fish_main' \n")
  stop("This will lead to duplicates in table 04. Please fix! (error occured in section 'Table 04 c')")
} else {
  cat("No duplicates in 'data_04_fish_main' \n")
}

```

### d. Comparison with sample file   
One example station  
```{r}

example_station <- "23B"

df_samples %>%
  filter(AQUAMONITOR_CODE %in% example_station & TISSUE == "LI-Lever") %>% 
  select(AQUAMONITOR_CODE, TISSUE, DESCRIPTION, BIOTA_SAMPLENO, X_BULK_BIO) %>%
  arrange(TISSUE, BIOTA_SAMPLENO)


```



### e. Same station in what will become table 04   
Uses 'example_station' from previous chunk  
Example of unpooled sample: 
- SAMPLE_NO_Liver = 6 (the SAMPLE_NO for the liver) - corresponds to fish specimen 7  
Example of pooled sample: 
- SUBNO = 22 (a single line) corresponds to a single pooled liver sample, with SAMPLE_NO = 8   
- BULKID = 22 (three lines) corresponds to the three fish (9,14,20) the liver sample is taken from,
each of which has a length and weight, at 9 and 14 has in addition a muscle sample   
```{r}

data_04_fish_main %>% 
  filter(STATION_CODE == example_station) %>%
  # select(STATION_CODE, SUBNO, BULKID) %>%
  mutate(Sort = case_when(
    is.na(BULKID) ~ SUBNO,
    BULKID == "" ~ SUBNO,
    !is.na(BULKID) ~ as.numeric(BULKID))
  ) %>%
  arrange(Sort) %>%
  select(-Sort)

```


### f. Check number of individuals per sample 
```{r}
cat("Muscle - number of fish per sample:\n")
data_04_fish_main %>%
  filter(!is.na(SAMPLE_NO_Muscle)) %>%
  count(STATION_CODE, SUBNO, NOIMP) %>%
  xtabs(~NOIMP + STATION_CODE, .)

cat("\n\nLiver - number of fish per sample:\n")
data_04_fish_main %>%
  filter(!is.na(SAMPLE_NO_Liver)) %>%
  count(STATION_CODE, SUBNO, NOIMP) %>%
  xtabs(~NOIMP + STATION_CODE, .)
```

### g. Fish table, add 33F (special case)  
Assuming that 
    + Liver and muscle samples are from the same fish
    + All samples are pooled  
```{r}
#
# We cheat and just do this manually
#

# # One line per fish, for length and weight measurements   
# df1 <- tibble(STATION_CODE = "33F", LATIN_NAME = "Platichthys flesus",
#               Fish_no = 1:15,
#               NOIMP = 1,
#               SAMPLE_NO_Muscle = NA, SAMPLE_NO_Liver = NA,
#               SUBNO = as.numeric(1:15), BULKID = as.character(rep(16:18, each = 5)))
# 
# # One line per pooled sample (= 3 lines altogether))
# df2 <- tibble(STATION_CODE = "33F", LATIN_NAME = "Platichthys flesus",
#               Fish_no = NA,
#               NOIMP = 5,
#               SAMPLE_NO_Muscle = 1:3, SAMPLE_NO_Liver = 1:3,
#               SUBNO = as.numeric(16:18), 
#               BULKID = "")
# df <- rbind(df1, df2)
# 
# data_04_fish_33F <-  df %>% left_join(df_smpno %>% select(LATIN_NAME, SMPNO), by = "LATIN_NAME")

```

### h. Check 33F
```{r}
# 
# # Sample file
# df_samples %>%
#   filter(AQUAMONITOR_CODE == "33F") %>% 
#   select(AQUAMONITOR_CODE, TISSUE, DESCRIPTION, BIOTA_SAMPLENO, X_BULK_BIO) %>%
#   arrange(TISSUE, BIOTA_SAMPLENO)
# 
# # Table 04  
# data_04_fish_33F %>%
#   mutate(Sort = case_when(
#     BULKID == "" ~ SUBNO,
#     !is.na(BULKID) ~ as.numeric(BULKID))
#   ) %>%
#   arrange(Sort) %>%
#   select(-Sort)

```

### i1. Extra samples that may be needed for fish biological effects  

```{r}

fish_species <- "Gadus morhua"
sel_tissues <- c("Blod", "Galle", "Liver - microsome")

df1 <- data_ind2 %>%
  filter(LATIN_NAME %in% fish_species & 
           (TISSUE_NAME %in% sel_tissues | PARAM %in% "EROD"))  %>%
  # filter(STATION_CODE == "53B") %>%
  rename(Fish_no = SAMPLE_NO) %>%       # because for biological effects, the sample numbers refer to specimen number
  distinct(STATION_CODE, Fish_no)

df2 <- data_04_fish_main %>%
  # filter(STATION_CODE == "53B") %>%
  distinct(STATION_CODE, Fish_no) %>%
  mutate(In_table_04 = TRUE)

# df1
# df2

check <- df1 %>%
  left_join(df2, by = c("STATION_CODE", "Fish_no")) %>%
  filter(is.na(In_table_04))

cat(nrow(check), "specimens lack in 'data_04_fish_main'")

if (nrow(check) > 0){
   
  stop("You must create 'data_04_fish_extra' for the lacking specimens. See next chunk. (Error in section 'Table 04 i1')")
  
} else {
  
  cat("\nNo specimens lacking, no need to create extra table 04 data for biological effect data")
  
}


```

### i2. Fish table, fish specimens only used for biological effects  
- If any speciemns in 23B, 53B, 30B are only used for biological effects, they have not been included in df_samples and therefore they don't make it into 'data_04_fish_main'   
- Note: code is not finished - see comments    
```{r}

if (nrow(check) > 0){
  
  # NOTE: this code is not finished - must add new SUBNO numbers (wher it says 'extra_subno_numbers' in the code
  # These SUBNO numbers must continue where the former SUBNO numbers stop for this station
  
  # NOTE: at least works if check has a single station 
  max_subno <- data_04_fish_main %>%
    filter(STATION_CODE %in% unique(check$STATION_CODE)) %>%
    pull(SUBNO) %>%
    max()
  
  extra_subno_numbers <- max_subno + 1
    
  data_04_fish_extra <- check %>%
    mutate(LATIN_NAME = "Gadus morhua",
           NOIMP = 1,
           SAMPLE_NO_Muscle = NA, SAMPLE_NO_Liver = 4,
           SUBNO = extra_subno_numbers, BULKID = "",
           SMPNO = 1)
  
  data_04_fish_main <- bind_rows(
    data_04_fish_main, 
    data_04_fish_extra)
  
} else {
  
  cat("These extra data ('data_04_fish_extra') is not needed in this case (see previous chunk) ")
  
}

```


### j1. Fish table, join together  
```{r}

no_flatfish <- identical(
  unique(subset(data_ind2, TISSUE_NAME == "Lever")$LATIN_NAME), 
  "Gadus morhua") 

if (no_flatfish){

  data_04_fish <- data_04_fish_main
   
} else if (nrow(check) == 0){

  data_04_fish <- rbind(data_04_fish_main, data_04_fish_33F)
   
} else {

   data_04_fish <- rbind(data_04_fish_main, data_04_fish_33F, data_04_fish_extra)
  
}

cat("'data_04_fish' created - it has", nrow(data_04_fish), "lines \n")

```


### j2. Check fish table
```{r}

check <- data_04_fish %>%
    group_by(STATION_CODE, SMPNO, SUBNO) %>%
    mutate(n = n()) %>%
    filter(n > 1) %>%
    arrange(STATION_CODE, SMPNO, SUBNO)

if (nrow(check) > 0){
  cat(nrow(check), "duplicate data in 'data_04_fish' \n")
  stop("This will lead to duplicates in table 04. Please fix! (error occured in section 'Table 04 j2')")
} else {
  cat("No duplicates in 'data_04_fish' \n")
}

if (FALSE){

  check2 <- data_04_fish_main %>%
    group_by(STATION_CODE, SMPNO, SUBNO) %>%
    mutate(n = n()) %>%
    filter(n > 1) %>%
    arrange(STATION_CODE, SMPNO, SUBNO)
  nrow(check2)
  
  check2 <- data_04_fish_extra %>%
    group_by(STATION_CODE, SMPNO, SUBNO) %>%
    mutate(n = n()) %>%
    filter(n > 1) %>%
    arrange(STATION_CODE, SMPNO, SUBNO)
  nrow(check2)
  
}
```


### k. Mussels 
```{r}
data_04_bluemussel <- data_ind2 %>%
  filter(LATIN_NAME %in% "Mytilus edulis" & MYEAR %in% selected_year) %>%
  mutate(SAMPLE_NO2 = SAMPLE_NO) %>%
  group_by(STATION_CODE, LATIN_NAME, SAMPLE_NO2) %>%
  summarise(N = n(), .groups = "drop")
data_04_bluemussel <- data_04_bluemussel %>% 
  left_join(df_smpno %>% select(LATIN_NAME, SMPNO), 
            by = "LATIN_NAME")
```

### l. Snails
```{r}
data_04_snail <- data_ind2 %>%
  filter(LATIN_NAME %in% c("Nucella lapillus", "Littorina littorea") & MYEAR %in% selected_year) %>%
  group_by(STATION_CODE, LATIN_NAME) %>%
  summarise(SAMPLE_NO2 = first(SAMPLE_NO),
             .groups = "drop") %>%
  left_join(df_smpno %>% select(LATIN_NAME, SMPNO), 
            by = "LATIN_NAME")
data_04_snail
```


### m. Eider duck  
For eider duck, the same SAMPLE_NO2 = 1 for blood and SAMPLE_NO2 = 1 for egg doesn't mean that this is the same individual (obviously). So we use SUBNO 1-15 for blood SAMPLE_NO 1-15, and SUBNO 16-30 for egg SAMPLE_NO 1-15.  
```{r}

data_04_eider <- 
  data_ind2 %>%
  filter(LATIN_NAME %in% c("Somateria mollissima") & MYEAR %in% selected_year) %>% # View()
  # excude siloxans - some samples have only siloxans
  filter(!PARAM %in% c("D4", "D5", "D6")) %>%   
  count(STATION_CODE, LATIN_NAME, TISSUE_NAME, SAMPLE_NO) %>% # View()
  arrange(TISSUE_NAME, SAMPLE_NO) %>%            # first blood SAMPLE_NO 1-15, then egg SAMPLE_NO 1-15.   
  rename(SAMPLE_NO2 = SAMPLE_NO) %>%
  left_join(df_smpno %>% select(LATIN_NAME, SMPNO), by = "LATIN_NAME") %>%
  ungroup()
data_04_eider$SUBNO <- 1:nrow(data_04_eider)     # SUBNO given = 1-30
data_04_eider

```

### n. Put together data_04  
For blue mussel and snail, SUBNO = SAMPLE_NO2  
For fish, we have constructed SUBNO (see above)  
For eider, SUBNO 
```{r}
data_04_1 <- data.frame(
  RECID = "04", 
  CRUIS = "AA711", 
  STNNO = data_04_fish$STATION_CODE,
  SMPNO = data_04_fish$SMPNO,
  SUBNO = data_04_fish$SUBNO,
  NOIMP = data_04_fish$NOIMP,
  ORGSP = NA,
  SEXCO = NA,
  STAGE = NA,
  CONES = NA,
  ASTSA = NA,
  NODIS = NA,
  BULKID = data_04_fish$BULKID,
  stringsAsFactors = FALSE
  )

data_04_2 <- data.frame(
  RECID = "04", 
  CRUIS = "AA711", 
  STNNO = data_04_bluemussel$STATION_CODE,
  SMPNO = data_04_bluemussel$SMPNO,
  SUBNO = data_04_bluemussel$SAMPLE_NO2,
  NOIMP = 20,
  ORGSP = NA,
  SEXCO = NA,
  STAGE = NA,
  CONES = NA,
  ASTSA = NA,
  NODIS = NA,
  BULKID = "",
  stringsAsFactors = FALSE
  )

data_04_3 <- data.frame(
  RECID = "04", 
  CRUIS = "AA711", 
  STNNO = data_04_snail$STATION_CODE,
  SMPNO = data_04_snail$SMPNO,
  SUBNO = data_04_snail$SAMPLE_NO2,
  NOIMP = 50,
  ORGSP = NA,
  SEXCO = "F",
  STAGE = NA,
  CONES = NA,
  ASTSA = NA,
  NODIS = NA,
  BULKID = "",
  stringsAsFactors = FALSE
  )

data_04_4 <- data.frame(
  RECID = "04", 
  CRUIS = "AA711", 
  STNNO = data_04_eider$STATION_CODE,
  SMPNO = data_04_eider$SMPNO,
  SUBNO = data_04_eider$SUBNO,
  NOIMP = 1,
  ORGSP = NA,
  SEXCO = "F",
  STAGE = NA,
  CONES = NA,
  ASTSA = NA,
  NODIS = NA,
  BULKID = "",
  stringsAsFactors = FALSE
  )

# Without Somateria /Eider duck:
# data_04 <- rbind(data_04_1, data_04_2, data_04_3)

# With Somateria /Eider:
data_04 <- rbind(data_04_1, data_04_2, data_04_3, data_04_4)


# Checks
# data_04_1 %>% filter(STNNO %in% "36B")
# data_04 %>% filter(STNNO %in% "36B")
# data_04 %>% filter(is.na(SUBNO))
# data_04 %>% filter(SUBNO == "")

```

### o. Check the crucial columns
```{r}
check <- apply(
  is.na(data_04 %>% select( RECID,  CRUIS,  STNNO,  SMPNO,  SUBNO,  NOIMP)), 
  2, sum
  )
if (sum(check == rep(0, 6)) == 6){
  cat("All records have values for RECID,  CRUIS,  STNNO,  SMPNO,  SUBNO,  NOIMP\n\n")
} else {
  cat("ERROR! One or more columns RECID,  CRUIS,  STNNO,  SMPNO,  SUBNO,  NOIMP lack data\n\n")
}

apply(is.na(data_04), 2, sum)
# data_04 %>% filter(STNNO %in% c("24B", "53B", "96B")) %>% arrange(STNNO, SMPNO, SUBNO)


# data_04 %>% filter(is.na(SUBNO))


```

## Table 21
We make this table before table 10 in order to generate the AMLNK values we need    
Note that the AMLNK values are just random, i.e. alphabetical for non-biological effects and
then biological effects added in the end. They have only meaning internally in this data set
(for coupling to table 10)    
SO FAR WE JUST USE LAST YEAR'S TABLE :-)  

### Generate lookup
```{r}
df_amlnk <- data_10_2016 %>%
  group_by(PARAM, AMLNK) %>%
  summarise(n = n(),
            .groups = "drop")
df_amlnk$n <- NULL
```


### Add 'new' parameters  
- GOSOI (gonadal somatic index = (gonad weight/whole organism weight) x 100), https://vocab.ices.dk/?CodeID=51654     
- INTF% = intersex, https://vocab.ices.dk/?CodeID=53682       
- C13D and N15D = isotope ratio δ13C:12C (https://vocab.ices.dk/?CodeID=197194) and isotope ratio δ15N:14N (https://vocab.ices.dk/?CodeID=197195)   

```{r}

params_new <- c("GOSOI", "INT%", "C13D", "N15D")
# params_new <- c("GOSOI", "INT%")

for (param in params_new){
  if (sum(df_amlnk$PARAM %in% param) == 0){
    # add one row to df_amlnk
    df_amlnk <- bind_rows(
      df_amlnk,
      tibble(
        PARAM = param, 
        AMLNK = max(df_amlnk$AMLNK) + 1)    # add 1 to the last AMLNK number  
    )
  }
}

```



## Table 10  
Table 10 is the combined rows of the following tables:
    - data_10_fish (part c)  
    - data_10_bluemussel (part d)  
    - data_10_length2 (part g3)  
    - data_10_vdsi (part h)  

### a1. Make data_conc_1   
**Note make sure that all estimates are on wet-weight basis**    
- Note that siloxanes "D4", "D5", "D6" are not included because they are not in the ICES vocabulary  
```{r}

par_ices <- unique(data_10_2016$PARAM) %>% sort()
# Add some
par_ices <- c(par_ices, "HCHA", "HCHG", "HCB") %>% unique() %>% sort() 
#
# Should also have added "D4", "D5", "D6"
# But siloxans are not in the ICES vocabulary!
# (I also searched fro e.g. 'Decamethylcyclopentasiloxane' (= D5))

par_niva <- data_ind2 %>% filter(MYEAR %in% selected_year) %>% pull(PARAM) %>% unique() %>% sort()
par_ices[!par_ices %in% par_niva]

data_conc_1 <- data_ind2 %>% 
  filter(MYEAR %in% selected_year) %>%
  mutate(SAMPLE_NO2 =                              # setting SAMPLE_NO2 (which is used in 2018 version)
           case_when(!is.na(SAMPLE_NO) ~  SAMPLE_NO,
                     is.na(SAMPLE_NO) ~  SAMPLE_NO2),
         BASIS = "W"                             # MAKE SURE ALL MEASUREMENS ARE WET WEIGHT 
  )

# data_ind2 %>% 
#   filter(STATION_CODE == "23B" & PARAM %in% "EROD") %>% 
#   select(STATION_CODE, PARAM, SAMPLE_NO, SAMPLE_NO2)

```

### a2. Change parameter names where ICES name differs  
```{r}

# Change BDE to BD for the following. NOTE: "BDE28"  "BDE47"  "BDE99" should NOT be changed (!)
sel <- data_conc_1$PARAM %in% c("BDE100", "BDE126", "BDE153", "BDE154", "BDE183", "BDE209")
data_conc_1$PARAM[sel] <- sub("BDE", "BD", data_conc_1$PARAM[sel])

sel <- data_conc_1$PARAM %in% "PFUdA"; sum(sel)
data_conc_1$PARAM[sel] <- "PFUnda"
sel <- data_conc_1$PARAM %in% "PFDcA"; sum(sel)
data_conc_1$PARAM[sel] <- "PFDA"

sel <- data_conc_1$PARAM %in% "TBT"; sum(sel)   # Note this is TBT cation weight = 2.44 * TBTIN (TBT tin weight)
data_conc_1$PARAM[sel] <- "TBSN+"

sel <- data_conc_1$PARAM %in% "Delta13C"; sum(sel)  
data_conc_1$PARAM[sel] <- "C13D"
sel <- data_conc_1$PARAM %in% "Delta15N"; sum(sel)  
data_conc_1$PARAM[sel] <- "N15D"

# This did NOT do in 2018:
# Fat:
sel <- data_conc_1$PARAM %in% "Fett"; sum(sel)
data_conc_1$PARAM[sel] <- "FATWT%"

# The new (2021) "SCCP" and "MCCP" should be OK. For LOQ-values: value = LOQ, Flag1 = "<")
# (Note: the "eskl. LOQ" vesrions have these set to value = 0!)
# but for previous years probably the "including LOQ-version" has been used

```

### a2. Filter data to keep only those parameters that we need
```{r}

# Parameters 
par_niva2 <- data_conc_1 %>% pull(PARAM) %>% unique() %>% sort()
cat("\nCheck parameters not found in this year's data:  \n")
par_ices[!par_ices %in% par_niva2]   # "FATWT%" "LNMEA"  "TBP"    "WTMEA"  
                                     # We will add "FATWT%" and "LNMEA" later

# Let us use TBSN+ instead of TBTIN
par_ices <- par_ices[!par_ices %in% "TBTIN"]
par_ices <- c(par_ices, "TBSN+")
# par_ices

# Filter data to keep only those parameters that we need
data_conc_2 <- data_conc_1 %>% filter(PARAM %in% par_ices)

cat("\nCheck final parameters:\n")
tab <- data_conc_2 %>% xtabs(~PARAM, .)
tab

# cat("\n\nCheck excluded parameters:\n")
# data_conc_1 %>% filter(!PARAM %in% par_ices) %>% xtabs(~PARAM, .)

# Parameters with data over LOQ that are excluded:
cat("\n\nParameters with data over LOQ that are excluded:\n")
tab <- data_conc_1 %>% 
  filter(!PARAM %in% par_ices & is.na(FLAG1)) %>% 
  xtabs(~PARAM, .) %>% 
  sort(decreasing = TRUE)
tab
# dput(names(tab)) 

```


### a3. Check that basis is only W (wet weight basis)
```{r}

data_conc_2 %>% xtabs(~BASIS, .)

```

### a4. Check availability of dry weight  
- Should be lacking only for blood and egg   
```{r}
# Dry weight:
# sel <- data_conc_2$PARAM %in% "DRYWT%"; sum(sel)

# Check availability of dry weight for CB118 (as example for liver)
data_conc_2 %>%
  filter(PARAM %in% c("DRYWT%", "CB118")) %>%
  select(STATION_CODE, TISSUE_NAME, SAMPLE_NO2, PARAM, VALUE_WW) %>% 
  arrange(STATION_CODE, TISSUE_NAME, SAMPLE_NO2, PARAM) %>%
  spread(PARAM, VALUE_WW) %>%
  filter(!is.na(CB118)) %>%
  count(STATION_CODE, TISSUE_NAME, Lack_dryweight = is.na(`DRYWT%`)) %>%
  filter(Lack_dryweight)

# Check availability of dry weight for HG (as example for muscle) 
data_conc_2 %>%
  filter(PARAM %in% c("DRYWT%", "HG")) %>%
  select(STATION_CODE, TISSUE_NAME, SAMPLE_NO2, PARAM, VALUE_WW) %>% 
  arrange(STATION_CODE, TISSUE_NAME, SAMPLE_NO2, PARAM) %>%
  spread(PARAM, VALUE_WW) %>%
  filter(!is.na(HG)) %>%
  count(STATION_CODE, TISSUE_NAME, Lack_dryweight = is.na(`DRYWT%`)) %>%
  filter(Lack_dryweight)

# Check bird data
# data_conc_2 %>%
#   filter(STATION_CODE == "19N") %>%
#   count(TISSUE_NAME, PARAM)

```

### a5. Check uncertainty (UNCRT)   
If problems here, check either   
    - script 34 (for chemical variables from Eurofins or NILU)  
    - script 02 (for VDSI and the biological effect variables in cod)  
```{r}
labels <- c("1 - Under LOQ", "2 - Over LOQ, NA", 
          "3 - Over LOQ, zero", "4 - Over LOQ, OK")

# xxx
data_conc_2 %>% 
  filter(PARAM == "BDE47" & is.na(FLAG1)) %>%
  xtabs(~Lab + is.na(UNCRT), .)

df_uncertainty <- data_conc_2 %>%
  mutate(Uncert = 
           case_when(
             !is.na(FLAG1) ~ labels[1],
             is.na(UNCRT) ~ labels[2],
             UNCRT == 0 ~ labels[3],
             UNCRT != 0 ~ labels[4]),
         Uncert = factor(Uncert, levels = labels)
  )

tab <- df_uncertainty %>%
  count(PARAM, Lab, Uncert) %>%
  pivot_wider(names_from = "Uncert", values_from = "n", values_fill = 0, names_sort = TRUE)
tab

sel <- names(tab) %in% labels
tot <- apply(tab[sel], 1, sum, na.rm = TRUE)       # note hard-coded columns
sel <- names(tab) %in% c("2 - Over LOQ, NA", "3 - Over LOQ, zero")
not_ok <- apply(tab[sel], 1, sum, na.rm = TRUE)
sel <- names(tab) %in% c("2 - Over LOQ, NA", "3 - Over LOQ, zero", "4 - Over LOQ, OK")
over_loq <- apply(tab[sel], 1, sum, na.rm = TRUE)

tab <- tab %>%
  mutate(Total = tot,
         NotOK = round(not_ok/over_loq*100, 1)
         ) %>%
  filter(Total > 0) %>%
  arrange(NotOK)

tab

# To print list of parameters
# tab %>%
#   filter(NotOK == 100) %>%
#   pull(PARAM) %>% 
#   dput()


```


### a6-1. GOSOI - gonadal somatic index, defined as (gonad weight/whole organism weight) x 100   
    - Must be given for all EROD samples  
    - Taken from data_ind_fish
    - Creates `data_conc`    
```{r}   

# Make data to go into 'data_to_add' below (Gonadal somatic index )
data_for_leftjoin <- data_ind_fish %>%
  mutate(VALUE_WW = Gonad_weight/Weight*100) %>%
  filter(!is.na(VALUE_WW)) %>%
  select(STATION_CODE, Fish_no, VALUE_WW)
  
# Check data 1
# data_conc_2[sel,] %>%
#  filter(PARAM %in% "EROD") %>%
#   count(STATION_CODE)

# Check data 2
# data_conc_2[sel,] %>%
#  filter(PARAM %in% "EROD") %>%
#  xtabs(~STATION_CODE + SAMPLE_NO2, .)

# Copy EROD data and replace the columns we need
# These will be extra rows in the data set
data_to_add <- data_conc_2 %>%
  filter(PARAM %in% "EROD") %>%
  mutate(TISSUE_NAME = "WO",
         MATRX = "WO",
         PARAM = "GOSOI",       
         BASIS = NA,
         UNIT = "idx") %>%       # for "index" (MUNIT will be set to the same)
  select(-VALUE_WW, -UNCRT) %>%  # this we replace by left_join below
  safe_left_join(data_for_leftjoin, 
                 by = c("STATION_CODE", "SAMPLE_NO2" = "Fish_no"),
                 na_matches = "never"
  ) %>%
  filter(!is.na(VALUE_WW)) %>%
  mutate(UNCRT = 5,     # We guess uncertainity is SD = 5%
         METCU = "%")

# Show data that will be added
if (FALSE){
  data_to_add %>%
    select(STATION_CODE, SAMPLE_NO2, PARAM, VALUE_WW, UNIT)
}

# Add to data (if not already added)
if (sum(data_conc_2$PARAM %in% "GOSOI") == 0){      # Make sure we add data only once
  data_conc <- bind_rows(data_conc_2, data_to_add)
}

```

### a6-2. Remove EROD without GOSOI    
- EROD withou GOSOI is an error in the OCES database ('Cross record/field check (condition not met)')    
```{r}

data_conc_back <- data_conc
# data_conc <- data_conc_back

data_conc <- data_conc %>%
  left_join(
    data_conc %>%
      filter(PARAM %in% "GOSOI") %>%
      select(STATION_CODE, SAMPLE_NO2, VALUE_WW) %>%
      rename(GOSOI = VALUE_WW),
    by = c("STATION_CODE", "SAMPLE_NO2")
  ) %>%
  filter(!(PARAM %in% "EROD" & is.na(GOSOI)))

message(nrow(data_conc_back) - nrow(data_conc), " EROD observations without GOSOI deleted")

```




### b. Unit lookup table  
- From UNIT (Nivabase) to MUNIT (ICES)  
```{r}

xtabs(~UNIT, data_conc)
xtabs(~MUNIT, data_10_2016)

# we need only these units, as the rest will be dropped
df_munit <- data.frame(
  UNIT = c("MG_P_KG", "NG_P_G", "UG_P_KG", "PERCENT",
           "ng/min/mg protein", "pmol/min/mg protein", 
           "idx"),
  MUNIT = c("mg/kg", "ng/kg", "ug/kg", "%",
            "ng/min/mg protein", "pmol/min/mg protein", 
            "idx"), 
  stringsAsFactors = FALSE)

#
# Check units for biol effects
#

# data_conc %>% 
#   filter(PARAM %in% c("PA1O", "PYR1O", "BAP3O", "PA1OH", "PYR1OH", "BAP3OH", "ALAD", "EROD")) %>%
#   count(PARAM, UNIT)
# 
# data_10_2016 %>% 
#   filter(PARAM %in% c("PA1O", "PYR1O", "BAP3O", "PA1OH", "PYR1OH", "BAP3OH", "ALAD", "EROD")) %>%
#   count(PARAM, MUNIT)

```


### c1. Fish, get SUBNO

#### c1a1. Definitions
```{r}
# Define fish species
fish_species <- c('Gadus morhua', 'Platichthys flesus', 'Limanda limanda')

# Define MATRX
df_matrx <- tibble(
  TISSUE_NAME = c("Muskel", "Blod", "Galle", "Liver - microsome", "Lever"),
  MATRX = c("MU", "BL", "BI", "LIMIC", "LI")
)

# xtabs(~MATRX + PARAM, data_10_2016 %>% filter(PARAM %in% c("ALAD", "BAP3OH", "EROD", "PA1OH", "PYR1OH")))
# xtabs(~MUNIT + PARAM, data_10_2016 %>% filter(PARAM %in% c("ALAD", "BAP3OH", "EROD", "PA1OH", "PYR1OH")))
```

#### c1a2. Check when SAMPLE_NO_Muscle =/= Fish_no   

This could potentially cause trouble, as biological effects follow Fish_no, not SAMPLE_NO_Muscle as assumed in c1a3.   
But this is no problem: only 2 cases, for 02B and 36B - which don't have biological effects   
```{r}

check <- data_04_fish %>%
  filter(Fish_no != SAMPLE_NO_Muscle)  
check


```


#### c1a3. Muscle part (not including biological effects)    
(biological effects includes EROD, which is in liver, but samples are numbered following specimen number)
- Join "SAMPLE_NO2" (from concentration data) = "SAMPLE_NO_Muscle" (in 'data_04_fish')
```{r}

sel_tissues <- "Muskel"
data_conc_fishmuscle <- data_conc %>%
  filter(LATIN_NAME %in% fish_species & TISSUE_NAME %in% sel_tissues)  %>%
  left_join(df_matrx, by = c("TISSUE_NAME", "MATRX"))

# Add SMPNO and SUBNO
df_for_join <- data_04_fish %>% 
  select(STATION_CODE, LATIN_NAME, SAMPLE_NO_Muscle, SMPNO, SUBNO) %>%
  filter(!is.na(SAMPLE_NO_Muscle))

n1 <- nrow(data_conc_fishmuscle)
data_conc_fishmuscle <- data_conc_fishmuscle %>%
  left_join(df_for_join, by = c("STATION_CODE", "LATIN_NAME", "SAMPLE_NO2" = "SAMPLE_NO_Muscle")) %>%
  mutate(MATRX = "MU")

n2 <- nrow(data_conc_fishmuscle)
if (n1 == n2){
  cat("Creating data_conc_fishmuscle - join OK\n")
} else {
  cat("Creating data_conc_fishmuscle - join NOT OK!\n")
}

# data_conc_fishmuscle %>% filter(STATION_CODE == "24B" & PARAM %in% "HG") %>% as.data.frame()
# data_conc_fishmuscle %>% filter(STATION_CODE == "23B" & PARAM %in% "EROD") %>% as.data.frame()

# data_conc_fishmuscle %>% 
#   filter(is.na(SMPNO)) %>% 
#   select(STATION_CODE, PARAM, SAMPLE_NO, SAMPLE_NO2, VALUE_WW)


# apply(is.na(data_conc_fishmuscle), 2, sum)

# Check SMPNO - lacking for one special case, where fish no 15 
#    has no ordinary muscle data, but has biol. effect data
not_ok <- is.na(data_conc_fishmuscle$SMPNO)
cat("Lacking SMPNO + SUBNO:", sum(not_ok), "records\n")  # 0

#
# If >0 records lack SMPno + SUBNO, we can skip the manual write
#

manual_write <- FALSE

if (manual_write){
  
  # If these are all biol. effects...
  lacking_subno_is_biol_effects <- 
    data_conc_fishmuscle$PARAM[not_ok] %in% c("PA1OH", "PYR1OH", "BAP3OH", "ALAD", "EROD")
  
  # ...we fix this special case
  if (mean(lacking_subno_is_biol_effects) == 1){
    data_conc_fishmuscle$SUBNO[not_ok] <- data_conc_fishmuscle$SAMPLE_NO2[not_ok]  # set SUBNO to SAMPLE_NO2
    data_conc_fishmuscle$SMPNO[not_ok] <- 1  # set SMPN to 1 (assumong this is cod)
  }
  
  # Otherwise:
  View(data_conc_fishmuscle[not_ok,], title = "Not ok")
  
}


if (sum(not_ok) == 0){
  cat("All OK\n")
} else {
  stop("Still lacking SMPNO + SUBNO!")
}

#
# In first try, this resukts in "Still lacking SMPNO + SUBNO!"
# This code shows that this is lacking for BLod, Gall Liver - microsome and WO for 
#  SAMPLE_NO 2,9, 13 and 14 in station 53B
# It also shows that this is because 'data_04_fish' doesn't contain these 
#  (because there are noe liver + muscel samples from these fish)

if (FALSE){
  
  df <- data_conc_fishmuscle[not_ok,] %>%
    arrange(SAMPLE_NO, PARAM)
  xtabs(~TISSUE_NAME + SAMPLE_NO + STATION_CODE, df)
  

  # Check data_04_fish
  data_04_fish %>%
    filter(STATION_CODE == "53B")  


}

# Check if problems




```

#### c1a4. Biological effects + length and weight  
(biological effects includes EROD, which is in liver, but samples are numbered as muscle)  
- Join "SAMPLE_NO2" (from concentration data) = "Fish_no" (in 'data_04_fish')   
- I.e., SAMPLE_NO2 follows numbered following specimen number - in contrast with muscle data (see previous chunk)  
```{r}

sel_tissues <- c("Blod", "Galle", "Liver - microsome", "WO")
data_conc_fishbiol <- data_conc %>%
  filter(LATIN_NAME %in% fish_species & 
           (TISSUE_NAME %in% sel_tissues | PARAM %in% "EROD"))  %>%
  left_join(df_matrx, by = c("TISSUE_NAME", "MATRX"))

# Add SMPNO and SUBNO
df_for_join <- data_04_fish %>% 
  select(STATION_CODE, LATIN_NAME, Fish_no, SMPNO, SUBNO) %>%
  filter(!is.na(Fish_no))

n1 <- nrow(data_conc_fishbiol)
data_conc_fishbiol <- data_conc_fishbiol %>%
  left_join(df_for_join, by = c("STATION_CODE", "LATIN_NAME", "SAMPLE_NO2" = "Fish_no")) %>%
  mutate(MATRX = "MU")

n2 <- nrow(data_conc_fishbiol)
if (n1 == n2){
  cat("Creating data_conc_fishbiol - join OK\n")
} else {
  cat("Creating data_conc_fishbiol - join NOT OK!\n")
}

check <- data_conc_fishbiol %>%
  filter(is.na(SUBNO))

if (nrow(check) > 0){
  check2 <- unique(check$STATION_CODE)
  check3 <- unique(check$SAMPLE_NO2)
  message("Lacking SUBNO for the following SAMPLE_NO: ", paste(check3, collapse = ","),
          "in STATION_CODE: ", paste(check2, collapse = ","),
       " (go back to Table 04.c to fix 'data_04_fish')")
  stop("'data_04_fish' must be fixed (error in Table 10 c1a4)\n")
}

# data_conc_fishbiol %>% filter(STATION_CODE == "24B" & PARAM %in% "HG") %>% as.data.frame()
# data_conc_fishbiol %>% filter(STATION_CODE == "23B" & PARAM %in% "EROD") %>% as.data.frame()

# data_conc_fishbiol %>% 
#   filter(is.na(SMPNO)) %>% 
#   select(STATION_CODE, PARAM, SAMPLE_NO, SAMPLE_NO2, VALUE_WW)


# apply(is.na(data_conc_fishbiol), 2, sum)

# Check SMPNO - lacking for one special case, where fish no 15 
not_ok <- is.na(data_conc_fishbiol$SMPNO)
cat("Lacking SMPNO + SUBNO:", sum(not_ok), "records\n")  # 0

#
# If >0 records lack SMPno + SUBNO, we can skip the manual write
#

manual_write <- FALSE

if (manual_write){
  
  # If these are all biol. effects...
  lacking_subno_is_biol_effects <- 
    data_conc_fishbiol$PARAM[not_ok] %in% c("PA1OH", "PYR1OH", "BAP3OH", "ALAD", "EROD")
  
  # ...we fix this special case
  if (mean(lacking_subno_is_biol_effects) == 1){
    data_conc_fishbiol$SUBNO[not_ok] <- data_conc_fishbiol$SAMPLE_NO2[not_ok]  # set SUBNO to SAMPLE_NO2
    data_conc_fishbiol$SMPNO[not_ok] <- 1  # set SMPN to 1 (assumong this is cod)
  }
  
}


if (sum(not_ok) == 0){
  cat("All OK\n")
} else {
  cat ("Still lacking SMPNO + SUBNO!")
}

#
# In first try, this resukts in "Still lacking SMPNO + SUBNO!"
# This code shows that this is lacking for BLod, Gall Liver - microsome and WO for 
#  SAMPLE_NO 2,9, 13 and 14 in station 53B
# It also shows that this is because 'data_04_fish' doesn't contain these 
#  (because there are noe liver + muscel samples from these fish)

if (FALSE){
  
  df <- data_conc_fishbiol[not_ok,] %>%
    arrange(SAMPLE_NO, PARAM)
  xtabs(~TISSUE_NAME + SAMPLE_NO + STATION_CODE, df)
  

  # Check data_04_fish
  data_04_fish %>%
    filter(STATION_CODE == "53B")  


}

# Check if problems




```


#### c1a5. Check 33F   

**Not needed **  


#### c1b. Liver part (EXCLUDING EROD, which is in liver but is numbered following specimen number)  
```{r}

data_conc_fishliver <- data_conc %>%
  filter(LATIN_NAME %in% fish_species & TISSUE_NAME %in% "Lever")

df_for_join <- data_04_fish %>% 
  select(STATION_CODE, LATIN_NAME, SAMPLE_NO_Liver, SMPNO, SUBNO) %>%
  filter(!is.na(SAMPLE_NO_Liver))

n1 <- nrow(data_conc_fishliver)
data_conc_fishliver <- data_conc_fishliver %>%
  left_join(df_for_join, by = c("STATION_CODE", "LATIN_NAME", "SAMPLE_NO2" = "SAMPLE_NO_Liver")) %>%
  mutate(MATRX = "LI") %>%
  filter(!PARAM %in% "EROD")   # exclude EROD
  
n2 <- nrow(data_conc_fishliver)
if (n1 == n2){
  cat("Join OK\n")
} else {
  cat("Join NOT OK\n")
}

# Check SMPNO
not_ok <- is.na(data_conc_fishliver$SMPNO)
cat("Number of SMPNO not ok: ", sum(not_ok))  # should be zero

# Check
# data_conc_fishmuscle[not_ok,]

if (sum(not_ok) != 0){
  data_conc_fishliver[not_ok,] %>%
    count(STATION_CODE, LATIN_NAME)
  
  df_liver %>% pull(STATION_CODE) %>% unique() %>% sort()
  # 19B, 28B, 43B2, 80B, 96B
  df_samples %>% pull(AQUAMONITOR_CODE) %>% unique() %>% sort()
  
  data_conc_fishliver[not_ok,] %>%
    count(STATION_CODE, LATIN_NAME, SAMPLE_NO2) %>% View("Liver samp")
  
}

# apply(is.na(data_conc_fishliver), 2, sum)

```

#### c1c. If some SMPNO lacking: Check data
```{r}

# data_04_fish %>% filter(STATION_CODE == "23B")
check <- FALSE

if (check){
  data_conc %>% 
    filter(STATION_CODE == "28B" & PARAM %in% "CB118") %>%
    arrange(SAMPLE_NO) %>%
    select(STATION_CODE, SAMPLE_NO)
  
  data_04_fish %>% 
    filter(STATION_CODE == "28B")
  
  data_04_fish_main %>% 
    filter(STATION_CODE == "28B")
  
}


```

#### c1d. Combine, add MUNIT and AMLNK   
- makes `data_conc_fish`  
```{r}

data_conc_fish <- rbind(data_conc_fishmuscle, data_conc_fishbiol, data_conc_fishliver)
# apply(is.na(data_conc_fish), 2, sum)

# Add MUNIT and AMLNK
data_conc_fish <- data_conc_fish %>%
  # add MUNIT
  safe_left_join(df_munit, by = "UNIT", 
                 na_matches = "never", check = "BCV") %>%
  # add AMLNK
  safe_left_join(df_amlnk, by = "PARAM", 
                 na_matches = "never", check = "BCV") %>%
  # Fix MUNIT for dry weight and fat weight  
  mutate(
    MUNIT = case_when(
      PARAM %in% c("DRYWT%", "FATWT%") ~ "%",
      TRUE ~ MUNIT)
    )

cat("'data_conc_fish' created with", nrow(data_conc_fish), "number of rows \n")

# apply(is.na(data_conc_fish), 2, sum)

#
# Unit for con biol. effects - NOT NEEDED ANYMORE, FIXED IN df_munit!
#
# sel <- data_conc_fish$PARAM %in% c("BAP3OH", "PA1OH", "PYR1OH"); sum(sel)
# data_conc_fish$MUNIT[sel] <- "ng/g"
# sel <- data_conc_fish$PARAM %in% c("ALAD"); sum(sel)
# data_conc_fish$MUNIT[sel] <- "ng/min/mg protein"
# sel <- data_conc_fish$PARAM %in% c("EROD"); sum(sel)
# data_conc_fish$MUNIT[sel] <- "pmol/min/mg protein"

#
# Basis for EROD
#  PARAM EROD and CYP1A should be reported in MATRX LIS9 or LIMIC (WGBEC2013)
#
# Info from Anders in mail 22 NOv 2018: "Microsomalfraksjonen lages ved å ultrasentrifugere S9-fraksjonen" 
#
sel <- data_conc_fish$PARAM %in% c("EROD"); sum(sel)
data_conc_fish$MATRX[sel] <- "LIS9"

```

### c2. Set uncertainty = 0 to NA  
No longer needed, we set UNCRT directly
```{r}
# sel <- !is.na(data_conc_fish$UNCERTAINTY) & data_conc_fish$UNCERTAINTY == 0
# table(sel)
# data_conc_fish$UNCERTAINTY[sel] <- NA

```

### c3. data_10, fish part
```{r}

data_10_fish <- data.frame(
  RECID = "10", 
  CRUIS = "AA711", 
  STNNO = data_conc_fish$STATION_CODE,
  SMPNO = data_conc_fish$SMPNO,
  SUBNO = data_conc_fish$SUBNO,
  MATRX = data_conc_fish$MATRX,
  DEPHU = NA,
  DEPHL = NA,
  PARAM = data_conc_fish$PARAM,
  MUNIT = data_conc_fish$MUNIT,
  BASIS = case_when(data_conc_fish$PARAM %in% c("DRYWT%", "FATWT%") ~ "",
                    TRUE ~ "W"),
  AMLNK = data_conc_fish$AMLNK,
  VFLAG = NA,
  QFLAG = data_conc_fish$FLAG1,
  VALUE = data_conc_fish$VALUE_WW,
  PERCR = NA,
  SIGND = NA,
  UNCRT = data_conc_fish$UNCRT, 
  METCU = data_conc_fish$METCU,
  DETLI = NA,
  LMQNT = NA,      #  could use data_conc_fish$QUANTIFICATION_LIMIT but unclear whether 
  stringsAsFactors = FALSE
  )

xtabs(~PARAM +  is.na(MATRX), data_conc_fish)
```

### c4. Set quantification limit = 1 for EROD
```{r}

sel <- data_10_fish$PARAM == "EROD"; sum(sel)
data_10_fish$DETLI[sel] <- 1

```

### c5. Checks
```{r}

cat("Table 10 - missing values\n")
missing <- apply(is.na(data_10_fish), 2, sum)
missing

critical1 <- c("RECID", "CRUIS", "STNNO", "SMPNO", "SUBNO")
critical2 <- c("MATRX", "PARAM", "MUNIT", "BASIS", "AMLNK", "VALUE", "UNCRT", "METCU")
if (sum(missing[critical1]) > 0)
  stop("Part Table 10.c5. Critical ID columns have missing values. Check 'data_conc_fish'")
if (sum(missing[critical2]) > 0)
  warning("Part Table 10.c5. Some important columns have missing values (e.g. AMLNK), but will be fixed later")
# This includes lacking AMLNK for HCHA, HCHG, HCB (see  Table 21.b3b)

# xtabs(~PARAM, data_10_fish)

cat("\n\nTable 10\n")
data_10_fish %>% 
  count(STNNO, MATRX, SUBNO) %>% 
  xtabs(~MATRX + STNNO, .)

cat("\n\nOriginal data\n")
data_conc %>%
  count(STATION_CODE, TISSUE_NAME, SAMPLE_NO) %>%
  xtabs(~TISSUE_NAME + STATION_CODE, .)

cat("\n\nMissing uncertainties  \n")
data_10_fish %>%
  filter(is.na(UNCRT)) %>%
  xtabs(~PARAM, .)

cat("\n\nMissing units  \n")
data_10_fish %>%
  filter(is.na(MUNIT)) %>%
  xtabs(~PARAM, .)

if (FALSE){
  
  cat("\n\nMissing values of VALUE:  \n")
  data_10_fish %>%
    filter(is.na(VALUE)) %>%
    View()
  
  data_conc_fish %>%
    filter(is.na(SUBNO)) %>%
    View("SUBNO missing")
  
  data_conc_fish %>%
    filter(is.na(AMLNK)) %>%
    View("AMLNK missing")
}


```
### c6. Check SMPNO
```{r}

xtabs(~is.na(SMPNO), data_10_fish)  
xtabs(~is.na(SMPNO), data_conc_fish)  
xtabs(~is.na(SMPNO), data_conc_fishmuscle)  
xtabs(~is.na(SMPNO), data_conc_fishliver)  

xtabs(~STATION_CODE + is.na(SMPNO), data_conc_fish)  

# xtabs(~PARAM + is.na(SMPNO), data_conc_fish)  

```

### c7. Check last year (2018)
```{r}

# fn1 <- "../Milkys_2018/ICES/Test/NIVA2017CF.NO"
# dat1 <- read_ices_file(fn1) %>%  add_field_codes()
# dat1[["10"]] %>% 
#   count(STNNO, MATRX, SUBNO) %>% 
#   xtabs(~MATRX + STNNO, .)

```

### d. Blue mussel  
Add SMPNO, MUNIT, AMLNK
```{r}
data_conc_bluemussel <- data_conc %>%
  filter(LATIN_NAME %in% 'Mytilus edulis') %>%
  left_join(df_smpno %>% select(LATIN_NAME, SMPNO), by = "LATIN_NAME") %>%
  left_join(df_munit, by = "UNIT") %>%
  left_join(df_amlnk, by = "PARAM") %>%
  # Fix MUNIT for dry weight and fat weight  
  mutate(
    MUNIT = case_when(
      PARAM %in% c("DRYWT%", "FATWT%") ~ "%",
      TRUE ~ MUNIT)
    )
  
```

### e1. Set uncertainty = 0 to NA  
No longer needed
```{r}
# sel <- !is.na(data_conc_bluemussel$UNCERTAINTY) & data_conc_bluemussel$UNCERTAINTY == 0
# table(sel)
# data_conc_bluemussel$UNCERTAINTY[sel] <- NA

```

### e2. data_10, blue mussel part
For blue mussel, we set SUBNO = SAMPLE_NO2
```{r}

data_10_bluemussel <- data.frame(
  RECID = "10", 
  CRUIS = "AA711", 
  STNNO = data_conc_bluemussel$STATION_CODE,
  SMPNO = data_conc_bluemussel$SMPNO,
  SUBNO = data_conc_bluemussel$SAMPLE_NO2,
  MATRX = "SB",
  DEPHU = NA,
  DEPHL = NA,
  PARAM = data_conc_bluemussel$PARAM,
  MUNIT = data_conc_bluemussel$MUNIT,
  BASIS = case_when(data_conc_bluemussel$PARAM %in% "DRYWT%" ~ "",
                    TRUE ~ "W"),
  AMLNK = data_conc_bluemussel$AMLNK,
  VFLAG = NA,
  QFLAG = data_conc_bluemussel$FLAG1,
  VALUE = data_conc_bluemussel$VALUE_WW,
  PERCR = NA,
  SIGND = NA,
  UNCRT = data_conc_bluemussel$UNCRT, 
  METCU = data_conc_bluemussel$METCU,
  DETLI = NA,
  LMQNT = NA,   # alternatively data_conc_bluemussel$QUANTIFICATION_LIMIT,
  stringsAsFactors = FALSE
  )

# xtabs(~PARAM + BASIS, data_10_bluemussel)
```


### f. Dry weight and fat
- Add dry weight and fat to the data  
- Dry weight and fat, add SUBNO
- Put on "table 10 form"
NOT NEEDED in 2019



### g. Length Ver. 2: Makes 'data_10_length2'  
Adds length and weight to the data using df_ind and data_04_fish
```{r}

df_fishlength <- data_04_fish %>%
  filter(!is.na(Fish_no)) %>%
  left_join(df_ind, by = c("STATION_CODE", "Fish_no")) %>%
  select(STATION_CODE, SMPNO, SUBNO, Length, Weight) %>%
  gather("Param", "VALUE", Length, Weight) %>%
  mutate(PARAM = case_when(Param == "Length" ~ "LNMEA",
                           Param == "Weight" ~ "WTMEA")) %>%
  mutate(MUNIT = case_when(Param == "Length" ~ "cm",
                           Param == "Weight" ~ "g")) %>%
  left_join(df_amlnk,  by = "PARAM")                 # Adds AMLNK


data_10_length2 <- data.frame(
  RECID = "10", 
  CRUIS = "AA711", 
  STNNO = df_fishlength$STATION_CODE,
  SMPNO = df_fishlength$SMPNO,
  SUBNO = df_fishlength$SUBNO,
  MATRX = "WO",
  DEPHU = NA,
  DEPHL = NA,
  PARAM = df_fishlength$PARAM,
  MUNIT = df_fishlength$MUNIT,
  BASIS = "",
  AMLNK = df_fishlength$AMLNK,
  VFLAG = NA,
  QFLAG = NA,
  VALUE = df_fishlength$VALUE,
  PERCR = NA,
  SIGND = NA,
  UNCRT = NA, 
  METCU = "",
  DETLI = NA,
  LMQNT = NA,
  stringsAsFactors = FALSE
  )
```


### h. Snail data: VDSI, TBSN+ and DRYWT%
```{r}

# data_conc %>% filter(PARAM %in% "TBSN+") %>% data.frame()
df_snail <- data_conc %>% 
  filter(LATIN_NAME %in% c("Nucella lapillus", "Littorina littorea")) %>% # xtabs(~PARAM, .)
  left_join(df_smpno) %>%
  left_join(df_amlnk) %>%
  mutate(SAMPLE_NO2 = 1) %>%     # Note - we just assume a single pooled sample from each snail station
  data.frame()
# df_snail %>% head(6)

data_10_vdsi <- data.frame(
  RECID = "10", 
  CRUIS = "AA711", 
  STNNO = df_snail$STATION_CODE,
  SMPNO = df_snail$SMPNO,
  SUBNO = df_snail$SAMPLE_NO2,
  MATRX = "SB",
  DEPHU = NA,
  DEPHL = NA,
  PARAM = df_snail$PARAM,
  UNIT = df_snail$UNIT,
  BASIS = case_when(df_snail$PARAM %in% c("DRYWT%", "VDSI", "INTF%") ~ "",
                    TRUE ~ "W"),
  AMLNK = df_snail$AMLNK,
  VFLAG = NA,
  QFLAG = df_snail$FLAG1,
  VALUE = df_snail$VALUE_WW,
  PERCR = NA,
  SIGND = NA,
  UNCRT = df_snail$UNCRT, 
  METCU = df_snail$METCU,
  DETLI = NA,
  LMQNT = NA,    # alternatively df_snail$QUANTIFICATION_LIMIT,
  stringsAsFactors = FALSE
  )

data_10_vdsi <- data_10_vdsi %>%
  left_join(df_munit) %>%
  mutate(
    MUNIT = case_when(
      PARAM %in% c("DRYWT%", "FATWT%") ~ "%",
      PARAM %in% "VDSI" ~ "IDX",
      TRUE ~ MUNIT)
  ) %>% 
  select(-UNIT)

cat("---------------------------------------------------------------\n")
cat("Parameters: \n---------------------------------------------------------------\n")
xtabs(~PARAM, data_10_vdsi)
cat("\n\n---------------------------------------------------------------\n")
cat("Units: \n---------------------------------------------------------------\n")
xtabs(~addNA(MUNIT), data_10_vdsi)

```

### i. Eider duck    
  
- SUBNO 16-30 should have EH - Egg homogenate of yolk and albumin    
- SUBNO 1-15 should have BL - Blood  
```{r}

# Use left join to data_04_eider to add SUBNO 
# df_eider <- 
df1 <- data_conc %>% 
  filter(LATIN_NAME %in% c("Somateria mollissima")) %>% # View()
  mutate(MATRX = case_when(
    grepl("Egg", TISSUE_NAME) ~ "EH",
    grepl("Blod", TISSUE_NAME) ~ "BL")
  ) %>%
  as.data.frame()
nrow(df1)
df2 <- df1 %>%
  left_join(df_smpno, by = "LATIN_NAME")    # adds SMPNO and SMLNK
df3 <- df2 %>%
  left_join(df_amlnk, by = "PARAM") %>%     # adds AMLNK
  left_join(df_munit, by = "UNIT")          # adds MUNIT
# Add SUBNO
data_conc_eider <- df3 %>%
  left_join(data_04_eider %>% select(LATIN_NAME, STATION_CODE, TISSUE_NAME, SAMPLE_NO2, SUBNO), 
            by = c("LATIN_NAME", "STATION_CODE", "TISSUE_NAME", "SAMPLE_NO2")
            )
# data_conc_eider %>%
#   select(LATIN_NAME, STATION_CODE, TISSUE_NAME, SAMPLE_NO2, SUBNO, PARAM, VALUE_WW) %>%
#   arrange(LATIN_NAME, STATION_CODE, PARAM, SUBNO) %>%
#   view()

if (nrow(df1) == nrow(data_conc_eider)){
  cat("Joins ok \n\n")
} else {
  cat("Joins NOT OK! \n\n")
}
# View(data_conc_eider)

data_10_eider <- data.frame(
  RECID = "10", 
  CRUIS = "AA711", 
  STNNO = data_conc_eider$STATION_CODE,
  SMPNO = data_conc_eider$SMPNO,
  SUBNO = data_conc_eider$SUBNO,
  MATRX = data_conc_eider$MATRX,
  DEPHU = NA,
  DEPHL = NA,
  PARAM = data_conc_eider$PARAM,
  MUNIT = data_conc_eider$MUNIT,
  BASIS = case_when(data_conc_eider$PARAM %in% "DRYWT%" ~ "",
                    TRUE ~ "W"),
  AMLNK = data_conc_eider$AMLNK,
  VFLAG = NA,
  QFLAG = data_conc_eider$FLAG1,
  VALUE = data_conc_eider$VALUE_WW,
  PERCR = NA,
  SIGND = NA,
  UNCRT = data_conc_eider$UNCRT, 
  METCU = data_conc_eider$METCU,
  DETLI = NA,
  LMQNT = NA,    # alternatively data_conc_eider$QUANTIFICATION_LIMIT
  stringsAsFactors = FALSE
  )

# xtabs(~PARAM + BASIS, data_10_bluemussel)
```


### j. Combine fish, blue mussel and fat weight/dryweight
```{r}

# Length version 1
# data_10 <- rbind(data_10_fish, data_10_bluemussel, data_10_fatdry, data_10_length, data_10_vdsi)

# Length version 2
# note: data_10_fatdry deleted
data_10 <- rbind(data_10_fish, data_10_bluemussel, data_10_length2, data_10_vdsi, data_10_eider)
  

# Round all values (5 decimals)
data_10$VALUE <- round(data_10$VALUE, 5)

# data_10 %>% filter(PARAM == "TBSN+") %>% View()
# data_10 %>% filter(PARAM == "VDSI") %>% View()

xtabs(~is.na(UNCRT), data_10)

```

### k1 - Fix some stuff
```{r}
# Delete records without value
sel <- is.na(data_10$VALUE); sum(sel)  
xtabs(~STNNO + PARAM, data_10[sel,])

data_10 <- data_10[!sel,]

# Set SMPNO to be numeric
data_10$SMPNO <- as.numeric(data_10$SMPNO)

# Change PFOSA (deprecated) to "PFOSD" (CAS 754-91-6)
# Description of PFOSA in ICES vocabulary:
#   Code: PFOSA
#   Description: 	Perfluorooctylsulfonate acid amide (deprecated)
#   Long Description: 	This is an old request which had no CAS number at the time. However, still 
#     cannot find CAS for perfluorooctylsulfonate acid amide in 2016. Is not a synonym of the 
#     commonly known PFOSA with CAS 754-91-6. Have to deprecate to clarify. See PFOSD.
# Description of PFOSD in ICES vocabulary: (seems that PFOSA has been misspelled PHOSA)
#   Code: 	PFOSD
#   Description: 	Perfluorooctanesulfonamide - branched + linear forms
#   Long Description: 	Perfluorooctanesulfonamide - Clarification that this is total, i.e., both 
#     forms, was added in 2018. This is the commonly known "PHOSA". ICES PHOSA has no CAS reference
#     and has been deprecated.

sel <- data_10$PARAM %in% "PFOSA"; sum(sel)
data_10$PARAM[sel] <- "PFOSD"
cat("PARAM changed from PFOSA to PFOSD for", sum(sel), "records\n")

# Check that all records have SMPNO
sum(is.na((data_10$SMPNO)))  # should be zero

# Set BASIS to "" for length, wet weigth etc. + VDSI
# xtabs(~PARAM, data_10)
sel <- data_10$PARAM %in% c("FATWT%", "DRYWT%", "LNMEA", "WTMEA", "VDSI"); sum(sel)
data_10$BASIS[sel] <- ""

```

### k2. Check why we lack SMPNO (if needed)
```{r}

if (sum(is.na(data_10$SMPNO)) > 0){
  # Which stations lack SMPNO?
  xtabs(~STNNO + is.na(SMPNO), data_10)
  
  # Which part of the data lack SMPNO? 
  xtabs(~is.na(SMPNO), data_10_fish)  
  xtabs(~is.na(SMPNO), data_10_bluemussel)  
  xtabs(~is.na(SMPNO), data_10_length2)  
  xtabs(~is.na(SMPNO), data_10_vdsi)  
} else {
  cat("SMPNO is OK\n")
}
```

### k3 - METCU may not be filled in when UNCRT is blank
```{r}

cat("Before: \n")
xtabs(~is.na(UNCRT) + addNA(METCU), data_10)

# Set 
sel <- is.na(data_10$UNCRT)
data_10$METCU[sel] <- NA

cat("\nAfter: \n")
xtabs(~is.na(UNCRT) + addNA(METCU), data_10)

```

### k4 - MATRX BL required for PARAM ALAD
```{r}
sel <- data_10$PARAM %in% "ALAD"
# table(addNA(data_10$MATRX[sel]))

data_10$MATRX[sel] <- "BL"
table(addNA(data_10$MATRX[sel]))

```

### k5 - GOSOI   
    - Effects parameters LISOI and GOSOI should report MUNIT index in the valid range of 0-50
    - Effects parameter GOSOI should be reported in MATRX GO
```{r}

sel <- data_10$PARAM %in% "GOSOI"

# Isn't this OK?
table(addNA(data_10$MUNIT[sel]))
summary(data_10$VALUE[sel])

# Fix MUNIT
table(addNA(data_10$MATRX[sel]))
data_10$MATRX[sel] <- "GO"

```

### k6 - PARAM Bile metabolites, fix    
    - PARAM Bile metabolites must be reported in matrix BI (bile) (WGBEC2013)  
    - Recommended MUNIT for Bile metabolites is nanogram/gram (WGBEC2013) (BUT written "ng/g")  
    - PARAM bile metabolites require reporting of DETLI (WGBEC2013)  
```{r}

sel <- data_10$PARAM %in% c("PA1O", "PYR1O", "BAP3O", "PA1OH", "PYR1OH", "BAP3OH")

# Fix MATRX
table(addNA(data_10$MATRX[sel]))
data_10$MATRX[sel] <- "BI"

# Fix MUNIT
table(addNA(data_10$MUNIT[sel]))
data_10$MUNIT[sel] <- "ng/g"

# DETLI is (SO FAR) set to zero
df_statistics <- data_10[sel,] %>%
  group_by(PARAM) %>%
  summarise(Min = min(VALUE), Median = median(VALUE), Max = max(VALUE),
            .groups = "drop")

# Add DETLI
data_10 <- data_10 %>%
  safe_left_join(df_statistics %>% select(PARAM, Min), 
                 na_matches = "never",
                 check = "V") %>%
  mutate(DETLI = case_when(!is.na(Min) ~ Min,
                           is.na(Min) ~ DETLI)
         ) %>%
  select(-Min)

# Change QFLAG from < to D
# But no QFLAGS for these, so don't have to worry about that

```


### k7. UNCRT for DRYWT% FATWT%  LNMEA  WTMEA    
Set to 10% (SD = 5%)
```{r}

# Det.limit - never there
table(is.na(data_conc$DETECTION_LIMIT))

# Lacking UNCRT:
#   DRYWT% FATWT%  LNMEA   PFBS  WTMEA
sel <- is.na(data_10$UNCRT)
xtabs(~PARAM, data_10[sel,])
xtabs(~PARAM, data_10[sel,]) %>% names() %>% dput()

# Set UNCRT
sel <- data_10$PARAM %in% c("DRYWT%", "FATWT%", "LNMEA", "WTMEA"); sum(sel)
data_10$UNCRT[sel] <- 10
data_10$METCU[sel] <- "%"

# View(data_10[sel,])

```


### k8. DETLI for DRYWT% FATWT%  LNMEA  WTMEA    
Set to 10% (SD = 5%)
```{r}

# Det.limit - never there
table(is.na(data_conc$DETECTION_LIMIT))

```

### k9a. For less-thans (QFLAG = <), set UNCRT = NA
```{r}

cat("Before: \n")
xtabs(~addNA(QFLAG) + addNA(METCU), data_10)

sel_lt <- data_10$QFLAG %in% "<"; sum(sel)
data_10$UNCRT[sel_lt] <- NA
data_10$METCU[sel_lt] <- NA

cat("\nAfter: \n")
xtabs(~addNA(QFLAG) + addNA(METCU), data_10)

```


### k9b. When UNCRT present, set METCU  
```{r}

# This is all old

# sel_uncrt <- !is.na(data_10$UNCRT)
# 
# 
# cat("UNCRT given vs less thans \n") 
# # (could also use sel_lt from previous :)
# table(sel_uncrt, sel_lt)
# 
# cat("\nCheck METCU \n") 
# table(addNA(data_10$METCU), sel_uncrt)
# 
# data_10$METCU[sel_uncrt] <- "SD"
# data_10$METCU[!sel_uncrt] <- NA
# 
# cat("\nCheck METCU again \n") 
# table(addNA(data_10$METCU), sel_uncrt)

```

### k9c. Set DETLI for PAH metabolites   

- Set to   

```{r}

sel <- data_10$PARAM %in% c("PA1OH", "PYR1OH", "BAP3OH")  
sum(sel)

data_10$DETLI[sel] <- 1

```

### k9c. Change QFLAG and set LMQNT  
- Change QFLAG from < to Q   
```{r}

sel_uncrt <- !is.na(data_10$QFLAG)
message(sum(sel_uncrt), " values are under LOQ - we set QFLAG and LMQNT for these")  

data_10$QFLAG[sel_uncrt] <- "Q"
data_10$LMQNT[sel_uncrt] <- data_10$VALUE[sel_uncrt]

```


### k10a. Detection limits in tab 10 and original data   
```{r}
xtabs(~is.na(DETLI), data_10)
# xtabs(~is.na(DETECTION_LIMIT), data_conc)


```
### k10b. Detection limits, metals in liver
```{r}

#
# Check metals in liver
#
# For getting metal parameters:
# pars <- data_10$PARAM %>% unique()
# pars[nchar(pars) == 2] %>% dput()

pars_metals <- c("HG", "SN", "CD", "AG", "CU", "CR", "AS", "PB", "CO", "NI", "ZN")

matrix <- "LI"
df <- data_10 %>%
  filter(PARAM %in% pars_metals & QFLAG %in% "Q" & MATRX %in% matrix) %>%
  ungroup() %>%
  group_by(PARAM) %>%
  summarise(
    loq_median = median(VALUE),
    loq_min =  min(VALUE),
    loq_max = max(VALUE)
    ) %>%
  arrange(loq_min)
df

#
# Set DETLI for metals in liver, based on table shown below
#
# note that "CU" "AS" "ZN" had no data under LOQ, but were se from looking at graphs...

matrix <- "LI"
sel <- with(data_10, MATRX %in% matrix & PARAM %in% c("CD","CO"))
data_10$DETLI[sel] <- 0.001
sel <- with(data_10, MATRX %in% matrix & PARAM %in% "PB")
data_10$DETLI[sel] <- 0.005
sel <- with(data_10, MATRX %in% matrix & PARAM %in% c("CR","NI","SN", "CU"))
data_10$DETLI[sel] <- 0.010
sel <- with(data_10, MATRX %in% matrix & PARAM %in% c("AG", "ZN", "AS"))
data_10$DETLI[sel] <- 0.050

# pars_metals[!pars_metals %in% df$PARAM]
# "HG" "CU" "AS" "ZN"

# For use in the next chink (blue mussel)
df_loq_metals <- bind_rows(
  df %>% select(PARAM, loq_min),
  data.frame(
    PARAM = c("CU","AS","ZN"),
    loq_min = c(0.010, 0.050, 0.050)
  ))
              

```

### k10c. Detection limits, metals in blue mussel
```{r}

matrix <- "SB"
df1 <- data_10 %>%
  filter(PARAM %in% pars_metals & QFLAG %in% "Q" & MATRX %in% matrix) %>%
  ungroup() %>%
  group_by(PARAM) %>%
  summarise(
    loq_median = median(VALUE),
    loq_min =  min(VALUE),
    loq_max = max(VALUE)
    ) 
df1

df2 <- data_10 %>%
  filter(PARAM %in% pars_metals & is.na(QFLAG) & MATRX %in% matrix) %>%
  ungroup() %>%
  group_by(PARAM) %>%
  summarise(
    val_median = median(VALUE),
    val_min =  min(VALUE)
    ) %>%
  arrange(val_min)
df2 %>%
  left_join(df_loq_metals)

#
# Conclusion: we use the same limits as we used for liver 
#
matrix <- "SB"
sel <- with(data_10, MATRX %in% matrix & PARAM %in% c("CD","CO"))
data_10$DETLI[sel] <- 0.001
sel <- with(data_10, MATRX %in% matrix & PARAM %in% "PB")
data_10$DETLI[sel] <- 0.005
sel <- with(data_10, MATRX %in% matrix & PARAM %in% c("CR","NI","SN", "CU"))
data_10$DETLI[sel] <- 0.010
sel <- with(data_10, MATRX %in% matrix & PARAM %in% c("AG", "ZN", "AS"))
data_10$DETLI[sel] <- 0.050


```

### k10d. Detection limits, mercury in muscle
```{r}

matrix <- "MU"
df1 <- data_10 %>%
  filter(PARAM %in% pars_metals & QFLAG %in% "Q" & MATRX %in% matrix) %>%
  ungroup() %>%
  group_by(PARAM) %>%
  summarise(
    loq_median = median(VALUE),
    loq_min =  min(VALUE),
    loq_max = max(VALUE)
    )  %>%
  arrange(loq_min)
df1

df2 <- data_10 %>%
  filter(PARAM %in% pars_metals & is.na(QFLAG) & MATRX %in% matrix) %>%
  ungroup() %>%
  group_by(PARAM) %>%
  summarise(
    val_median = median(VALUE),
    val_min =  min(VALUE)
    ) %>%
  arrange(val_min)
df2

matrix <- "MU"
sel <- with(data_10, QFLAG %in% "Q" & MATRX %in% matrix & PARAM %in% c("CD","CO"))

```

### k11. Set MUNIT for FAT% and test for remaining missing units    
```{r}

sel <- with(data_10, PARAM %in% c("DRYWT%","FATWT%") & !MUNIT %in% "%")
data_10$MUNIT[sel] <- "%"

cat("MUNIT set for", sum(sel), "DRYWT% records.")

# Missing units
cat("\n\nMissing units: \n----------------------------------------------------------\n")
check <- data_10 %>%
  filter(is.na(MUNIT)) 
cat(nrow(check), "records lack MUNIT \n")

if (nrow(check) > 0){
  cat("Missing MUNIT - Parameters: \n")
  xtabs(~PARAM, check)
  stop("All records must have MUNIT! (Table 10 k11)")
}


```

### k12. Check 'VALUE <= DETLI but QFLAG is not specified'    
(As it is called in the DATSU check)  
```{r}

check <- data_10 %>%
  filter(VALUE <= DETLI) 
# nrow(check)

# All are PAH metabolites  
xtabs(~PARAM + is.na(QFLAG), check)

# Change DETLI to 90% of VALUE for these cases  
data_10 <- data_10 %>%
  mutate(DETLI = case_when(
    VALUE <= DETLI & is.na(QFLAG) ~ 0.9*VALUE,
    TRUE ~ DETLI)
  )

```

### k13. Check BASIS for GOSOI measurements     
DATSU: 'BASIS is for contaminant and chemical analysis only'  
```{r}

xtabs(~addNA(BASIS), data_10)


# Set BASIS to 'not available' for these cases  
data_10 <- data_10 %>%
  mutate(BASIS = case_when(
    PARAM %in% "GOSOI" ~ as.character(NA),
    TRUE ~ BASIS)
  )

```
### k14. Check type of error (METCU)      
DATSU: 'BASIS is for contaminant and chemical analysis only'  
```{r}

xtabs(~addNA(METCU), data_10)


# Set BASIS to 'not available' for these cases  
data_10 <- data_10 %>%
  mutate(BASIS = case_when(
    PARAM %in% "GOSOI" ~ as.character(NA),
    TRUE ~ BASIS)
  )

```


## Table 21 - analytic methods

### a. Check PARAM relative to table 10
```{r}
# data_21 <- data_21_2016  # restore from "backup"
data_21_back <- data_21  # make backup

# We use 2018 data instead
data_21 <- data_2018[["21"]] %>%
  mutate(AMLNK = as.numeric(AMLNK))

column_order_21 <- names(data_21)

```


### b1. Check if some parameters in data_10 lack AMLNK  
```{r}

check1 <- data_10 %>%
  count(PARAM, AMLNK) %>%
  filter(is.na(AMLNK))

check1

```

### b2. Check if some parameters in data_10 are lacking a data_21 entry  
```{r}

check2 <- data_10 %>%
  count(PARAM, AMLNK) %>%
  left_join(data_21 %>% select(AMLNK, ALABO), by = "AMLNK") %>% 
  filter(is.na(ALABO))

check2

```

### b3a. Modify AMLNK in data_10   
So the AMLNK for TBSN+ in data_10 points to the TBT line in data_21  
```{r}

if ("TBSN+" %in% check1$PARAM) {
  
  # First, find AMLNK for TBTIN (note that we use 'old' data, as these have been deleted from 'data_10'....)
  amlnk <- data_2018[["10"]] %>% 
    filter(PARAM == "TBTIN") %>% 
    pull(AMLNK) %>% unique()
  amlnk
  
  # Then, use this in the TBSN+ rows of table 10
  data_10 <- data_10 %>%
    mutate(AMLNK = case_when(
      PARAM %in% "TBSN+" ~ as.numeric(amlnk),  # use the AMLNK value we found...
      TRUE ~ AMLNK)                            # ...else, use existing AMLNK                         
    )
  
  cat("Modified AMLNK in data_10 so AMLNK for TBSN+ in data_10 points to the TBT line in data_21 ")
  
}


```
### b3b. Add extra parameters  
Must both add a new AMLNK to table 10 and add a line with this AMLNK to table 21
```{r}

#
# Functions
#

table_10_amlnk_get <- function(existing_data_10, param){
  max(existing_data_10$AMLNK, na.rm = TRUE) + 1
}

table_10_amlnk_set <- function(existing_data_10, param, amlnk){
  sel <- existing_data_10$PARAM %in% param
  if (sum(!is.na(existing_data_10$AMLNK[sel])) > 0){
    stop("AMLNK is given for this PARAM in table 10 (Table 21 b3b)")
  }
  existing_data_10$AMLNK[sel] <- amlnk
  cat("AMLNK set to", amlnk, "for", sum(sel), param, "records \n")
  existing_data_10
}

table_21_add <- function(existing_data_21, amlnk){
  if (amlnk %in% existing_data_21$AMLNK){
    stop("This AMLNK already exists in table 21 (Table 21 b3b)")
  }
  cat(length(amlnk), "row with AMLNK", amlnk, "added to table 21 \n")
  bind_rows(
    existing_data_21,
    tibble(
      RECID = "21",
      AMLNK = amlnk,
      ALABO = "NIVA",
      METFP = "NON")
  )
} 

#
# DO IT
#

amlnk <- data_10 %>% table_10_amlnk_get("HCHA")
data_10 <- data_10 %>% table_10_amlnk_set("HCHA", amlnk)
data_21 <- data_21 %>% table_21_add(amlnk)

cat("\n")
amlnk <- data_10 %>% table_10_amlnk_get("HCHG")
data_10 <- data_10 %>% table_10_amlnk_set("HCHG", amlnk)
data_21 <- data_21 %>% table_21_add(amlnk)

cat("\n")
amlnk <- data_10 %>% table_10_amlnk_get("HCB")
data_10 <- data_10 %>% table_10_amlnk_set("HCB", amlnk)
data_21 <- data_21 %>% table_21_add(amlnk)



```


### b4. Repeat checks (same as in b1 and b2)  
```{r}

# Check if some parameters in data_10 lack AMLNK  
check1 <- data_10 %>%
  count(PARAM, AMLNK) %>%
  filter(is.na(AMLNK))


# Check if some parameters in data_10 are lacking a data_21 entry  
check2 <- data_10 %>%
  count(PARAM, AMLNK) %>%
  left_join(data_21 %>% select(AMLNK, ALABO), by = "AMLNK") %>% 
  filter(is.na(ALABO))

if (nrow(check1) == 0 & nrow(check2) == 0){
  cat("All data_10 data have AMLNK, and all point to a line in data_21")
} else {
  stop("Either some data_10 data lack AMLNK, or some AMLNK point to a non-existing line in data_21. You must edit one of them. (Error in section 'Table 21 b4')")
}

```
### b5. Find records in table 21 tha are not used  
```{r}
# data_21 %>% colnames() %>% dput()
amlnk_missing <- data_21 %>% filter(!AMLNK %in% data_10$AMLNK) %>% pull(AMLNK)
param_in_21_but_not_in_10 <- data_10_2016 %>%
  filter(AMLNK %in% amlnk_missing) %>%
  group_by(PARAM, AMLNK) %>%
  summarise(N = n(), .groups = "drop")

cat("Parameters in data_21 that are not in data_10: \n")
param_in_21_but_not_in_10 %>% pull(PARAM)
#  "TBBPA" "TBP"
```

### b6. Those records are deleted from table 21
```{r}

data_21 <- data_21 %>% 
  filter(!AMLNK %in% param_in_21_but_not_in_10$AMLNK)

cat(length(param_in_21_but_not_in_10$AMLNK), "lines deleted from data_21")

```


### c. Add PARAM and MUNIT (will be removed afterwards)  
These are fetched from data_10 
```{r}

# Backup (for debugging)
# data_21_back <- data_21  # make backup
# data_21 <- data_21_back  # restore from backup

# xtabs(~addNA(MUNIT), data_10)

# Add PARAM (will be removed afterwards)
n1 <- nrow(data_21)
data_21 <- data_21 %>%
  safe_left_join(data_10 %>% count(PARAM, AMLNK, MUNIT) %>% select(-n), 
            by = "AMLNK",
            na_matches = "never",
            check = "BCV")
n2 <- nrow(data_21)

if (FALSE){
  
  amlnk_dubl <- data_10 %>% distinct(PARAM, AMLNK, MUNIT) %>% count(AMLNK) %>% filter(n > 1) %>% pull(AMLNK)
  data_10 %>% count(PARAM, AMLNK, MUNIT) %>% filter(AMLNK == amlnk_dubl)

  amlnk_dubl <- data_10 %>% count(PARAM, AMLNK, MUNIT) %>% count(AMLNK) %>% filter(n > 1) %>% pull(AMLNK)
  data_10 %>% count(PARAM, AMLNK, MUNIT) %>% filter(AMLNK == amlnk_dubl)

  data_conc %>%
    filter(PARAM == "HG") %>%
    count(LATIN_NAME, UNIT)

  data_conc %>%
    filter(PARAM == "HG") %>%
    group_by(LATIN_NAME, TISSUE_NAME, UNIT) %>%
    summarise(min(VALUE_WW), median(VALUE_WW), max(VALUE_WW))
  # If warning - check AMLNK
  data_10 %>%
    count(PARAM, AMLNK, MUNIT) %>%
    count(AMLNK) %>%
    filter(n > 2)
}


# Check 
cat("Number of AMLNK without a corresponding PARAM (should be 0):", sum(is.na(data_21$PARAM)), "\n")

# xtabs(~PARAM + MUNIT, data_10)
# xtabs(~PARAM + MUNIT, data_21)
# data_21 <- data_21 %>% select(-PARAM, -MUNIT)    # remove these variables (if error)
```

### d. Check existing METOA codes  
For biol. effects only
```{r}
xtabs(~METOA, data_21)
# xtabs(~PARAM + METOA, data_21 %>% filter(METOA != ""))
```



### e1. Create 'ICES_METOA_codes2.xlsx'  
```{r}
#
# This was used to create 'ICES_METOA_codes2.xlsx' (based on 'ICES_METOA_codes.xlsx')
# Also some manual additions to that file was done (length, wight, VDSI etc.) so don't repeat this code
#

# # Methods - made using script "03extra01....R"
# df_metoa <- readxl::read_excel("Input_data/ICES_METOA_codes.xlsx")
# df_metoa$PARAM <- replace_parameter_synonyms(df_metoa$NAME)
# 
# # From 10(a) above
# # Change BDE to BD for the following. NOTE: "BDE28"  "BDE47"  "BDE99" should NOT be changed (!)
# sel <- df_metoa$PARAM %in% c("BDE100", "BDE126", "BDE153", "BDE154", "BDE183", "BDE209"); sum(sel)
# df_metoa$PARAM[sel] <- sub("BDE", "BD", df_metoa$PARAM[sel])
# 
# # Change some of the rest
# sel <- df_metoa$PARAM %in% "PFUdA"; sum(sel)
# df_metoa$PARAM[sel] <- "PFUnda"
# sel <- df_metoa$PARAM %in% "PFDcA"; sum(sel)
# df_metoa$PARAM[sel] <- "PFDA"
# sel <- df_metoa$PARAM %in% "TBT"; sum(sel)
# df_metoa$PARAM[sel] <- "TBSN+"
# 
# # Pick the first of these (simple way to handle duplicates...)
# pars <- c("AG", "AS", "CB118", "CD", "CO", "CR", "CU", "NI", "PB", "SN", "TDEPP", "ZN")
# df_metoa <- bind_rows(
#   df_metoa %>% filter(!PARAM %in% pars),
#   df_metoa %>% filter(PARAM %in% pars) %>% group_by(PARAM) %>% summarise_all(funs(first))
#   )
# 
# df_metoa %>% select(METHOD_ID, NAME, PARAM, everything()) %>% select(-PARAM_ICES) %>%
#   openxlsx::write.xlsx("Input_data/ICES_METOA_codes2.xlsx")
```

### e2. Read method data   
```{r}

df_metoa <- readxl::read_excel("../Milkys_2018/Input_data/ICES_METOA_codes2.xlsx")

# Change PFOSA to PFOSD in the method data
sel <- df_metoa$PARAM %in% "PFOSA"; sum(sel)
df_metoa$PARAM[sel] <- "PFOSD"


```

### f. Rename original METOA and add new ones  
By join with df_metoa. Note that both PARAM and UNIT are used for the join!  
```{r}
# data_21_back <- data_21  # make backup
# data_21 <- data_21_2016  # restore from backup

n1 <- nrow(data_21)
data_21 <- data_21 %>%
  rename(METOA_orig = METOA) %>%
  left_join(df_metoa %>% select(PARAM, UNIT_ICES, METOA), by = c("PARAM", "MUNIT" = "UNIT_ICES"))
n2 <- nrow(data_21)
if (n2 != n1)
  warning("W A R N I N G !  LEFT JOIN CHANGED THE NUMBER OF LINES!")

# table(is.na(data_21$METOA_orig))

sel_lacking <- with(data_21, is.na(METOA) & (is.na(METOA_orig) | METOA_orig %in% ""))
# Should be zero
cat("Number of parameters without METOA or METOA_orig (should be zero):", sum(sel_lacking), "\n")
# data_21[sel_lacking,] %>% select(PARAM, MUNIT, METOA, METOA_orig)

# Where we lack METOA, we use METOA_orig
sel <- is.na(data_21$METOA)
data_21$METOA[sel] <- data_21$METOA_orig[sel]
cat("Number of codes gotten from METOA from last year:", sum(sel), "\n")

# Check
lacking_metoa <- is.na(data_21$METOA) | data_21$METOA %in% ""

# Should be zero
cat("Number of parameters without METOA (should be zero):", sum(lacking_metoa), "\n")

if (sum(lacking_metoa) > 0){
  df <- data_21[lacking_metoa,]
  df
  data_10 %>% filter(AMLNK %in% df$AMLNK) %>% pull(PARAM) %>% unique()
}



```

### Add METOA for GOSOI, FATWT and VDSI
```{r}
sel <- data_21$PARAM %in% "VDSI"; sum(sel)
data_21$METOA[sel] <- "VISO"  # Visual observation/identification

sel <- data_21$PARAM %in% c("GOSOI", "FATWT%"); sum(sel)
data_21$METOA[sel] <- "GRV"

```


```{r}
# sel <- data_21$PARAM %in% "PFOSA"; sum(sel)
# data_21$PARAM[sel] <- "Perfluorooctanesulfonic acid amide"
# cat("PARAM changed from PFOSA to Perfluorooctanesulfonic acid amide for", sum(sel), "records\n")

```


### Pick correct variables in correct order  
```{r}
vars <- c("RECID", "AMLNK", "ALABO", "METDC", "REFSK", "METST", "METFP", 
  "METPT", "METCX", "METPS", "METOA", "AGDET", "SREFW", "SPECI", 
  "RLIST", "ORGSP", "SIZRF", "FORML")
data_21 <- data_21[vars]
```




## Table 20 - sampling methods  
Just 3 lines - we will just use last year's table  
Note that the eider duck sampling (SMTYP) we will denote as 'HAN' = "by hand'

```{r}

# This is as follows:

# RECID SLABO SMLNK SMTYP NETOP MESHS SAREA LNSMB SPEED PDMET SPLIT Line_no
# 20;NIVA;11;HAN;;;;;;;
# 20;NIVA;9;GIL;;;;;;;
# 20;NIVA;8;BOT;;;;;;;

# Change to (vode below):

# 20;NIVA;11;HAN;;;;;;;
# 20;NIVA;8;GIL;;;;;;;

data_20 <- structure(
  list(
    RECID = c("20", "20", "20"), 
    SLABO = c("NIVA", "NIVA", "NIVA"), 
    SMLNK = c("11", "9", "8"), 
    SMTYP = c("HAN", "GIL", "BOT"), 
    NETOP = c("", "", ""), MESHS = c("", "", ""), SAREA = c("", "", ""), 
    LNSMB = c("", "", ""), SPEED = c("", "", ""), PDMET = c("", "", ""), 
    SPLIT = c("", "", ""), 
    row.names = c(NA, 
-3L), class = "data.frame")

```


## Table 91 - stations  
We will base ourself on using last year's table  
It lacks  
    + 19B (Svalbard)
    + 28A2 Ålesund havn
    + 97A3 Bodø havn
      
**NOTE:** A check of the data in script 26 shows that **97A3 (Bodø havn)** is still lacking in table 91. See `Start making data_91` below.  

### a. Get 2018 data (submitted in 2019)  
```{r}

data_91 <- data_2018[["91"]]
column_order_91 <- names(data_91)

```
    
  
### b. Check lacking stations  
```{r}

sel <- data_03$STNNO %in% data_91$STNNO
cat("\nTable 03 stations not in table 91: \n")
data_03[!sel,] %>% pull(STNNO)

sel <- data_10$STNNO %in% data_91$STNNO
cat("\nTable 10 stations not in table 91: \n")
data_10[!sel,] %>% pull(STNNO) %>% unique()

```



### c1. Add 97A3 (the actual Bodø harbour)     
```{r}

new_stations <- "97A3"
new_lat <- 67.2963
new_lon <- 14.3956

data_91_extra <- tibble(
  RECID = "91",
  CRUIS = "AA711",
  STNNO = new_stations, 
  LATIT = decimal2minute(new_lat),   # change to ICES format ("+67 17.778")
  LONGI = decimal2minute(new_lon),   # change to ICES format
  POSYS = "EST",
  SDATE = "",   # set later
  STATN = "97A3 Bodø harbour",
  MPROG = "CEMP",
  WLTYP = "CF", 
  MSTAT = "IH",       # impacted
  PURPM = "T~S"
)

check <- new_stations %in% data_91$STNNO

if (sum(check) == 0){
  data_91 <- bind_rows(data_91, data_91_extra)
  cat(nrow(data_91_extra), "new stations added\n")
} else if (sum(check) == 1){
  cat("Stations were already in data set\n")
} else {
  cat("Some, but not all, stations are already in data set. CHECK.\n")
}

```
### c2. Add 71A and 20B  

- 71A = Bjørkøya, Grenland   
- 20B = Longyearbyen  
```{r}

# data_2018[["91"]] %>% View()

new_stations <- c("71A", "20B")
new_lat <- c(59.02333, 78.2760)	
new_lon <- c(9.75367, 15.4060)

data_91_extra <- tibble(
  RECID = "91",
  CRUIS = "AA711",
  STNNO = new_stations, 
  LATIT = decimal2minute(new_lat),   # change to ICES format ("+67 17.778")
  LONGI = decimal2minute(new_lon),   # change to ICES format
  POSYS = "EST",
  SDATE = "",   # set later
  STATN = "71A Bjørkøya (Risøyodden)",
  MPROG = "CEMP",
  WLTYP = "CF", 
  MSTAT = "RH",      # Representative of general conditions in terms of hazardous substances
  PURPM = "T~S"
)

check <- new_stations %in% data_91$STNNO

if (sum(check) == 0){
  data_91 <- bind_rows(data_91, data_91_extra)
  cat(nrow(data_91_extra), "new stations added\n")
} else if (sum(check) == 1){
  cat("Stations were already in data set\n")
} else {
  cat("Some, but not all, stations are already in data set. CHECK.\n")
}

```
### c3. Check if we have all stations  
```{r}

check <- setdiff(unique(data_10$STNNO), data_91$STNNO)

if (length(check) > 0)
  stop("Lacking station(s)", paste(check, collapse = ", "))

```


### d. Remove stations not needed  
```{r}

sel <- data_91$STNNO %in% data_03$STNNO  
cat("\nStations not in table 03: \n")
data_91[!sel,] %>% pull(STNNO)

sel <- data_91$STNNO %in% unique(data_10$STNNO)  
cat("\nStations not in table 10: \n")
data_91[!sel,] %>% pull(STNNO)

# Remove those data lines nor found in table 10  
data_91 <- data_91[sel,]
cat("\n", sum(!sel), "stations deleted from table 91 \n")

```

### e. Fix coordinates of station 76A2  
ICES station dictionary: "Latitude, Longitude (58.7327 9.28104) Latitude range +/-, Longitude range +/- (0.00718, 0.01381)"  
- This also fits with "Kartbase.xlsx"  
```{r}

station <- "76A2" 
sel <- with(data_91, STNNO %in% station)

cat("Existing coordinates for", station, ":", 
    minute2decimal(data_91$LATIT[sel]), ",", minute2decimal(data_91$LONGI[sel]), "\n")

data_91$LONGI[sel] <- decimal2minute(9.28104)

cat("New coordinates for", station, ":", 
    minute2decimal(data_91$LATIT[sel]), ",", minute2decimal(data_91$LONGI[sel]), "\n")


```


### f. Sample dates   
- Taken from raw data, 
```{r}

data_for_dates <- data_ind2 %>%
  filter(!is.na(SAMPLE_DATE)) %>%
  distinct(STATION_CODE, SAMPLE_DATE)

df_dates <- data_for_dates %>%
  rename(STNNO = STATION_CODE)

# Check stations with >1 date (19N has two)
check_dates_1 <- df_dates %>%
  group_by(STNNO) %>%
  mutate(n = n()) %>%
  filter(n > 1) %>%
  arrange(STNNO, SAMPLE_DATE)

# If some stations with >1 date:
if (nrow(check_dates_1) > 0){
  message("Some stations have >1 date")
  
  # Mean of dates for 19N
  df_dates <- df_dates %>%
    group_by(STNNO) %>%
    summarise(
      SAMPLE_DATE = mean(SAMPLE_DATE) %>% floor_date("day"),
      .groups = "drop") %>%
    mutate(
      Date = as.character(SAMPLE_DATE),
      SDATE = paste0(substr(Date,1,4),substr(Date,6,7),substr(Date,9,10))
    )
  
  # df_dates %>% filter(STNNO == "19N")
  # df_dates
  
  # Check again stations with >1 date
  check_dates_2 <- df_dates %>%
    group_by(STNNO) %>%
    mutate(n = n()) %>%
    filter(n > 1) %>%
    arrange(STNNO, SAMPLE_DATE)
  
  if (nrow(check_dates_2) == 0){
    warning("Mean date used for stations with more than have >1 date. Check 'check_dates_1'. \n")
  } else {
    stop("Some stations with more than have >1 date!")
  }
  
}
cat("\n")

# Check if we have all dates we need
sel <- data_91$STNNO %in% df_dates$STNNO  
if (sum(!sel) > 0){
  stop(sum(!sel), " stations not found in 'df_dates'!")
} else {
  message("All stations are found in 'df_dates'")
}

```

### g. Add new dates
```{r}


data_91 <- data_91 %>%
  # Remove old date
  select(-SDATE) %>%
  # Add SAMPLE_DATE
  left_join(df_dates %>% select(STNNO, SDATE),  by = "STNNO")

# Correct column order
data_91 <- data_91[column_order_91]



```



## Table 90 - list of cruises  

### New table 90  
Original table is just 1 line - we will just use last year's table  
REVISION: Use
* SHIPC = AA36 unspecified fishing vessel for all fish stations except 30B  
* SHIPC = 58TB 	TRYGVE BRAARUD for 30B
```{r}
# Original
data_90

data_90 <- tibble(
  RECID = rep(90,3),
  SHIPC = c("AA36", "AA71", "58TB"),
  CRUIS = c(1,2,3),
  OWNER = NA,
  PRDAT = NA
)

data_90

```

### Also change CRUIS in other tables  
I.e.:  "data_90" "data_91" "data_03" "data_04" "data_10"
```{r}
# data_03

data_03 <- data_03 %>%
  mutate(CRUIS = case_when(
    grepl("B", STNNO) & STNNO != "30B" ~ 1,
    STNNO %in% "30B" ~ 3,
    grepl("F", STNNO) ~ 1,
    TRUE ~ 2)
  )

# For checking
if (FALSE){
  for (cr in unique(data_03$CRUIS)){
    cat(cr, "\n")
    print(data_03 %>% filter(CRUIS == cr) %>% pull(STNNO) %>% unique())
  }
}

data_04 <- data_04 %>%
  mutate(CRUIS = case_when(
    grepl("B", STNNO) & STNNO != "30B" ~ 1,
    STNNO %in% "30B" ~ 3,
    grepl("F", STNNO) ~ 1,
    TRUE ~ 2)
  )
    
    
data_10 <- data_10 %>%
  mutate(CRUIS = case_when(
    grepl("B", STNNO) & STNNO != "30B" ~ 1,
    STNNO %in% "30B" ~ 3,
    grepl("F", STNNO) ~ 1,
    TRUE ~ 2)
  )
    

data_91 <- data_91 %>%
  mutate(CRUIS = case_when(
    grepl("B", STNNO) & STNNO != "30B" ~ 1,
    STNNO %in% "30B" ~ 3,
    grepl("F", STNNO) ~ 1,
    TRUE ~ 2)
  )
    
# data_92 - fixed below

# For checking table 10
if (FALSE){
  for (cr in 1:3){
    cat(cr, "\n")
    print(data_10 %>% filter(CRUIS == cr) %>% pull(STNNO) %>% unique())
  }
}

```






## Table 92 - station info   
Only temp for 23B, 30B, 53B   
Line 21, PARAMs EROD and CYP1A require the reporting of bottom temperature (BotTemp) in record 92 (WGBEC2013)
```{r}
data_92 <- tibble(
  RECID = "92",
  CRUIS = c(1,3,1),          # See table 90
  STNNO = c("23B","30B","53B"),
  RSRVD = NA,
  MATRX = "SI",
  PARAM = "BotTemp",
  MUNIT = "degC",
  VALUE = 7.75,
  DCFLG = NA
  )
```



## Finishing


### Check 1 - duplicates in 04
```{r}
# Are there duplicates in table 04? 
check <- data_04 %>%
    group_by(CRUIS, STNNO, SMPNO, SUBNO) %>%
    mutate(n = n()) %>%
    filter(n > 1) %>%
    arrange(CRUIS, STNNO, SMPNO, SUBNO) %>%
    select(RECID:NOIMP, SUBNO, NOIMP, BULKID, everything())
  
cat("Number of duplicates in table 04: ", nrow(check), "\n")

if (nrow(check) > 0){

  cat("Please remove the duplicates.\n\n")

  tab <- xtabs(~SUBNO + STNNO, check)
  tab
  
  # check

}

if (FALSE){

  data_04_fish %>%
    group_by(STATION_CODE, SMPNO, SUBNO) %>%
    mutate(n = n()) %>%
    filter(n > 1) %>%
    arrange(STATION_CODE, SMPNO, SUBNO)
  
  }

```


### Check 2 - duplicates in 10
```{r}
# Are there duplicates in table 10? 
check <- data_10 %>%
  group_by(CRUIS, STNNO, SMPNO, SUBNO, MATRX, PARAM) %>%
  mutate(n = n()) %>%
  filter(n > 1) %>%
  arrange(CRUIS, STNNO, SMPNO, SUBNO, MATRX, PARAM) %>%
  select(CRUIS, STNNO, SMPNO, SUBNO, MATRX, PARAM, VALUE, everything())

cat("Number of duplicates in table 10: ", nrow(check), "\n")

if (nrow(check) > 0){

  cat("Please remove the duplicates.\n\n")

  tab <- xtabs(~PARAM + STNNO, check)
  cat("Parameters: \n-----------------------------\n")
  apply(tab, 1, sum) %>% sort(decreasing = TRUE) %>% print()
  cat("Stations: \n-----------------------------\n")
  apply(tab, 2, sum) %>% sort(decreasing = TRUE) %>% print()
  cat("Stations / parameters: \n-----------------------------\n")
  tab2 <- cbind(tab, Sum = apply(tab, 1, sum)) 
  tab2 <- tab2[order(tab2[,"Sum"]),]
  tab2
  
}

```


### Check 3 - join 10 + 91
```{r}
#
# Stations: Measurement table vs station table
#
check <- data_10 %>%
  safe_left_join(data_91, by = c("CRUIS", "STNNO"), 
                 na_matches = "never", check = "VMN")

cat("Checked measurement table (10) vs station table (91). If no error, check has passed.")

# If lack of matches, do this:
if (FALSE){
  df <- full_join(
    tibble(STNNO = unique(data_10$STNNO), tab10 = 1),
    tibble(STNNO = unique(data_91$STNNO), tab91 = 1)
  ) %>%
    filter(is.na(tab10) | is.na(tab91))
  df
}
```


### Check 4 - join 10 + 04
```{r}

#
# Stations and sample numbers: Measurement table vs sample table
#
check <- data_10 %>%
  safe_left_join(data_04, by = c("CRUIS", "STNNO", "SMPNO", "SUBNO"), 
                 na_matches = "never", check = "VMN")

cat("Checked measurement table (10) vs sample table (04). If no error, check has passed.")

# If errors, do this:
if (FALSE){
  
  # If  y is not unique on CRUIS, STNNO, SMPNO and SUBNO, do this:
  data_04 %>%
    group_by(CRUIS, STNNO, SMPNO, SUBNO) %>%
    mutate(n = n()) %>%
    filter(n > 1) %>%
    arrange(CRUIS, STNNO, SMPNO, SUBNO) %>%
    select(RECID:NOIMP, SUBNO, NOIMP, BULKID, everything())
  
  # If  x has unmatched sets of joining values:
  full_join(
    data_10 %>% distinct(STNNO, SMPNO, SUBNO) %>% mutate(tab10 = 1),
    data_04 %>% distinct(STNNO, SMPNO, SUBNO) %>% mutate(tab04 = 1)
  ) %>%
    filter(is.na(tab10) | is.na(tab04))
}

```


### Check 5 - join 91 + 92
```{r}

check <- data_91 %>%
  safe_left_join(data_92, by = c("CRUIS", "STNNO"), 
                 na_matches = "never", check = "VN")

cat("Checked station table (91) vs station information table (92). If no error, check has passed.")

if (FALSE){
  data_04 %>%
    group_by(CRUIS, STNNO, SMPNO, SUBNO) %>%
    mutate(n = n()) %>%
    filter(n > 1)
}

```


### Check 6 - Missing values  

#### a. Number of missing values per variable
```{r}

# Missing values
cat("Missing values: \n----------------------------------------------------------\n")
cat("Table 04\n")
apply(is.na(data_04), 2, sum)
cat("Table 10\n")
apply(is.na(data_10), 2, sum)
cat("\n")
```

#### b. Missing uncertainties
```{r}

cat("Missing uncertainties: \n----------------------------------------------------------\n")
xtabs(~is.na(UNCRT) + addNA(QFLAG), data_10)

cat("\n\n")
cat("Missing uncertainties (data above LOQ) by parameter: \n----------------------------------------------------------\n")
data_10 %>% 
  mutate(Status = case_when(
    !is.na(QFLAG) ~ "Under LOQ",
    !is.na(UNCRT) ~ "Over LOQ, has UNCRT",
    is.na(UNCRT) ~ "Over LOQ, lacks UNCRT")
  ) %>%
  xtabs(~PARAM + Status, .)

```

#### c. Missing units
```{r}
cat("Missing units: \n----------------------------------------------------------\n")
check <- data_10 %>%
  filter(is.na(MUNIT)) 
cat(nrow(check), "records lack MUNIT \n")

if (nrow(check) > 0){
  cat("Missing MUNIT - Parameters: \n")
  xtabs(~PARAM, check)
  stop("All records must have MUNIT! (Check 6c)")
}

  
```


### Check 7
```{r}

# Check GOSOI (gonado-somatic index) for EROD samples
check <- data_10 %>%
  filter(PARAM %in% c("EROD", "GOSOI")) %>%
  select(STNNO, SMPNO, SUBNO, PARAM, VALUE) %>%
  spread(PARAM, VALUE)

percent_ok <- mean(!is.na(check$GOSOI))*100

cat(percent_ok, "% of the EROD values have gonado-somatic index (GOSOI) \n")

if (percent_ok < 100){
  warning("GOSOI is needed for all EROD values! (Check 7)")
}

```


### Check 8
```{r}
  
if (FALSE){

    check1 <- data_10 %>% left_join(data_04[,2:6])
  nrow(data_10) == nrow(check1)
  sum(is.na(data_04$NOIMP))
  sum(is.na(check1$NOIMP))
  
  check2 <- data_04 %>% left_join(data_10[,2:6])
  nrow(data_04) == nrow(check2)  # false but that is ok
  sum(is.na(data_10$MATRX))
  sum(is.na(check2$MATRX))
  
  xtabs(~STNNO + MATRX, subset(check1, is.na(NOIMP)))
  xtabs(~PARAM + MATRX, subset(check1, is.na(NOIMP) & STNNO == "13B"))
  xtabs(~SUBNO + MATRX, subset(check1, is.na(NOIMP) & STNNO == "13B"))
  
  xtabs(~STNNO, subset(data_03, STNNO %in% c("227G", "227G2")))
  xtabs(~STNNO, subset(data_04, STNNO %in% c("227G", "227G2")))
  xtabs(~STNNO + PARAM, subset(data_10, STNNO %in% c("227G", "227G2")))
  xtabs(~STATION_CODE + PARAM, subset(data_conc, STATION_CODE %in% c("227G", "227G2")))
  xtabs(~STATION_CODE + PARAM, subset(data_conc, STATION_CODE %in% c("97A2", "97A3")))
  
  subset(check1, STNNO == "13B") %>% View()
  
  subset(data_04, STNNO == "13B")
  subset(data_04, STNNO == "13B") %>% xtabs(~SUBNO, .)
  
  subset(data_conc_fishmuscle, STATION_CODE == "13B") %>% View()
  subset(data_conc_fishliver, STATION_CODE == "13B") %>% View()
  
  subset(data_conc_fishmuscle, STATION_CODE == "13B") %>% xtabs(~SUBNO, .)
  subset(data_conc_fishliver, STATION_CODE == "13B") %>% xtabs(~SUBNO, .)
  
  subset(data_conc_fish, STATION_CODE == "13B") %>% xtabs(~SUBNO, .)
  subset(data_10, STNNO == "13B") %>% xtabs(~SUBNO, .)
  
  subset(data_04, STNNO == "13B") %>% xtabs(~SUBNO, .)
  subset(data_10, STNNO == "13B") %>% xtabs(~SUBNO, .)
  subset(data_10_fish, STNNO == "13B") %>% xtabs(~SUBNO, .)
  subset(data_10_fatdry, STNNO == "13B") %>% xtabs(~SUBNO, .)
  subset(data_10_length2, STNNO == "13B") %>% xtabs(~SUBNO, .)
}

# xtabs(~PARAM + is.na(UNCRT), data_10)
```


### Write ICES submission file    
- automatically sets version number (finds the highest existing version number and adds one)  
- Set up to use UTF-8 when writing, but apparently we need to change the file manually to UTF-8 afterwards anyway (in Notepad++, set encoding to ANSI, then "convert to UTF-8")  
```{r}

write_ices_file <- FALSE
 write_ices_file <- TRUE

#
# Set "scipen" options so we avoid writing scientific notation (almost always)
# See ?write.table
#
cat("Set the way numbers are written as characters: \n")
cat("Existing setting: \n")
as.character(0.000004)   # check
old_options <- options(scipen = 3)
cat("New setting: \n")
as.character(0.000004)   # check again
cat("\n\n")


if (write_ices_file){

  if (!dir.exists(save_folder_txt)){
    message("Creating folder '", save_folder_txt, "'")
    dir.create(save_folder_txt)
  }
  
  fns <- dir(save_folder_txt, paste0("NIVA", selected_year, "CF*"))
  existing_versions <- substr(fns, 12, 13) %>% as.numeric()
  
  if (length(existing_versions) == 0){
    version <- 1
  } else {
    version <- max(existing_versions) + 1
  }
  
  string <- paste0("NIVA", selected_year, "CF_%02i.NO")
  fn <- sprintf(string, version)
  icesfile <- file(paste0(save_folder_txt, "/", fn), "w+")
  write.table(data_00, icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(data_03, icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(data_04, icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(data_10, icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(data_20, icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(data_21, icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(data_90, icesfile, sep = ";", quote = FALSE, na = "",
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(data_91, icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(data_92, icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  close(icesfile)
  cat("Version", version, "written as text file at", as.character(Sys.time()))

  } 

# Reset options
options(old_options)

```

### Save as R data
```{r}

if (write_ices_file){
  

  if (!dir.exists(save_folder_rds)){
    message("Creating folder '", save_folder_rds, "'")
    dir.create(save_folder_rds)
  }
  
  datalist <- list(data_00, data_03, data_04, data_10, data_20, 
                   data_21, data_90, data_91, data_92)
  names(datalist) <- c("00", "03", "04", "10", "20", "21", "90", "91", "92")
  
  string <- paste0("842_NIVA", selected_year, "CF_%02i.rds")
  fn_rds <- sprintf(string, version)
  
  saveRDS(datalist, paste0(save_folder_rds, "/", fn_rds))

  cat("Version", version, "written as rds (R data) file at", as.character(Sys.time()), "\n")
  cat("Filename: ",  paste0(save_folder_rds, "/", fn_rds), "\n")
  
}

if (FALSE){
  
  # For reading file
  datalist <- readRDS("Files_to_ICES/2020/Rdata/842_NIVA2020CF_09.rds")
  
}

```


### Save as Excel file  
```{r}

if (FALSE){
  
  library(writexl)
  string <- paste0("842_NIVA", selected_year, "CF_%02i.xlsx")
  fn_xl <- sprintf(string, version)
  
  write_xlsx(datalist, paste0(save_folder_excel, "/", fn_xl))

  cat("Version", version, "written as excel file at", as.character(Sys.time()), "\n")
  cat("Filename: ",  paste0(save_folder_rds, "/", fn_rds), "\n")
  
}

```


## Post-submission


### Restore all tables from saved version  
NOTE:   
- Versions 1-36 are written using this script  
- Versions 38-40 are written using script `26_Fix_submitted_ICES_file`   
    - The order of the tables in the latter rds files ( a list) differs form the order in vesrion 1-36     
    - In these versions, 
- (Version 37 doesn't exist, at least not as rds file)
```{r}
# Set this to true for restoring the tables 
# NOTE: will overwrite existing data_XX objects
restore <- FALSE
# restore <- TRUE

  
# Pick version (if needed)

if (restore) {
  
  
  # Find newest rds file
  fns <- dir(save_folder_rds)
  existing_versions <- substr(fns, 16, 17) %>% as.numeric()
  if (length(existing_versions) == 0){
    version <- 1
  } else {
    version <- max(existing_versions)
  }
  
  cat("Reading newest version - 'version' =", version, "\n")
  
  # Make file names
  
  string <- paste0("842_NIVA", selected_year, "CF_%02i.rds")
  fn_rds <- sprintf(string, version)
  string <- paste0("NIVA", selected_year, "CF_%02i.NO")
  fn <- sprintf(string, version)
  
  data_list <- readRDS(paste0(save_folder_rds, "/", fn_rds))
  names(data_list) <- c("00", "03", "04", "10", "20", "21", "90", "91", "92")
  
  data_00 <- data_list[["00"]]
  data_03 <- data_list[["03"]]
  data_04 <- data_list[["04"]]   # sample table
  data_10 <- data_list[["10"]]   # concentration data
  data_20 <- data_list[["20"]]
  data_21 <- data_list[["21"]]
  data_90 <- data_list[["90"]]
  data_91 <- data_list[["91"]]   # station data
  data_92 <- data_list[["92"]]
  
}

```

## Checks of data  

### Check positions     
- ICES position strings getting the warning 'Station position appears to be on land'   
- Checked using norgeskart.no  
```{r}

# Turns ICES position string into decimal longitude, latitude
copy_decimal <- function(string){ 
  str2 <- strsplit(string, ",")[[1]]
  y1 <- strsplit(str2[1], " ")[[1]] %>% as.numeric()
  y <- y1[1] + y1[2]/60
  x1 <- strsplit(str2[2], " ")[[1]] %>% as.numeric()
  x <- x1[1] + x1[2]/60
  paste0(round(y, 5), ",", round(x,5)) %>% writeLines("clipboard")
}

# ICES position strings getting the warning 'Station position appears to be on land'   
# Run line by line
# For each line, paste into search box of norgeskart.no
"+59 53.737,+05 6.514" %>% copy_decimal()
"+69 39.180,+18 58.440" %>% copy_decimal()
"+79 0.240,+12 6.600" %>% copy_decimal()

```


### Check specific lines in data
```{r}

check <- FALSE
 check <- TRUE
if (check) {
  dat <- readLines(fn_txt)  
  # Here, check which stations which "appears to be on land"
  lines <- c(1178, 1236, 1273, 1319, 1365)
  colnumber <- c(3,6,9,15)
  dat[lines] %>%
    strsplit(split = ";") %>%
    map(~.[colnumber])
}


```

### Check SHIPC references    
Mail from Anna (accessions@ices.dk) 4. sept 2019:  
Due to some critical errors, it is presently impossible to upload the NIVA 2018 CF file to DOME.
The SHIPC references in all records but record 90 are erroneous, and don’t match the ‘parent’ 90.SHIPC code.  
```{r}

# Which tables contain SHIPC?
sel <- data %>% map_lgl(~("SHIPC" %in% colnames(.)))
names(data)[sel] 

# Which tables contain CRUIS?
sel <- data %>% map_lgl(~("CRUIS" %in% colnames(.)))
names(data)[sel] 


data_90$CRUIS %>% unique() %>% sort()
data_91$CRUIS %>% unique() %>% sort()
data_03$CRUIS %>% unique() %>% sort()
data_04$CRUIS %>% unique() %>% sort()
data_10$CRUIS %>% unique() %>% sort()

if (FALSE){
  data_90
}
```

### Check STNNO references    
Mail from Anna (accessions@ices.dk) 12. sept 2019: 
'There is actually a problem with a missing record 91 for the STNNO=97A3 that is used in other record types'.  
Also lacking from record 92 but that doesnæt matter (only contains bottom temperatures)      
```{r}

# Which tables contain STNNO?
sel <- data %>% map_lgl(~("STNNO" %in% colnames(.)))
names(data)[sel] 

# Checks
for (station_code in c("97A2", "97A3")){
  cat(station_code, "\n")
  for (table in list(data_03, data_04, data_10, data_91, data_92)){
    cat(station_code %in% table$STNNO, "")
  }
  cat("\n")
}


```

### Check DRYWT 
```{r}

# Check parameters per matrix (excluding LI, MU)
df <- data_10 %>%
  filter(!MATRX %in% c("LI","MU")) %>%
  group_by(MATRX) %>%
  summarise(PARAM = paste(unique(PARAM), collapse = ", ")) %>%
  as.data.frame()
for (i in seq_len(nrow(df)))
  cat(df[i,1], "\n", df[i,2], "\n\n")


# Count samples
data_10 %>%
  count(STNNO, SMPNO, SUBNO, MATRX) %>% 
  nrow()   # 1116 samples

# Count measurements per matrix
data_10 %>%
  count(MATRX)

# Percent of samples with dry weight, by matrix 
data_10 %>%
  group_by(STNNO, SMPNO, SUBNO, MATRX) %>% 
  summarise(Drywt_n = sum(PARAM %in% "DRYWT%"), .groups = "drop") %>%    # count(Drywt_n)  # just checking
  group_by(MATRX) %>%  
  summarise(Drywt_perc = 100*mean(Drywt_n), N = n(), .groups = "drop") 
  # Lacking for 23% of SB samples
  # Lacking for all:
  #   cod biol. effect parameters (MATRX = BI, BL, LIS9) 
  #   gonad somatic index GOSOI (MATRX = GO)
  #   fish length LNMEA and weight WTMEA (MATRX = WO)

# Percent of "SB" samples with dry weight, BY STATION 
data_10 %>%
  group_by(STNNO, SMPNO, SUBNO, MATRX) %>% 
  summarise(Drywt_n = sum(PARAM %in% "DRYWT%")) %>%    # ungroup() %>% count(Drywt_n)  # just checking
  group_by(STNNO, MATRX) %>%
  summarise(Drywt_perc = 100*mean(Drywt_n), N = n(), .groups = "drop") %>%
  filter(Drywt_perc < 100 & MATRX %in% "SB")
  # 19N ONLY (Eider duck)

```

### Check a parameter, inc.. uncertainty     
```{r}

param <- "TBSN+"
metcu <- "SD"     # standard deviation  

if (FALSE){
  
  unique(data[["10"]]$PARAM) %>% sort()
  head(data[["10"]])

  }

data[["10"]] %>%
  filter(PARAM %in% param & (METCU %in% metcu | is.na(METCU))) %>% # View()
  ggplot(aes(STNNO, VALUE, color = QFLAG)) +
  geom_pointrange(aes(ymin = VALUE - UNCRT, ymax = VALUE + UNCRT)) +
  facet_wrap(vars(MATRX))


```


## Notes  


```

"BDE119", "BDE138", "BDE156", "BDE17", "BDE184", "BDE191", "BDE196", 
"BDE197", "BDE202", "BDE206", "BDE207", "BDE49", "BDE66", 
"BDE71", "BDE77", "BDE85", 
"CB105", "CB114", "CB122", "CB123", "CB128", "CB141", "CB149", "CB156", 
"CB157", "CB167", "CB170", "CB18", "CB183", "CB187", "CB189", 
"CB194", "CB206", "CB209", "CB31", "CB33", "CB37", "CB47", "CB66", 
"CB74", "CB99", 

"MCCP eksl. LOQ" ~ "MCCP", 
"SCCP eksl. LOQ" ~ "SCCP",
"Dieldrin" ~ "DIELD", 
"Mirex" = "MIREX", 
"Oktaklorstyren (OCS)" ~ "OCS",
"Oxyklordan" ~ "OCDAN", 

"Nonaklor, trans-" ~ "TNONC",
"alfa-Klordan (cis)" ~ "CCDAN"

"DDTEP" ~ "DDTEP", 
"DDTOP" ~ "DDTOP", 
"PROTV" ~ "PROTV"

"Tributyltinn (TBT)-Sn" ~ "TBTIN"
"MBTIN" ~ "MBSN+",

 
"D4", "D5", "D6", 
D4: octamethylcyclotetrasiloxane	556-67-2	
D5: decamethylcyclopentasiloxane	541-02-6	
D6: dodecamethylcyclohexasiloxane	540-97-6	


# "Heptaklor epoksid"
# only cis-heptachlorepoxide "(alpha)" and trans-heptachlorepoxide "(beta)"	


HCEPT 	 

 	trans-heptachlorepoxide "(beta)" 
, "MCCP eksl. LOQ", 
"Toksafen Parlar 50", "Toksafen Parlar 26", "Toksafen Parlar 62", - only TTOX, total toxaphene

"DDDOP" - not existing 

"Oxyklordan", "SCCP eksl. LOQ", "PROTV", "gamma-Klordan (trans)", 
"CB105", "CB128", "CB156", "CB167", "CB187", "CB99", "Sum-HepCB", 
"CB170", "CB74", "PeCB", "CB183", "Sum-HexCB", "BDE77", "CB209", 
"CB157", "Sum-PenCB", "CB114", , "CB123", 
"CB66", "Sum-TriCB", "BDE17", "KPAH", "P_S", "PAH16", "Sum-TetCB", 
"CB141", "DDEOP", "CB189", "CB194", "CB206", "CB31", "CB47", 
"Krysen", "MBTIN", "Monobutyltinn (MBT)-Sn", "CB149", "DBTIN", 
"Dibutyltinn-Sn (DBT-Sn)", "PFHxS", "QCB", "TBA", 
"TPTIN", "Trifenyltinn (TPhT)-Sn", "CB122", "BDE85", "Total PFOS/PFOA inkl. LOQ", 
"CB33", "CB37", "CB18", "Endrin", "BDE184", "BDE71", "HCHB", 
"INTF%")



"P_S", "PAH16", "PeCB", "Perfluor-3,7-dimetyloktansyre (PF37DMOA)", 
"Perfluorbutansyre (PFBA)", "Perfluordekansulfonat (PFDS)", "Perfluordodekansyre (PFDoA)", 
"Perfluorheptansulfonat (PFHpS)", "Perfluorpentansyre (PFPeA)", 
"Perfluortetradekansyre (PFTA)", "Perfluortridekansyre (PFTrA)", 
"PFAS", "PFHxS", "PROTV", "QCB", "SCCP eksl. LOQ", "Sum-HepCB", 
"Sum-HexCB", "Sum-PenCB", "Sum-TetCB", "Sum-TriCB", "TBA", "TCHT", 
"Tetrabutyltinn (TTBT)-Sn", "Toksafen Parlar 26", "Toksafen Parlar 50", 
"Toksafen Parlar 62", "Total PFOS/PFOA inkl. LOQ", "TPTIN", "trans-Heptaklorepoksid", 
"Tributyltinn (TBT)-Sn", "Trifenyltinn (TPhT)-Sn", "Trisykloheksyltinn (TCHT)-Sn", 
"TTBT")
> 
> cat("\n\nCheck excluded parameters:\n")

```