---
title: ""
author: "DHJ"
date: "2 1 2022"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

Following up data queries from Rob, found here  
`H:\seksjon 212\MIME 2021\Data_queries`    
from mails from Rob Fryer 21.11 and 13.12.2021  
- queries.doc          = Rob's original file - refers to file in 'dodgy_data_biota_211121'  
- queries_norway.doc   = DHJ's comments to Rob's file  



## 0. Library + functions  
Functions are from `"H:\Documents\seksjon 212\Milkys 2017\Analyse\21_read_ices_file.R"`
```{r}

# library(tidyverse)
library(dplyr)
library(purrr)
library(tidyr)     # pivot_wider
library(ggplot2)
source("845_Check_ICES_submission_functions.R")

```


## 1. Query 3 + 4. HCHA and HCHG at 36B, year 2000:     

### Filename
```{r}

fn <- "Files_to_ICES/Submission files 1981-2014/NIVA2000CF.NO"

```

### Read data
```{r}

data_2000 <- read_ices_file(fn, sep = ",") %>%   # in newer files, used semicolon
  add_field_codes() %>%
  set_numeric()

check_all_links(data_2000)

```

### Check w.w. conc vs. lipid weigth    
```{r}

df_plot <- data_2000[["10"]] %>% 
  filter(STNNO %in% "36B" & MATRX %in% "LI" & PARAM %in% c("EXLIP%", "HCHA")) %>%
  select(STNNO, SMPNO, SUBNO, PARAM, VALUE, QFLAG) %>%
  pivot_wider(names_from = PARAM, values_from = c(VALUE, QFLAG)) 

ggplot(df_plot, aes(`VALUE_EXLIP%`, VALUE_HCHA)) +
  geom_text(aes(label = SUBNO))

```

### Check all params vs liid weight    
- Note: MT = metallothionein   
```{r, fig.width=10, fig.height=8}

df_plot_allpar <- data_2000[["10"]] %>% 
  filter(STNNO %in% "36B" & MATRX %in% "LI") %>%
  select(STNNO, SMPNO, SUBNO, PARAM, VALUE, QFLAG)

df_plot_lip <- data_2000[["10"]] %>% 
  filter(STNNO %in% "36B" & MATRX %in% "LI" & PARAM %in% "EXLIP%") %>%
  select(STNNO, SMPNO, SUBNO, VALUE) %>%
  rename(Lip_perc = VALUE)

df_plot_allpar <- df_plot_allpar %>%
  left_join(df_plot_lip)

ggplot(df_plot_allpar, aes(Lip_perc, VALUE)) +
  geom_point(data = df_plot_allpar %>% filter(SUBNO ==4), shape = 1, size = 3, color = "red") +
  geom_text(aes(label = SUBNO), size = 3) +
  facet_wrap(vars(PARAM), scales = "free_y")

```

### Lines to delete from file   
- We keep metals and fat, they seem normal
- note that MT = metallothionein (we keep that one too)     
```{r}

data_to_exclude <- data_2000[["10"]] %>% 
  filter(
    STNNO %in% "36B" & MATRX %in% "LI" & SUBNO == 4,        # specify sample
    nchar(PARAM) != 2 & !PARAM %in% c("EXLIP%", "FATWT%")   # specify non-metals, and not fat
    ) %>%
  select(STNNO, SMPNO, SUBNO, PARAM, VALUE, QFLAG, Line_no) 

data_to_exclude


```
### Delete lines and write edited version   
```{r}

fn_orig <- "Files_to_ICES/Submission files 1981-2014/NIVA2000CF.NO"
fn_new <- "Files_to_ICES/Submission files 1981-2014/NIVA2000CF_edit_2022-01-03.NO"

data_raw <- readLines(fn_orig) # 1990

# Check
# data_raw[11117]

data_raw_fixed <- data_raw[-pull(data_to_exclude, Line_no)]

# writeLines(data_raw_fixed, con = fn_new)

```


### Check 1 - read and check links  

```{r}

data_fixed <- read_ices_file(fn_new, sep = ",") %>%
  add_field_codes() %>%
  set_numeric()

check_all_links(data_fixed)

```
### Check 2 - check data values  
```{r}

stationcode <- "36B"
matrix <- "LI"

df_plot_allpar <- data_fixed[["10"]] %>%
  filter(STNNO %in% stationcode & MATRX %in% matrix) %>%
  select(STNNO, SMPNO, SUBNO, PARAM, VALUE, QFLAG)

df_plot_lip <- data_fixed[["10"]] %>%
  filter(STNNO %in% stationcode & MATRX %in% matrix & PARAM %in% c("EXLIP%", "FATWT%")) %>%
  select(STNNO, SMPNO, SUBNO, VALUE) %>%
  rename(Lip_perc = VALUE)

df_plot_allpar <- df_plot_allpar %>%
  left_join(df_plot_lip)

ggplot(df_plot_allpar, aes(Lip_perc, VALUE)) +
  geom_point(data = df_plot_allpar %>% filter(SUBNO == 4), shape = 1, size = 3, color = "red") +
  geom_text(aes(label = SUBNO), size = 3) +
  facet_wrap(vars(PARAM), scales = "free_y")


```




## 2. Query 2. 2016. Chromium concentrations in cod liver at station 98B1 Austnesfjord, Lofoten  

Note: Same high concentrations for the following stations:
- "23B Karihavet area", "24B Bergen harbour", "53B Inner Sørfjord" 
       

### Filename
```{r}

fn <- "Files_to_ICES/Submission files 2015-2019/NIVA2016CF.NO"

```

### Read data
```{r}

# readLines(fn, 3)

data_2016_raw <- read_ices_file(fn, sep = ";")

length(data_2016_raw)

# debugonce(set_numeric)
data_2016 <- data_2016_raw %>%   # comma in older files, semicolon in newer
  add_field_codes() %>%
  set_numeric()

# We set 'print_values = TRUE' as there were some missing 
check_all_links(data_2016, print_values = TRUE)

```

### Check values     
```{r}

selected_stations <- c("98B1", "23B", "24B", "53B")
matrix <- "LI"
params <- "CR"

df_plot <- data_2016[["10"]] %>% 
  filter(MATRX %in% matrix & PARAM %in% params) %>%
  # filter(STNNO %in% stations) %>%
  select(STNNO, SMPNO, SUBNO, PARAM, VALUE, QFLAG) %>%
  mutate(
    Selected = STNNO %in% selected_stations
  )

gg <- ggplot(df_plot, aes(STNNO, VALUE, color = Selected)) +
  geom_jitter(width = 0.2)

gg
gg + scale_y_log10()

```

### Lines to delete from file   
- We keep metals and fat, they seem normal
- note that MT = metallothionein (we keep that one too)     
```{r}

data_to_exclude <- data_2016[["10"]] %>% 
  filter(MATRX %in% matrix & PARAM %in% params) %>%
  filter(STNNO %in% stations) %>%
  select(STNNO, SMPNO, SUBNO, PARAM, VALUE, QFLAG, Line_no) 

data_to_exclude


```

### Delete lines and write edited version   
```{r}

fn_orig <- "Files_to_ICES/Submission files 2015-2019/NIVA2016CF.NO"
fn_new <- "Files_to_ICES/Submission files 2015-2019/NIVA2016CF_edit_2022-01-03_temp.NO"

# we call it 'temp' because we will delete 3 rows from table 03, see results of 'check_all_links'   

data_raw <- readLines(fn_orig) 

# Check
# data_raw[11117]

data_raw_fixed <- data_raw[-pull(data_to_exclude, Line_no)]

# writeLines(data_raw_fixed, con = fn_new)

```


### Check 1 - read and check links  
- The file 'NIVA2016CF_edit_2022-01-03_temp' is manually edited to 'NIVA2016CF_edit_2022-01-03' based on this   
- Specifically: deleting "Line numbers: 85,90,98"   
- NOTE: when deleting lines based on line numbers, *start with the last number!*!  (line numbers will be messed up if you start with the first... )   
- See check 3 below

```{r}

data_fixed <- read_ices_file(fn_new, sep = ";") %>%   # remember sep value
  add_field_codes() %>%
  set_numeric()

# debugonce(check_link)
check_all_links(data_fixed, print_values = TRUE)

```

### Check 2 - check data values    
```{r}

# Code copied from above

selected_stations <- c("98B1", "23B", "24B", "53B")
matrix <- "LI"
params <- "CR"

df_plot <- data_fixed[["10"]] %>% 
  filter(MATRX %in% matrix & PARAM %in% params) %>%
  # filter(STNNO %in% stations) %>%
  select(STNNO, SMPNO, SUBNO, PARAM, VALUE, QFLAG) %>%
  mutate(
    Selected = STNNO %in% selected_stations
  )

gg <- ggplot(df_plot, aes(STNNO, VALUE, color = Selected)) +
  geom_jitter(width = 0.2)

gg
gg + scale_y_log10()

```


### Check 3 - check manual edits   
- Reading and checking manually fixed file (3 lnes deleted from part '03' in file, see Check 1)   
```{r}

fn_new2 <- "Files_to_ICES/Submission files 2015-2019/NIVA2016CF_edit_2022-01-03.NO"

data_fixed2 <- read_ices_file(fn_new2, sep = ";") %>%   # remember sep value
  add_field_codes() %>%
  set_numeric()

# debugonce(check_link)
check_all_links(data_fixed2, print_values = TRUE)

```


## 3. Queries 2022: 1. 1996 BDE47 and BD153 very high     

The PBDE concentrations from 1996 are dubiously high so delete them. Here is a graph of concentrations by PBDE and year (across all stations) with the 1996 data that are to be deleted plotted in red.


### Filename
```{r}

fn <- "Files_to_ICES/Submission files 1981-2014/NIVA1996CF_utf8.NO"

```

### Read data
```{r}

# readLines(fn, 3)

check_sep <- readLines(fn, 1)
comma <- gregexpr(",", check_sep, fixed = TRUE)[[1]] %>% length()
semicolon <- gregexpr(";", check_sep, fixed = TRUE)[[1]]

if (comma > semicolon){
  # debugonce(read_ices_file)
  dat_raw <- read_ices_file(fn, sep = ",")
} else {
  dat_raw <- read_ices_file(fn, sep = ";")
}

length(dat_raw)

# debugonce(set_numeric)
dat <- dat_raw %>%   # comma in older files, semicolon in newer
  add_field_codes() %>%
  set_numeric()

# We set 'print_values = TRUE' as there were some missing 
check_all_links(dat, print_values = TRUE)

```

### Check values     
```{r}

# selected_stations <- c("98B1", "23B", "24B", "53B")
matrix <- "LI"
params <- c("BDE47", "BD153")

df_plot <- dat[["10"]] %>% 
  filter(MATRX %in% matrix & PARAM %in% params) %>%
  # filter(STNNO %in% stations) %>%
  select(STNNO, SMPNO, SUBNO, PARAM, VALUE, QFLAG) # %>%
  # mutate(
  #   Selected = STNNO %in% selected_stations
  # )

gg <- ggplot(df_plot, aes(STNNO, VALUE, color = PARAM)) +
  geom_jitter(width = 0.2)

gg
gg + scale_y_log10()

```

### Lines to delete from file   
- We keep metals and fat, they seem normal
- note that MT = metallothionein (we keep that one too)     
```{r}

data_to_exclude <- dat[["10"]] %>% 
  filter(MATRX %in% matrix & PARAM %in% params) 

data_to_exclude

# Since we delete all these params, we also need to delete some methods
# params <- c("BDE47", "BD153")

table(data_to_exclude$AMLNK)
table(dat[["10"]]$AMLNK)

data_to_exclude2 <- dat[["21"]] %>% 
  filter(AMLNK %in% unique(data_to_exclude$AMLNK))  

```

### Delete lines and write edited version   
```{r}

fn_orig <- "Files_to_ICES/Submission files 1981-2014/NIVA1996CF_utf8.NO"
fn_new <- "Files_to_ICES/Submission files 1981-2014/NIVA1996CF_utf8_edit_2022-10-20.NO"

# we call it 'temp' because we will delete 3 rows from table 03, see results of 'check_all_links'   

data_raw <- readLines(fn_orig) 

# Check
# data_raw[11117]

lines_to_remove <- c(
  pull(data_to_exclude, Line_no),
  pull(data_to_exclude2, Line_no)
)

data_raw_fixed <- data_raw[-lines_to_remove]

# Uncomment to write:
# writeLines(data_raw_fixed, con = fn_new)

```


### Check 1 - read and check links   

- Discovered here that we also needed to remove two lines from table 21 (methods)

```{r}

data_fixed <- read_ices_file(fn_new, sep = ",") %>%   # remember sep value
  add_field_codes() %>%
  set_numeric()

# debugonce(check_link)
check_all_links(data_fixed, print_values = TRUE)

```

### Check 2 - check data values    
```{r}

# Code copied from above, except that dat is replaced by 'data_fixed'

matrix <- "LI"
params <- c("BDE47", "BD153")

df_plot <- data_fixed[["10"]] %>% 
  filter(MATRX %in% matrix & PARAM %in% params) %>%
  # filter(STNNO %in% stations) %>%
  select(STNNO, SMPNO, SUBNO, PARAM, VALUE, QFLAG)

nrow(df_plot)

```


## 3. Queries 2022: 1. 1996 BDE47 and BD153 very high     

The PBDE concentrations from 1996 are dubiously high so delete them. Here is a graph of concentrations by PBDE and year (across all stations) with the 1996 data that are to be deleted plotted in red.


### Filename
```{r}

fn <- "Files_to_ICES/Submission files 1981-2014/NIVA1996CF_utf8.NO"

```

### Read data
```{r}

# readLines(fn, 3)

check_sep <- readLines(fn, 1)
comma <- gregexpr(",", check_sep, fixed = TRUE)[[1]] %>% length()
semicolon <- gregexpr(";", check_sep, fixed = TRUE)[[1]]

if (comma > semicolon){
  # debugonce(read_ices_file)
  dat_raw <- read_ices_file(fn, sep = ",")
} else {
  dat_raw <- read_ices_file(fn, sep = ";")
}

length(dat_raw)

# debugonce(set_numeric)
dat <- dat_raw %>%   # comma in older files, semicolon in newer
  add_field_codes() %>%
  set_numeric()

# We set 'print_values = TRUE' as there were some missing 
check_all_links(dat, print_values = TRUE)

```

### Check values     
```{r}

# selected_stations <- c("98B1", "23B", "24B", "53B")
matrix <- "LI"
params <- c("BDE47", "BD153")

df_plot <- dat[["10"]] %>% 
  filter(MATRX %in% matrix & PARAM %in% params) %>%
  # filter(STNNO %in% stations) %>%
  select(STNNO, SMPNO, SUBNO, PARAM, VALUE, QFLAG) # %>%
  # mutate(
  #   Selected = STNNO %in% selected_stations
  # )

gg <- ggplot(df_plot, aes(STNNO, VALUE, color = PARAM)) +
  geom_jitter(width = 0.2)

gg
gg + scale_y_log10()

```

### Lines to delete from file   
- We keep metals and fat, they seem normal
- note that MT = metallothionein (we keep that one too)     
```{r}

data_to_exclude <- dat[["10"]] %>% 
  filter(MATRX %in% matrix & PARAM %in% params) 

data_to_exclude

# Since we delete all these params, we also need to delete some methods
# params <- c("BDE47", "BD153")

table(data_to_exclude$AMLNK)
table(dat[["10"]]$AMLNK)

data_to_exclude2 <- dat[["21"]] %>% 
  filter(AMLNK %in% unique(data_to_exclude$AMLNK))  

```

### Delete lines and write edited version   
```{r}

fn_orig <- "Files_to_ICES/Submission files 1981-2014/NIVA1996CF_utf8.NO"
fn_new <- "Files_to_ICES/Submission files 1981-2014/NIVA1996CF_utf8_edit_2022-10-20.NO"

# we call it 'temp' because we will delete 3 rows from table 03, see results of 'check_all_links'   

data_raw <- readLines(fn_orig) 

# Check
# data_raw[11117]

lines_to_remove <- c(
  pull(data_to_exclude, Line_no),
  pull(data_to_exclude2, Line_no)
)

data_raw_fixed <- data_raw[-lines_to_remove]

# Uncomment to write:
# writeLines(data_raw_fixed, con = fn_new)

```


### Check 1 - read and check links   

- Discovered here that we also needed to remove two lines from table 21 (methods)

```{r}

data_fixed <- read_ices_file(fn_new, sep = ",") %>%   # remember sep value
  add_field_codes() %>%
  set_numeric()

# debugonce(check_link)
check_all_links(data_fixed, print_values = TRUE)

```

### Check 2 - check data values    

```{r}

# Code copied from above, except that dat is replaced by 'data_fixed'

matrix <- "LI"
params <- c("BDE47", "BD153")

df_plot <- data_fixed[["10"]] %>% 
  filter(MATRX %in% matrix & PARAM %in% params) %>%
  # filter(STNNO %in% stations) %>%
  select(STNNO, SMPNO, SUBNO, PARAM, VALUE, QFLAG)

nrow(df_plot)

```
