---
title: ""
author: "DHJ"
date: "2 1 2022"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

Following up data queries from Rob, found here  
`H:\seksjon 212\MIME 2021\Data_queries`    
from mails from Rob Fryer 21.11 and 13.12.2021  
- queries.doc          = Rob's original file - refers to file in 'dodgy_data_biota_211121'  
- queries_norway.doc   = DHJ's comments to Rob's file  



## 0. Library + functions  
Functions are from `"H:\Documents\seksjon 212\Milkys 2017\Analyse\21_read_ices_file.R"`
```{r}

# library(tidyverse)
library(dplyr)
library(purrr)
library(tidyr)     # pivot_wider
library(ggplot2)
source("845_Check_ICES_submission_functions.R")

```


## 1. ICES file version 01

```{r}

fn <- "Files_to_ICES/2021/Test/NIVA2021CF_01.NO"

```


### Read data
```{r}

# readLines(fn, 3)

dat_txt <- readLines(fn)

# comma in older files, semicolon in newer - but this is now checked by 'read_ices_file'  
dat_raw <- read_ices_file(fn)

# debugonce(set_numeric)
dat <- dat_raw %>%   
  add_field_codes() %>%
  set_numeric()

```

### Check all links  
```{r}
# We set 'print_values = TRUE' as there were some missing 
check_all_links(dat, print_values = TRUE)

```


### PARAM bile metabolites require reporting of DETLI    

- Found to be 1 - we set it to 1  
- 135	warning  
    - Cross field warning check (condition not met)  
    - Line 747,750,753,756 PARAM bile metabolites require reporting of DETLI (WGBEC2013)  - most common
    - Line 11569, LIPIDWT% or FATWT% should be >= 1% for fish - 6 cases

#### Check  
```{r}

# example
dat_txt[c(747,750,753,756)]

# Check all QFLAG
# dat[["10"]] %>% xtabs(~addNA(QFLAG) + PARAM, .)

# Check  
dat[["10"]] %>%
  filter(PARAM %in% c("BAP3OH", "PYR1OH", "PA1OH")) %>%
  group_by(PARAM, QFLAG) %>%
  summarise(
    Min = min(VALUE),
    Max = max(VALUE),
  )

# Plot  
gg <- dat[["10"]] %>%
  filter(PARAM %in% c("BAP3OH", "PYR1OH", "PA1OH")) %>% 
  ggplot(aes(PARAM, VALUE)) + 
  geom_violin() +
  facet_grid(vars(QFLAG))

gg
gg + scale_y_log10()

```
#### Fix  
```{r}

sel <- with(dat[["10"]], PARAM %in% c("BAP3OH", "PYR1OH", "PA1OH"))
sum(sel)

dat[["10"]]$DETLI[sel] <- 1

```

### VALUE of VDSI must be > 0 if METCU is "%"  

- conclusion: this is an erronous check, the percentage is actually zero
- 8 error
    - Cross field check (error preventing data import)
    - Line 11491,11492,11497,11502 etc.: VALUE must be > 0 if METCU is "%"	 
    - 8 lines  

#### Check        
```{r}

# example
dat_txt[c(11491,11492,11497,11502)]

# Check  
dat[["10"]] %>%
  filter(PARAM %in% "VDSI" & VALUE == 0)

```

### Sampling coordinates are not in the station dictionary defined area for this station  

#### Check   
- 
```{r}

# 71A Bjørkøya (Risøyodden),+78 16.560,+15 24.360  


dat[["91"]] %>%
  filter(grepl("71A", STATN)) %>%
  select(STNNO, LATIT, LONGI, STATN)
# 2 lines, one for the actual 71A, the other for STNNO = 20B 

# - latitude/longitude for the last one is ok - checked below
# +78 16.560	+15 24.360
lat <- 78 + 16.560/60
lon <- 15 + 24.360/60

if (FALSE){
  # interactive map - copy-paste to the R console
  library(leaflet)
  leaflet() %>%
    addTiles() %>%
    addMarkers(lng = lon, lat = lat)
}


```

#### Fix
```{r}

sel <- with(dat[["91"]], grepl("71A", STATN) & STNNO == "20B")
sum(sel)

dat[["91"]]$STATN[sel] <- "20B Longyearbyen"

```

### Table 20 has one row without corresponding SMLNK values in table 3   

- See "Check all links" above   
- No warning from DATSU, but we fix it anyway  

#### Check  
```{r}

# From "Check all links" above

# *** 03 (sample occasion per station) and 20 (sampling methods) ***
# Variables used for join: SMLNK 
# Table 1 has 52 rows (0 rows with missing values in join variables)
# Table 2 has 3 rows (0 rows with missing values in join variables)
# All variable combinations in table 1 exist in table 2
# 
# Table 2 has 1 rows (1 variable combinations) without corresponding rows in table 1
# Line numbers: 12806 
#   Number of SMLNK values in missing rows: 1
#     - 9
  
xtabs(~SMLNK, dat[["03"]])

# xtabs(~SMLNK + STNNO, dat[["03"]])

```

#### Fix  
```{r}

# Delete BOT (bottom trawling)  
# It doesn't matter for the results, and we donæt know which fihsermen have used it anyway  
sel1 <- dat[["20"]]$SMTYP %in% "BOT"
dat[["20"]] <- dat[["20"]][!sel1,]

# Change SMLNK (link reference to table 03) to 8  
sel2 <- dat[["20"]]$SMTYP %in% "GIL"
dat[["20"]]$SMLNK[sel2] <- "8"

nrow(dat[["20"]])

```


### Check all links again  

```{r}

check_all_links(dat, print_values = TRUE)

```

### Write ICES submission file    
- automatically sets version number (finds the highest existing version number and adds one)  
- Set up to use UTF-8 when writing, but apparently we need to change the file manually to UTF-8 afterwards anyway (in Notepad++, set encoding to ANSI, then "convert to UTF-8")  
```{r}

write_ices_file <- FALSE
 write_ices_file <- TRUE

#
# Set "scipen" options so we avoid writing scientific notation (almost always)
# See ?write.table
#
cat("Set the way numbers are written as characters: \n")
cat("Existing setting: \n")
as.character(0.000004)   # check
old_options <- options(scipen = 3)
cat("New setting: \n")
as.character(0.000004)   # check again
cat("\n\n")


if (write_ices_file){

  if (!dir.exists(save_folder_txt)){
    message("Creating folder '", save_folder_txt, "'")
    dir.create(save_folder_txt)
  }
  
  fns <- dir(save_folder_txt, paste0("NIVA", selected_year, "CF*"))
  existing_versions <- substr(fns, 12, 13) %>% as.numeric()
  
  if (length(existing_versions) == 0){
    version <- 1
  } else {
    version <- max(existing_versions) + 1
  }
  message("Version number to write: ", version)
  
  # Back up
  dat_with_linenumbers <- dat

  # Remove Line_no
  for (i in seq_along(dat)){
    dat[[i]]$Line_no <- NULL
  }
  
  string <- paste0("NIVA", selected_year, "CF_%02i.NO")
  fn <- sprintf(string, version)
  icesfile_name <- paste0(save_folder_txt, "/", fn)
  icesfile <- file(icesfile_name, "w+")
  write.table(dat[["00"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["03"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["04"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["10"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["20"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["21"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["90"]], icesfile, sep = ";", quote = FALSE, na = "",
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["91"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["92"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  close(icesfile)
  cat("Version", version, "written as text file at", as.character(Sys.time()))

  } 

# Reset options
options(old_options)

```

### Read the file we just wrote and check all links yet again  

```{r}

# dat_txt <- readLines(icesfile_name)
# dat_txt[497]

dat <- read_ices_file(icesfile_name) %>%   
  add_field_codes() %>%
  set_numeric()

check_all_links(dat, print_values = TRUE)

```


### Save as R data
```{r}

if (write_ices_file){
  

  if (!dir.exists(save_folder_rds)){
    message("Creating folder '", save_folder_rds, "'")
    dir.create(save_folder_rds)
  }
  
  datalist <- dat

  string <- paste0("842_NIVA", selected_year, "CF_%02i.rds")
  fn_rds <- sprintf(string, version)
  
  saveRDS(datalist, paste0(save_folder_rds, "/", fn_rds))

  cat("Version", version, "written as rds (R data) file at", as.character(Sys.time()), "\n")
  cat("Filename: ",  paste0(save_folder_rds, "/", fn_rds), "\n")
  
}

if (FALSE){
  
  # For reading file
  datalist <- readRDS("Files_to_ICES/2020/Rdata/842_NIVA2020CF_09.rds")
  
}

```


### Save as Excel file  
```{r}

if (FALSE){
  
  library(writexl)
  string <- paste0("842_NIVA", selected_year, "CF_%02i.xlsx")
  fn_xl <- sprintf(string, version)
  
  write_xlsx(datalist, paste0(save_folder_excel, "/", fn_xl))

  cat("Version", version, "written as excel file at", as.character(Sys.time()), "\n")
  cat("Filename: ",  paste0(save_folder_rds, "/", fn_rds), "\n")
  
}

```


## 2. ICES file version 04

```{r}

fn <- "Files_to_ICES/2021/Submitted/NIVA2021CF_04.NO"

```

### Read data
```{r}

# readLines(fn, 3)

dat_txt <- readLines(fn)
# dat_txt[497]

# comma in older files, semicolon in newer - but this is now checked by 'read_ices_file'  
dat_raw <- read_ices_file(fn)

# debugonce(set_numeric)
dat <- dat_raw %>%   
  add_field_codes() %>%
  set_numeric()

```

### LIPIDWT% or FATWT% should be >= 1% for fish   

- Conclusion: this is not liver/muscle for fish, but blood from bird  
- 25 errors
- Line 11569,11645,11711 etc. "LIPDWT% or FATWT% should be >= 1% for fish"  


#### Check  
```{r}

# Example lines - we see that this is blood samples, probably from birds    
dat_txt[c(11569,11645,11711)]

# Table 10 plus added info including from table 03 
nrow(dat[["10"]])
check <- dat[["10"]] %>%
  left_join(dat[["03"]] %>% select(-RECID, -Line_no))  # important to remove these, so they are not used in the join
nrow(check)

# Confirm that this is eider duck blood  
check %>%
  filter(MATRX %in% "BL" & PARAM %in% "FATWT%") %>% # View("check")
  xtabs(~SPECI, .)

```

