---
title: ""
author: "DHJ"
date: "2 1 2022"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

Following up data queries from Rob, found here  
`H:\seksjon 212\MIME 2021\Data_queries`    
from mails from Rob Fryer 21.11 and 13.12.2021  
- queries.doc          = Rob's original file - refers to file in 'dodgy_data_biota_211121'  
- queries_norway.doc   = DHJ's comments to Rob's file  



## 0. Library + functions  
Functions are from `"H:\Documents\seksjon 212\Milkys 2017\Analyse\21_read_ices_file.R"`
```{r}

# library(tidyverse)
library(dplyr)
library(purrr)
library(tidyr)     # pivot_wider
library(ggplot2)
source("845_Check_ICES_submission_functions.R")

```


## 1. ICES file version 01

```{r}

fn <- "Files_to_ICES/2021/Test/NIVA2021CF_01.NO"

```


### Read data
```{r}

# readLines(fn, 3)

dat_txt <- readLines(fn)

# comma in older files, semicolon in newer - but this is now checked by 'read_ices_file'  
dat_raw <- read_ices_file(fn)

# debugonce(set_numeric)
dat <- dat_raw %>%   
  add_field_codes() %>%
  set_numeric()

```

### Check all links  
```{r}
# We set 'print_values = TRUE' as there were some missing 
check_all_links(dat, print_values = TRUE)

```


### PARAM bile metabolites require reporting of DETLI    

- Found to be 1 - we set it to 1  
- 135	warning  
    - Cross field warning check (condition not met)  
    - Line 747,750,753,756 PARAM bile metabolites require reporting of DETLI (WGBEC2013)  - most common
    - Line 11569, LIPIDWT% or FATWT% should be >= 1% for fish - 6 cases

#### Check  
```{r}

# example
dat_txt[c(747,750,753,756)]

# Check all QFLAG
# dat[["10"]] %>% xtabs(~addNA(QFLAG) + PARAM, .)

# Check  
dat[["10"]] %>%
  filter(PARAM %in% c("BAP3OH", "PYR1OH", "PA1OH")) %>%
  group_by(PARAM, QFLAG) %>%
  summarise(
    Min = min(VALUE),
    Max = max(VALUE),
  )

# Plot  
gg <- dat[["10"]] %>%
  filter(PARAM %in% c("BAP3OH", "PYR1OH", "PA1OH")) %>% 
  ggplot(aes(PARAM, VALUE)) + 
  geom_violin() +
  facet_grid(vars(QFLAG))

gg
gg + scale_y_log10()

```
#### Fix  
```{r}

sel <- with(dat[["10"]], PARAM %in% c("BAP3OH", "PYR1OH", "PA1OH"))
sum(sel)

dat[["10"]]$DETLI[sel] <- 1

```

### VALUE of VDSI must be > 0 if METCU is "%"  

- conclusion: this is an erronous check, the percentage is actually zero
- 8 error
    - Cross field check (error preventing data import)
    - Line 11491,11492,11497,11502 etc.: VALUE must be > 0 if METCU is "%"	 
    - 8 lines  

#### Check        
```{r}

# example
dat_txt[c(11491,11492,11497,11502)]

# Check  
dat[["10"]] %>%
  filter(PARAM %in% "VDSI" & VALUE == 0)

```

### Sampling coordinates are not in the station dictionary defined area for this station  

#### Check   
- 
```{r}

# 71A Bjørkøya (Risøyodden),+78 16.560,+15 24.360  


dat[["91"]] %>%
  filter(grepl("71A", STATN)) %>%
  select(STNNO, LATIT, LONGI, STATN)
# 2 lines, one for the actual 71A, the other for STNNO = 20B 

# - latitude/longitude for the last one is ok - checked below
# +78 16.560	+15 24.360
lat <- 78 + 16.560/60
lon <- 15 + 24.360/60

if (FALSE){
  # interactive map - copy-paste to the R console
  library(leaflet)
  leaflet() %>%
    addTiles() %>%
    addMarkers(lng = lon, lat = lat)
}


```

#### Fix
```{r}

sel <- with(dat[["91"]], grepl("71A", STATN) & STNNO == "20B")
sum(sel)

dat[["91"]]$STATN[sel] <- "20B Longyearbyen"

```

### Table 20 has one row without corresponding SMLNK values in table 3   

- See "Check all links" above   
- No warning from DATSU, but we fix it anyway  

#### Check  
```{r}

# From "Check all links" above

# *** 03 (sample occasion per station) and 20 (sampling methods) ***
# Variables used for join: SMLNK 
# Table 1 has 52 rows (0 rows with missing values in join variables)
# Table 2 has 3 rows (0 rows with missing values in join variables)
# All variable combinations in table 1 exist in table 2
# 
# Table 2 has 1 rows (1 variable combinations) without corresponding rows in table 1
# Line numbers: 12806 
#   Number of SMLNK values in missing rows: 1
#     - 9
  
xtabs(~SMLNK, dat[["03"]])

# xtabs(~SMLNK + STNNO, dat[["03"]])

```

#### Fix  
```{r}

# Delete BOT (bottom trawling)  
# It doesn't matter for the results, and we donæt know which fihsermen have used it anyway  
sel1 <- dat[["20"]]$SMTYP %in% "BOT"
dat[["20"]] <- dat[["20"]][!sel1,]

# Change SMLNK (link reference to table 03) to 8  
sel2 <- dat[["20"]]$SMTYP %in% "GIL"
dat[["20"]]$SMLNK[sel2] <- "8"

nrow(dat[["20"]])

```


### Check all links again  

```{r}

check_all_links(dat, print_values = TRUE)

```

### Write ICES submission file    
- automatically sets version number (finds the highest existing version number and adds one)  
- Set up to use UTF-8 when writing, but apparently we need to change the file manually to UTF-8 afterwards anyway (in Notepad++, set encoding to ANSI, then "convert to UTF-8")  
```{r}

write_ices_file <- FALSE
 write_ices_file <- TRUE

#
# Set "scipen" options so we avoid writing scientific notation (almost always)
# See ?write.table
#
cat("Set the way numbers are written as characters: \n")
cat("Existing setting: \n")
as.character(0.000004)   # check
old_options <- options(scipen = 3)
cat("New setting: \n")
as.character(0.000004)   # check again
cat("\n\n")


if (write_ices_file){

  if (!dir.exists(save_folder_txt)){
    message("Creating folder '", save_folder_txt, "'")
    dir.create(save_folder_txt)
  }
  
  fns <- dir(save_folder_txt, paste0("NIVA", selected_year, "CF*"))
  existing_versions <- substr(fns, 12, 13) %>% as.numeric()
  
  if (length(existing_versions) == 0){
    version <- 1
  } else {
    version <- max(existing_versions) + 1
  }
  message("Version number to write: ", version)
  
  # Back up
  dat_with_linenumbers <- dat

  # Remove Line_no
  for (i in seq_along(dat)){
    dat[[i]]$Line_no <- NULL
  }
  
  string <- paste0("NIVA", selected_year, "CF_%02i.NO")
  fn <- sprintf(string, version)
  icesfile_name <- paste0(save_folder_txt, "/", fn)
  icesfile <- file(icesfile_name, "w+")
  write.table(dat[["00"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["03"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["04"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["10"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["20"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["21"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["90"]], icesfile, sep = ";", quote = FALSE, na = "",
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["91"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["92"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  close(icesfile)
  cat("Version", version, "written as text file at", as.character(Sys.time()))

  } 

# Reset options
options(old_options)

```

### Read the file we just wrote and check all links yet again  

```{r}

# dat_txt <- readLines(icesfile_name)
# dat_txt[497]

dat <- read_ices_file(icesfile_name) %>%   
  add_field_codes() %>%
  set_numeric()

check_all_links(dat, print_values = TRUE)

```


### Save as R data
```{r}

if (write_ices_file){
  

  if (!dir.exists(save_folder_rds)){
    message("Creating folder '", save_folder_rds, "'")
    dir.create(save_folder_rds)
  }
  
  datalist <- dat

  string <- paste0("842_NIVA", selected_year, "CF_%02i.rds")
  fn_rds <- sprintf(string, version)
  
  saveRDS(datalist, paste0(save_folder_rds, "/", fn_rds))

  cat("Version", version, "written as rds (R data) file at", as.character(Sys.time()), "\n")
  cat("Filename: ",  paste0(save_folder_rds, "/", fn_rds), "\n")
  
}

if (FALSE){
  
  # For reading file
  datalist <- readRDS("Files_to_ICES/2020/Rdata/842_NIVA2020CF_09.rds")
  
}

```


### Save as Excel file  
```{r}

if (FALSE){
  
  library(writexl)
  string <- paste0("842_NIVA", selected_year, "CF_%02i.xlsx")
  fn_xl <- sprintf(string, version)
  
  write_xlsx(datalist, paste0(save_folder_excel, "/", fn_xl))

  cat("Version", version, "written as excel file at", as.character(Sys.time()), "\n")
  cat("Filename: ",  paste0(save_folder_rds, "/", fn_rds), "\n")
  
}

```


## 2. ICES file version 04

```{r}

fn <- "Files_to_ICES/2021/Submitted/NIVA2021CF_04.NO"

```

### Read data
```{r}

# readLines(fn, 3)

dat_txt <- readLines(fn)
# dat_txt[497]

# comma in older files, semicolon in newer - but this is now checked by 'read_ices_file'  
dat_raw <- read_ices_file(fn)

# debugonce(set_numeric)
dat <- dat_raw %>%   
  add_field_codes() %>%
  set_numeric()

```

### LIPIDWT% or FATWT% should be >= 1% for fish   

- Conclusion: this is not liver/muscle for fish, but blood from bird  
- 25 errors
- Line 11569,11645,11711 etc. "LIPDWT% or FATWT% should be >= 1% for fish"  


#### Check  
```{r}

# Example lines - we see that this is blood samples, probably from birds    
dat_txt[c(11569,11645,11711)]

# Table 10 plus added info including from table 03 
nrow(dat[["10"]])
check <- dat[["10"]] %>%
  left_join(dat[["03"]] %>% select(-RECID, -Line_no))  # important to remove these, so they are not used in the join
nrow(check)

# Confirm that this is eider duck blood  
check %>%
  filter(MATRX %in% "BL" & PARAM %in% "FATWT%") %>% # View("check")
  xtabs(~SPECI, .)

```
## 3. Submit imposex on individual basis   

- using data for individual snails (parameter VDS) instead of the mean for all snails per station (parameter VDSI)

### a. Read data
```{r}

# readLines(fn, 3)

fn <- "Files_to_ICES/2021/Test/NIVA2021CF_04.NO"

dat_txt <- readLines(fn)
# dat_txt[497]

# comma in older files, semicolon in newer - but this is now checked by 'read_ices_file'  
dat_raw <- read_ices_file(fn)

# debugonce(set_numeric)
dat <- dat_raw %>%   
  add_field_codes() %>%
  set_numeric()

```
#### Check all links  

```{r}

check_all_links(dat, print_values = TRUE)

```

### b. Check existing snail data  
```{r}

table(dat[["03"]]$SPECI)  

species <- c("Nucella lapillus", "Littorina littorea")
# species <- "Mytilus edulis"

df1 <- dat[["03"]] %>%
  filter(SPECI %in% species)
nrow(df1)

df2 <- dat[["04"]] %>%
  left_join(dat[["03"]] %>% select(-RECID, -Line_no)) %>%
  filter(SPECI %in% species)
nrow(df2)

dat_existing <- dat[["10"]] %>%
  left_join(dat[["04"]] %>% select(-RECID, -Line_no)) %>%
  left_join(dat[["03"]] %>% select(-RECID, -Line_no)) %>%
  filter(SPECI %in% species)
nrow(dat_existing)

xtabs(~PARAM + SUBNO, dat_existing)

```


### c. Data for input (from raw data)  

```{r, message=FALSE}

folder_vdsi <- "K:/Prosjekter/Sjøvann/JAMP/2021/opparbeiding/Snegl"
folder_vdsi <- "Input_files_2021/Snegl"

#
# 1. Imposex in Littorina
#

fn <- dir(folder_vdsi, pattern = ".xlsx")
fn <- fn[!grepl("imposex_mal", fn)]
fn <- fn[!grepl("intersex", fn)]          # skip intersex for now
fn <- fn[!grepl("~", fn, fixed = TRUE)]

# Names = Extract all characters to the left of the first underscore  
names(fn) <- stringr::str_extract(fn, "[^_]+")
# names(fn)[names(fn) == "131"] <- "131G"
# fn


read_excel_snail <- function(fn){
  df <- readxl::read_excel(paste0(folder_vdsi, "/", fn), skip = 6)
  df <- df[1:6]
  df <- df[!is.na(df$`Analysert av`),]
  df
}

dat_vdsi_list <- fn %>% map(read_excel_snail)

# read_excel_snail_s <- safely(read_excel_snail)
# dat_vdsi_list <- fn %>% map(read_excel_snail_s)



# Check number of rows - should be 50
#  unless oherwise noted in sheet (15G "Fant ikke flere snegl i år, men det er over 15 hunner.")
dat_vdsi_list %>% map_int(nrow)

# Check 15G
# dat_vdsi_list[["15G"]] %>% View()
# dat_vdsi_list[["36G"]] %>% View()
# dat_vdsi_list[["15G"]] <- dat_vdsi_list[["15G"]][1:50,]  # delete last line

# Pick only number, sex and Imposex 
# ...we will deal with length and penis length later  
dat_vdsi <- dat_vdsi_list %>% 
  map(~.[c(1,3,5)]) %>%
  bind_rows(.id = "STATION_CODE") %>%
  rename(VALUE_WW = VDSI, 
         Sex = `Kj<U+00F8>nn`) %>%
  mutate(PARAM = "VDSI",
         UNIT = "idx",
         LATIN_NAME = "Nucella lapillus",
         Sex = tolower(Sex))

# View(dat_vdsi)

cat("\n------------------------------------------\n")
cat("Imposex data: \n------------------------------------------\n")
xtabs(~is.na(VALUE_WW) + addNA(Sex), dat_vdsi)


#
# 2. Intersex
#
# dir(folder_vdsi)
fn <- "71G_fugloy_intersex_LITTORINA_2021.xlsx"
df <- readxl::read_excel(paste0(folder_vdsi, "/", fn), range = "A3:G53")

dat_intersex <- df %>%
  filter(F == 1) %>%
  mutate(STATION_CODE = "71G", 
         Sex = "f",
         PARAM = "Intersex",
         UNIT = "PERCENT",
         LATIN_NAME = "Littorina littorea") %>%
  mutate(VALUE_WW = `ISI   (intersex-stadie)`*100) %>%
  select(STATION_CODE, Sex, VALUE_WW, PARAM, LATIN_NAME)



```

### d. New rows for tables 04, 10 and 21     

- data for table 04 (station*species), 10 (measurements) and 21 methods     

```{r}

# View(dat[["04"]])
# View(dat[["10"]])
# View(dat[["21"]])

table(dat_vdsi$Sex)

xtabs(~CRUIS + STNNO, dat[["04"]])  
# Snails are in "Cruise number 2"  

# Imposex, all "G" stations except 76G:
dat_for_insert1 <- dat_vdsi %>%
  filter(Sex %in% "f") %>%
  rename(STNNO = STATION_CODE) %>%
  mutate(
    SMPNO = 3,               # dog whelk fixed number, see table 03
    SUBNO = Løpenr + 1,     # to avoid SUBNO=1, which already is "reserved" for RYWT and TBTIN   
    STNNO = ifelse(STNNO == "227G", "227G2", STNNO),
    PARAM = "VDS"
  )

# Intersex, 76G:
dat_for_insert2 <- dat_intersex %>%
  rename(STNNO = STATION_CODE) %>%
  mutate(
    SMPNO = 3,                          # dog whelk fixed number, see table 03
    SUBNO = 1:nrow(dat_intersex) + 1,   # to avoid SUBNO=1, which already is "reserved" for RYWT and TBTIN
    STNNO = ifelse(STNNO == "227G", "227G2", STNNO),
    PARAM = "INTS"
  )

dat_for_insert <- bind_rows(dat_for_insert1, dat_for_insert2)

# Only data for 

max(dat[["10"]]$AMLNK)

dat_new_04 <- dat_for_insert %>%
  mutate(
    RECID = "04",
    CRUIS = "2",
    NOIMP = "1",
    SEXCO = "F"
  ) %>%
  select(
    RECID, CRUIS, STNNO, SMPNO, SUBNO, NOIMP, SEXCO
  )


dat_new_10 <- dat_for_insert %>%
  mutate(
    RECID = "10",
    CRUIS = "2",
    MATRX = "WO",
    MUNIT = "ST",     # Effects parameters VDS and IMPS should have integer range 0-6 with MUNIT ST (WGBEC2010)
    BASIS = "W",
    AMLNK = max(dat[["10"]]$AMLNK) + 1,      # new reference number in table 21    
    VALUE = VALUE_WW
  ) %>%
  select(
    RECID, CRUIS, STNNO, SMPNO, SUBNO, MATRX, PARAM, MUNIT, BASIS, AMLNK, VALUE
  )

# For table 21 
# test <- dat[["21"]] %>%
#   filter(METOA %in% "VISO") 
# dput(test[1,])

dat_new_21 <- structure(
  list(
    RECID = "21", 
    AMLNK = max(dat[["10"]]$AMLNK) + 1, 
    ALABO = "NIVA", METDC = "", REFSK = "", METST = "", METFP = "NON", METPT = "", METCX = "", 
    METPS = "", METOA = "VISO", AGDET = "", SREFW = "", SPECI = "", 
    RLIST = "", ORGSP = "", SIZRF = "", FORML = "", Line_no = NA), 
  row.names = 1L, class = "data.frame")


```
#### Check  
```{r}

# CRUIS, STNNO, SMPNO, SUBNO

xtabs(~CRUIS, dat_new_10 )
xtabs(~CRUIS, dat_existing)

xtabs(~addNA(STNNO), dat_new_10 )
xtabs(~addNA(STNNO), dat_existing)

xtabs(~addNA(SMPNO), dat_new_10 )
xtabs(~addNA(SMPNO), dat_existing)

xtabs(~addNA(SUBNO), dat_new_10 )
xtabs(~addNA(SUBNO), dat_existing)
xtabs(~addNA(SUBNO), dat[["04"]] %>% filter(STNNO == "11G"))


```

### e. New data 1: remove rows    

```{r}

# New data set
dat_new <- dat

#
# Remove rows from table 04 - ot needed (only one line per station)
#


#
# Remove rows from table 10
#

nrow(dat_new[["10"]])

dat_new[["10"]] <- dat_new[["10"]] %>%
  filter(!PARAM %in% "VDSI")

nrow(dat_new[["10"]])

#
# Remove rows from table 21
#

amlnk <- dat[["10"]] %>%
  filter(PARAM %in% "VDSI") %>% 
  pull(AMLNK) %>%
  unique()

nrow(dat_new[["21"]])
dat_new[["21"]] <- dat_new[["21"]] %>%
  filter(!AMLNK %in% amlnk)
nrow(dat_new[["21"]])

```
#### Check  
```{r}

cat("3 vs. 4 (old / new) \n")
cols <- var_overlap(dat[["03"]], dat[["04"]], drop = c("RECID", "Line_no"))
check_link(dat[["03"]], dat[["04"]], cols)
cols <- var_overlap(dat[["03"]], dat[["04"]], drop = c("RECID", "Line_no"))
check_link(dat[["03"]], dat_new[["04"]], cols, print_values = TRUE)

cat("\n04 vs. 10 (old / new) \n")
cols <- var_overlap(dat[["04"]], dat[["10"]], drop = c("RECID", "Line_no"))
check_link(dat[["04"]], dat[["10"]], cols, print_values = TRUE)
cols <- var_overlap(dat[["04"]], dat_new[["10"]], drop = c("RECID", "Line_no"))
check_link(dat[["04"]], dat_new[["10"]], cols, print_values = TRUE)

# View(dat[["04"]])

```

### f. New data 2: add rows     
```{r}

dat_new[["04"]] <- bind_rows(
  dat_new[["04"]],
  dat_new_04
)

dat_new[["10"]] <- bind_rows(
  dat_new[["10"]],
  dat_new_10
)

dat_new[["21"]] <- bind_rows(
  dat_new[["21"]],
  dat_new_21
)

nrow(dat[["10"]])
nrow(dat_new[["10"]])

nrow(dat[["21"]])
nrow(dat_new[["21"]])

```

#### Checking 
```{r}

cat("Number of rows, original + new: \n")
dat %>% map_int(nrow)
dat_new %>% map_int(nrow)

cat("\nCheck of links, new: \n\n")
check_all_links(dat_new)  



```

#### Checks  
```{r}

# table(addNA(dat_new[["04"]]$SMPNO))
# table(addNA(dat_new[["04"]]$STNNO))
# table(addNA(dat_new_04$STNNO))

```


### g. Write ICES submission file    
- automatically sets version number (finds the highest existing version number and adds one)  
- Set up to use UTF-8 when writing, but apparently we need to change the file manually to UTF-8 afterwards anyway (in Notepad++, set encoding to ANSI, then "convert to UTF-8")  
```{r}

write_ices_file <- FALSE
 write_ices_file <- TRUE
 
save_folder_txt <- "Files_to_ICES/2021/Test"
selected_year <- 2021

#
# Set "scipen" options so we avoid writing scientific notation (almost always)
# See ?write.table
#
cat("Set the way numbers are written as characters: \n")
cat("Existing setting: \n")
as.character(0.000004)   # check
old_options <- options(scipen = 3)
cat("New setting: \n")
as.character(0.000004)   # check again
cat("\n\n")


if (write_ices_file){

  if (!dir.exists(save_folder_txt)){
    message("Creating folder '", save_folder_txt, "'")
    dir.create(save_folder_txt)
  }
  
  fns <- dir(save_folder_txt, paste0("NIVA", selected_year, "CF*"))
  existing_versions <- substr(fns, 12, 13) %>% as.numeric()
  
  if (length(existing_versions) == 0){
    version <- 1
  } else {
    version <- max(existing_versions) + 1
  }
  message("Version number to write: ", version)
  
  # Back up
  dat_with_linenumbers <- dat_new

  # Remove Line_no
  for (i in seq_along(dat_new)){
    dat_new[[i]]$Line_no <- NULL
  }
  
  string <- paste0("NIVA", selected_year, "CF_%02i.NO")
  fn <- sprintf(string, version)
  icesfile_name <- paste0(save_folder_txt, "/", fn)
  icesfile <- file(icesfile_name, "w+")
  write.table(dat_new[["00"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat_new[["03"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat_new[["04"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat_new[["10"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat_new[["20"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat_new[["21"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat_new[["90"]], icesfile, sep = ";", quote = FALSE, na = "",
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat_new[["91"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat_new[["92"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  close(icesfile)
  cat("Version", version, "written as text file at", as.character(Sys.time()))

  } 

# Reset options
options(old_options)

```

### h. Read the file we just wrote and check all links yet again  

```{r}

# icesfile_name <- "Files_to_ICES/2021/Test/NIVA2021CF_06.NO"

dat_check <- read_ices_file(icesfile_name) %>%   
  add_field_codes() %>%
  set_numeric()

if (FALSE){
  table(addNA(dat_new[["04"]]$STNNO))
  table(addNA(dat_new[["10"]]$STNNO))
  table(addNA(dat_check[["10"]]$STNNO))
}

check_all_links(dat_check, print_values = TRUE)

```

#### Checking if errors  
```{r}

if (FALSE){

xtabs(~CRUIS, dat_check[["03"]])
xtabs(~CRUIS, dat_check[["04"]])
xtabs(~CRUIS + STNNO, dat_check[["03"]])

xtabs(~SMPNO, dat_check[["03"]])

check <- dat_check[["04"]] %>%
  filter(Line_no %in% 497:522)

check2 <- dat_check[["04"]] %>%
  filter(Line_no %in% 497:522) %>%
  left_join(dat_check[["03"]] %>% select(CRUIS, STNNO, SMPNO ) %>% mutate(TEST = 1))
table(addNA(check2$TEST))
      

View(check2)  
# CRUIS, STNNO, SMPNO 

}

```


### Check all links again  

```{r}

check_all_links(dat, print_values = TRUE)

```

### Write ICES submission file    
- automatically sets version number (finds the highest existing version number and adds one)  
- Set up to use UTF-8 when writing, but apparently we need to change the file manually to UTF-8 afterwards anyway (in Notepad++, set encoding to ANSI, then "convert to UTF-8")  
```{r}

write_ices_file <- FALSE
 write_ices_file <- TRUE

#
# Set "scipen" options so we avoid writing scientific notation (almost always)
# See ?write.table
#
cat("Set the way numbers are written as characters: \n")
cat("Existing setting: \n")
as.character(0.000004)   # check
old_options <- options(scipen = 3)
cat("New setting: \n")
as.character(0.000004)   # check again
cat("\n\n")


if (write_ices_file){

  if (!dir.exists(save_folder_txt)){
    message("Creating folder '", save_folder_txt, "'")
    dir.create(save_folder_txt)
  }
  
  fns <- dir(save_folder_txt, paste0("NIVA", selected_year, "CF*"))
  existing_versions <- substr(fns, 12, 13) %>% as.numeric()
  
  if (length(existing_versions) == 0){
    version <- 1
  } else {
    version <- max(existing_versions) + 1
  }
  message("Version number to write: ", version)
  
  # Back up
  dat_with_linenumbers <- dat

  # Remove Line_no
  for (i in seq_along(dat)){
    dat[[i]]$Line_no <- NULL
  }
  
  string <- paste0("NIVA", selected_year, "CF_%02i.NO")
  fn <- sprintf(string, version)
  icesfile_name <- paste0(save_folder_txt, "/", fn)
  icesfile <- file(icesfile_name, "w+")
  write.table(dat[["00"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["03"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["04"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["10"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["20"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["21"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["90"]], icesfile, sep = ";", quote = FALSE, na = "",
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["91"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  write.table(dat[["92"]], icesfile, sep = ";", quote = FALSE, na = "", 
              col.names = FALSE, row.names = FALSE, fileEncoding = "UTF-8")
  close(icesfile)
  cat("Version", version, "written as text file at", as.character(Sys.time()))

  } 

# Reset options
options(old_options)

```

